% Unit 10: Nonhomogeneous Equations and Forced Response
% Condensed Cheat Sheet (Part I only, standalone)
% Prepared by Staples Education
% Content informed by the local curriculum modules (10.1 to 10.8) and the
% QUIZ_DATA micro-practice items for Unit 10.
% Shared style harness pulled from the unified preamble input.

\input{ode/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 10: Nonhomogeneous Equations and Forced Response}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Condensed Cheat Sheet: Particular Solutions and Forced Response}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 10 at a Glance: Core Sub-Topics]
A nonhomogeneous linear equation reads $L[y] = g(x)$, and its general solution
always splits as $y = y_h + y_p$ --- the complementary solution of $L[y]=0$ plus
\emph{any} particular solution of the forcing. This unit is the toolkit for
building $y_p$ and for reading how a system answers an input.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Undetermined coefficients} --- guess a trial of the same shape as
          the forcing, then match coefficients.
    \item \textbf{Variation of parameters} --- promote the constants $c_1,c_2$ to
          functions for \emph{any} continuous forcing.
    \item \textbf{The annihilator method} --- an operator that kills $g$ derives
          the trial form instead of memorizing it.
    \item \textbf{Transient and steady-state} --- the homogeneous part decays, the
          particular part persists.
    \item \textbf{Unit step and impulse response} --- $u(t)$, $\delta(t)$, and the
          system's answer to a unit kick from rest.
    \item \textbf{Duhamel's principle} --- convolve the impulse response with the
          input to answer \emph{every} forcing at once.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{10}   % subsections auto-number as the unit modules 10.1, 10.2, ...

\subsection{The Method of Undetermined Coefficients}

For a constant-coefficient $L[y] = g(x)$, the general solution is
$y = y_h + y_p$: the complementary solution $y_h$ solves $L[y]=0$ and carries the
arbitrary constants, while $y_p$ matches the forcing. When $g$ is an exponential,
polynomial, sine/cosine, or a product of these, guess a trial of the \emph{same
shape} and solve for its coefficients.

\begin{formulabox}[Complementary Plus Particular]
\[ y = y_h + y_p, \qquad L[y_h] = 0, \qquad L[y_p] = g(x). \]
Only $y_h$ holds the free constants, so it alone can absorb the initial or
boundary conditions --- which is why $y_p$ is never the whole story.
\end{formulabox}

\begin{formulabox}[Trial Solution Table (No Overlap)]
\[
\begin{array}{l|l}
\textbf{Forcing } g(x) & \textbf{Trial } y_p \\ \hline
e^{\alpha x} & A e^{\alpha x} \\
\text{polynomial of degree } n & A_n x^n + \cdots + A_1 x + A_0 \\
\cos\beta x \ \text{ or } \ \sin\beta x & A\cos\beta x + B\sin\beta x \\
e^{\alpha x}\cos\beta x & e^{\alpha x}(A\cos\beta x + B\sin\beta x) \\
\end{array}
\]
For a \emph{sum} of forcings, build a particular solution for each piece and add
them (superposition for the nonhomogeneous equation).
\end{formulabox}

\begin{warningbox}[Common Pitfall: Overlap with the Homogeneous Solution]
If any term of the trial already solves $L[y]=0$, substitution collapses it to
zero and the method fails. \textbf{Multiply the offending trial by $x$} (or by a
higher power $x^k$ for a root of multiplicity $k$) until no term overlaps $y_h$.
For $y''+y=\cos x$, the bare guess $A\cos x + B\sin x$ \emph{is} the homogeneous
solution; the corrected resonant trial is $x(A\cos x + B\sin x)$.
\end{warningbox}

\subsection{Variation of Parameters}

When the forcing is not one of the friendly shapes --- $\tan x$, $\sec x$,
$e^x/x$ --- undetermined coefficients has nothing to guess. Variation of
parameters promotes the constants $c_1,c_2$ of $y_h = c_1 y_1 + c_2 y_2$ to
functions $u_1(x),u_2(x)$ and works for \emph{any} continuous $g$.

\begin{formulabox}[Variation of Parameters Formulas]
Write the equation in standard form $y'' + p y' + q y = g$ (leading coefficient
one), let $W = y_1 y_2' - y_2 y_1'$ be the Wronskian, and set $y_p = u_1 y_1 + u_2 y_2$ with
\[ u_1' = -\frac{y_2\, g}{W}, \qquad u_2' = \frac{y_1\, g}{W}, \]
\[ y_p = -\,y_1\!\int \frac{y_2\, g}{W}\,dx \;+\; y_2\!\int \frac{y_1\, g}{W}\,dx. \]
\end{formulabox}

\begin{warningbox}[Common Pitfall: Forgetting to Normalize]
The $g$ in these formulas is the right-hand side \emph{after} dividing through so
the $y''$ coefficient is one. Reading $g$ off an unnormalized equation scales the
answer incorrectly. Always normalize first, then read $g$.
\end{warningbox}

\subsection{Differential Operators and the Annihilator Method}

Writing $D = \tfrac{d}{dx}$, a constant-coefficient operator factors like a
polynomial. An \textbf{annihilator} of $g$ is an operator $A(D)$ with $A(D)g = 0$.
Applying it to both sides of $L[y]=g$ turns the problem into a larger
\emph{homogeneous} equation whose extra solutions are exactly the trial form.

\begin{methodbox}[Annihilator Workflow]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item Factor $L(D)$ and write the complementary solution $y_h$.
    \item Pick the annihilator $A(D)$ of the forcing $g$.
    \item Apply $A(D)$: solve the homogeneous equation $A(D)L(D)\,y = 0$.
    \item Drop the terms already in $y_h$; the survivors are the trial $y_p$.
    \item Substitute into $L[y]=g$ and match coefficients.
\end{enumerate}
\end{methodbox}

\begin{formulabox}[Common Annihilators]
\[
\begin{array}{l|l}
\textbf{Function} & \textbf{Annihilator} \\ \hline
e^{\alpha x} & (D-\alpha) \\
x^{k-1} & D^{k} \\
\cos\beta x,\ \sin\beta x & (D^2 + \beta^2) \\
e^{\alpha x}\cos\beta x,\ e^{\alpha x}\sin\beta x & (D-\alpha)^2 + \beta^2 \\
\end{array}
\]
A repeated factor $(D-\alpha)^k$ signals the terms
$e^{\alpha x}, x e^{\alpha x}, \ldots, x^{k-1}e^{\alpha x}$. Annihilate a sum by
\emph{multiplying} the individual annihilators.
\end{formulabox}

\subsection{Transient and Steady-State Response}

For a \emph{stable} forced system (characteristic roots with negative real
parts), the complete solution $y = y_h + y_p$ separates by long-term behavior: the
homogeneous part fades while the particular part endures.

\begin{formulabox}[Transient Versus Steady-State]
\[ y(t) = \underbrace{y_h(t)}_{\text{transient}\ \to\ 0} \;+\; \underbrace{y_p(t)}_{\text{steady-state}}. \]
The \textbf{transient} carries the initial-condition dependence and decays because
its exponentials have negative real parts. The \textbf{steady-state} is fixed by
the forcing alone, so long after the start the response is essentially independent
of the initial data. For sinusoidal forcing the steady state is a sinusoid at the
\emph{drive} frequency, rescaled and phase-shifted, often read off the complex
gain $1/p(i\omega)$.
\end{formulabox}

\subsection{Unit Step and Impulse Response}

Two idealized inputs model switching and striking. The unit step $u(t)$ switches a
forcing on; the Dirac delta $\delta(t)$ is an instantaneous unit kick. They are
linked by calculus, and so are the responses they produce.

\begin{formulabox}[Step, Delta, and Their Responses]
\[ u(t) = \begin{cases} 0, & t<0 \\ 1, & t>0 \end{cases}, \qquad
   \delta(t) = u'(t), \qquad \int_{-\infty}^{\infty}\!\delta(t)\,dt = 1. \]
\textbf{Sifting property:} $\displaystyle\int_{-\infty}^{\infty} f(t)\,\delta(t-a)\,dt = f(a)$.
The \textbf{impulse response} $w(t)$ is the output of $L[y]=\delta(t)$ from rest;
the \textbf{step response} is its integral. Knowing $w$ unlocks every input by
convolution.
\end{formulabox}

\begin{warningbox}[Common Pitfall: Forgetting the Rest State]
The impulse response is defined from \textbf{rest} --- zero state just before the
kick. Starting with leftover energy mixes in a transient that is unrelated to the
impulse, so the clean $w(t)$ (here $w(0)=0$, $w'(0)=1$ for a second-order system)
is lost. Isolate the impulse by zeroing the initial conditions first.
\end{warningbox}

\subsection{Duhamel's Principle}

Duhamel's principle views a general input as a continuous succession of small
impulses. Each kick at time $\tau$ produces a scaled, time-shifted copy of the
impulse response, and superposition sums them into a convolution.

\begin{methodbox}[Duhamel's Convolution]
For a linear, time-invariant system starting from rest, the forced response to an
input $f(t)$ is
\[ y_p(t) = \int_0^t w(t-\tau)\,f(\tau)\,d\tau, \]
where $w$ is the impulse response (the Green's function of the initial-value
problem). The limits $0$ to $t$ encode \textbf{causality} --- only inputs already
applied can affect the present --- and the argument $t-\tau$ is the time elapsed
since each kick. \textbf{Linearity} and \textbf{time invariance} are exactly what
make the convolution valid.
\end{methodbox}

\begin{conceptbox}[Unit 10 Key Takeaway]
Every forced response is \textcolor{accentorange}{$y = y_h + y_p$}. Build $y_p$ by
\textcolor{accentorange}{undetermined coefficients} when the forcing is friendly
(remember to \textcolor{accentorange}{multiply by $x$} on overlap), by
\textcolor{accentorange}{variation of parameters} for anything continuous, or read
its form from an \textcolor{accentorange}{annihilator}. In a stable system the
\textcolor{accentorange}{transient} $y_h$ decays and the
\textcolor{accentorange}{steady-state} $y_p$ remains, and once the
\textcolor{accentorange}{impulse response $w$} is known,
\textcolor{accentorange}{Duhamel's convolution} answers every input at once.
\end{conceptbox}

\end{document}
