% Unit 7: First-Order Exactness and Methods
% Master Unit Reference Guide (Cheat Sheet, Problems, and Solutions)
% Prepared by Staples Education
% Content informed by the local curriculum modules (7.1 to 7.7) read from
% curriculum-data.js and the 10-problem master practice set for Unit 7.
% Shared style harness pulled from the unified preamble input.

\input{ode/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 7: First-Order Exactness and Methods}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Master Unit Reference Guide: Cheat Sheet, Problems, and Solutions}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Table of contents (master guide only)
% ---------------------------------------------------------------------------
\tableofcontents

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box (all 7 curriculum modules from curriculum-data.js)
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 7 at a Glance: Core Sub-Topics]
A first-order ODE written as $M\,dx + N\,dy = 0$ is \emph{exact} when its
left side is already the total differential $dF$ of some scalar potential $F$,
making the general solution simply $F(x,y)=C$. This unit develops the complete
toolkit: test the cross-partials to detect exactness, reconstruct the potential
via the integration loop, repair non-exact equations with special integrating
factors, connect the theory to conservative vector fields, and read the physics
through thermodynamic state functions.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{7.1 Exact Differential Equations} --- the criterion $M_y = N_x$
          and the implicit general solution $F(x,y)=C$.
    \item \textbf{7.2 Exact Equations Intuition} --- Clairaut's theorem:
          $F_{xy}=F_{yx}$ is why $M_y=N_x$ guarantees a potential.
    \item \textbf{7.3 Exact Equations Motivation and Theory} --- the integration
          loop: $F = \int M\,dx + g(y)$, fix $g'(y)$ by matching $F_y = N$.
    \item \textbf{7.4 Non-Exact Equations and Special Integrating Factors} ---
          integrating factor $\mu(x)$ when $\Delta/N$ is pure in $x$; $\mu(y)$
          when $-\Delta/M$ is pure in $y$ (where $\Delta = M_y - N_x$).
    \item \textbf{7.5 Almost Exact Differential Equations} --- decision tree:
          test both $\mu(x)$ and $\mu(y)$ routes systematically, then execute
          the repair and re-verify exactness before reconstructing $F$.
    \item \textbf{7.6 Exact Equations and Conservative Vector Fields} ---
          $M_y=N_x$ iff $\mathbf F=\langle M,N\rangle$ is conservative;
          path independence via the Fundamental Theorem for Line Integrals.
    \item \textbf{7.7 Applications in Thermodynamics} --- $dU$ is an exact
          state differential; $\delta Q$ is inexact; the integrating factor
          $1/T$ produces exact entropy $dS = \delta Q/T$.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{7}   % subsections auto-number as 7.1, 7.2, ...

% ---------------------------------------------------------------------------
\subsection{Exact Differential Equations}
% ---------------------------------------------------------------------------

A first-order ODE $M(x,y)\,dx + N(x,y)\,dy = 0$ is \emph{exact} on a simply
connected region when its left side equals the total differential $dF$ of some
potential function $F(x,y)$. The key theorem reduces exactness to a single
algebraic test on the cross-partial derivatives of $M$ and $N$.

\begin{formulabox}[Exactness Criterion and Implicit Solution]
The ODE $M\,dx + N\,dy = 0$ (with $M,N$ having continuous first partials) is
\textbf{exact} on a simply connected region if and only if
\[
  \frac{\partial M}{\partial y} = \frac{\partial N}{\partial x}.
\]
When exact, a potential $F(x,y)$ exists satisfying $F_x = M$ and $F_y = N$,
and the \textbf{general solution} is
\[
  F(x,y) = C.
\]
\end{formulabox}

\begin{warningbox}[Common Pitfall: Swapped Cross Partials]
Always differentiate $M$ with respect to \textbf{$y$} and $N$ with respect to
\textbf{$x$}. The test is $M_y = N_x$, never $M_x$ versus $N_y$. Computing the
same-variable partials tells you nothing about exactness and will produce a
false positive or false negative on every equation you test.
\end{warningbox}

% ---------------------------------------------------------------------------
\subsection{Exact Equations Intuition}
% ---------------------------------------------------------------------------

The cross-partial criterion is not an arbitrary rule --- it is Clairaut's
theorem in disguise. If $M = F_x$ and $N = F_y$ for some $F$, then
$M_y = F_{xy}$ and $N_x = F_{yx}$, and Clairaut guarantees the two mixed
partials agree whenever $F$ has continuous second partials. Exactness is
precisely the statement that $M$ and $N$ are partials of a common function.

\begin{formulabox}[Clairaut's Theorem and the Exact-Differential Connection]
On a simply connected region with $M,N$ continuously differentiable,
\[
  M\,dx + N\,dy = dF
  \iff
  \frac{\partial M}{\partial y}
  = \frac{\partial^{2}F}{\partial y\,\partial x}
  = \frac{\partial^{2}F}{\partial x\,\partial y}
  = \frac{\partial N}{\partial x}.
\]
The equality of mixed partials is both \textbf{necessary} (if $dF$ exists then
the cross partials must agree) and \textbf{sufficient} (on a simply connected
domain, equal cross partials guarantee a global potential).
\end{formulabox}

% ---------------------------------------------------------------------------
\subsection{Exact Equations Motivation and Theory}
% ---------------------------------------------------------------------------

Once exactness is confirmed, the potential $F$ is found by a two-step
integration loop. Integrating $M$ with respect to $x$ introduces an arbitrary
function $g(y)$ in place of a constant; that function is then pinned down by
the condition $F_y = N$.

\begin{methodbox}[Potential Function Reconstruction: The Integration Loop]
Given $M\,dx + N\,dy = 0$ confirmed exact by $M_y = N_x$:
\begin{enumerate}[leftmargin=1.6em, itemsep=3pt]
    \item \textbf{Integrate $M$ in $x$}, holding $y$ fixed:
          \[
            F(x,y) = \int M(x,y)\,dx + g(y).
          \]
    \item \textbf{Differentiate $F$ in $y$} and set the result equal to $N$:
          \[
            \frac{\partial F}{\partial y} = N(x,y)
            \quad\Longrightarrow\quad
            g'(y) = N(x,y) - \frac{\partial}{\partial y}\int M\,dx.
          \]
          (All $x$-terms must cancel; if they do not, re-check the exactness
          test.)
    \item \textbf{Integrate $g'(y)$} in $y$ to obtain $g(y)$.
    \item Write the \textbf{general solution} $F(x,y) = C$.
\end{enumerate}
\end{methodbox}

\begin{warningbox}[Common Pitfall: $g'(y)$ Must Be Free of $x$]
After forming $g'(y) = N - \partial_y\!\int M\,dx$, every $x$-term must
cancel. If $x$ survives in $g'(y)$, the equation was not exact (or an
arithmetic error occurred). Integrating an $x$-dependent expression as if it
were a function of $y$ alone produces a wrong answer without any algebra error
to signal the problem.
\end{warningbox}

% ---------------------------------------------------------------------------
\subsection{Non-Exact Equations and Special Integrating Factors}
% ---------------------------------------------------------------------------

When $M_y \neq N_x$, no potential $F$ exists for the equation as written.
Multiplying both sides by a non-zero \emph{integrating factor} $\mu$ can
change the cross partials of $\mu M$ and $\mu N$ until they match. Two
special forms of $\mu$ are algebraically testable and cover the majority of
non-exact equations encountered in practice.

\begin{formulabox}[Integrating Factors $\mu(x)$ and $\mu(y)$]
Let $\Delta = M_y - N_x$ (the \textbf{exactness defect}).
\begin{itemize}[leftmargin=1.4em, itemsep=4pt]
    \item \textbf{Test for $\mu(x)$:} Compute $\dfrac{\Delta}{N}$.
          If the result depends on $x$ \emph{only}, set $h(x)=\Delta/N$ and
          \[
            \mu(x) = \exp\!\left(\int h(x)\,dx\right).
          \]
    \item \textbf{Test for $\mu(y)$:} Compute $\dfrac{N_x - M_y}{M}
          = \dfrac{-\Delta}{M}$.
          If the result depends on $y$ \emph{only}, set $k(y)=-\Delta/M$ and
          \[
            \mu(y) = \exp\!\left(\int k(y)\,dy\right).
          \]
\end{itemize}
After multiplying the ODE through by $\mu$, solve the resulting exact equation
via the integration loop (Section~7.3).
\end{formulabox}

\begin{warningbox}[Common Pitfall: Always Re-Verify Exactness After Multiplying]
The formulas for $\mu(x)$ and $\mu(y)$ are derived by assuming the repaired
equation will be exact. After multiplying by $\mu$, explicitly re-compute the
cross partials $(\mu M)_y$ and $(\mu N)_x$ and confirm they are equal before
proceeding. A mis-simplified defect ratio can give a wrong $\mu$ that does
\emph{not} restore exactness.
\end{warningbox}

% ---------------------------------------------------------------------------
\subsection{Almost Exact Differential Equations}
% ---------------------------------------------------------------------------

An \emph{almost exact} equation is non-exact but possesses an integrating
factor of one of the two standard forms ($\mu(x)$ or $\mu(y)$). The key skill
is recognizing which test succeeds and executing the full repair-and-solve
workflow systematically.

\begin{methodbox}[Integrating-Factor Decision Tree]
Given $M\,dx + N\,dy = 0$ with $M_y \neq N_x$:
\begin{enumerate}[leftmargin=1.6em, itemsep=3pt]
    \item Compute the defect $\Delta = M_y - N_x$.
    \item \textbf{Try $\mu(x)$:} Form $\Delta/N$. Does all $y$ cancel?
          If yes, $\mu(x) = e^{\int \Delta/N\,dx}$.
    \item \textbf{Try $\mu(y)$:} Form $-\Delta/M$. Does all $x$ cancel?
          If yes, $\mu(y) = e^{\int(-\Delta/M)\,dy}$.
    \item Multiply the ODE by $\mu$, re-verify $(\mu M)_y = (\mu N)_x$, then
          use the integration loop to find $F$ and write $F = C$.
\end{enumerate}
\end{methodbox}

\begin{formulabox}[Why the Integrating Factor ODE Works]
Requiring $(\mu M)_y = (\mu N)_x$ for $\mu = \mu(x)$ yields
\[
  \mu\,M_y = \mu'\,N + \mu\,N_x
  \quad\Longrightarrow\quad
  \frac{\mu'}{\mu} = \frac{M_y - N_x}{N} = \frac{\Delta}{N},
\]
a separable ODE in $\mu(x)$ whose solution is $\mu = e^{\int \Delta/N\,dx}$.
The analogous calculation with $\mu = \mu(y)$ gives $\mu'/\mu = -\Delta/M$.
\end{formulabox}

% ---------------------------------------------------------------------------
\subsection{Exact Equations and Conservative Vector Fields}
% ---------------------------------------------------------------------------

Every exact equation $M\,dx + N\,dy = 0$ can be read as the zero-work
condition $\mathbf{F}\cdot d\mathbf{r}=0$ for the vector field
$\mathbf{F}=\langle M,N\rangle$. Exactness is equivalent to $\mathbf{F}$
being \emph{conservative}, and the potential $F$ of the ODE is the scalar
potential $\varphi$ with $\nabla\varphi=\mathbf{F}$.

\begin{formulabox}[Conservative Fields, Scalar Potentials, and Path Independence]
A $C^1$ vector field $\mathbf{F}=\langle M,N\rangle$ on a simply connected
region is \textbf{conservative} if and only if
\[
  \frac{\partial M}{\partial y} = \frac{\partial N}{\partial x}.
\]
When conservative, there exists a scalar potential $\varphi$ with
$\nabla\varphi = \mathbf{F}$, found by the same integration loop as the ODE
potential. Line integrals are then \textbf{path independent}:
\[
  \int_C \mathbf{F}\cdot d\mathbf{r}
  = \varphi\!\left(B\right) - \varphi\!\left(A\right)
\]
for any smooth curve $C$ from point $A$ to point $B$ (Fundamental Theorem for
Line Integrals). Equivalently, $\oint_C \mathbf{F}\cdot d\mathbf{r}=0$ for
every closed curve.
\end{formulabox}

\begin{warningbox}[Common Pitfall: The Domain Must Be Simply Connected]
The equivalence $M_y = N_x \iff$ conservative holds only on a
\textbf{simply connected} domain (a domain with no holes). On a
multiply-connected domain such as $\mathbb{R}^2\setminus\{(0,0)\}$, the field
$\mathbf{F}=\langle -y,\,x\rangle/(x^2+y^2)$ satisfies $M_y=N_x$ everywhere
it is defined yet is \emph{not} conservative: its integral around the unit
circle equals $2\pi\neq 0$.
\end{warningbox}

% ===========================================================================
% PART II: REVIEW GUIDE (PRACTICE QUESTIONS)
% Structural \clearpage before Part II in the master Reference Guide.
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part II --- Review Guide: Practice Problems}
\section*{Part II --- Review Guide: Practice Problems}

Work the following ten problems in order. Problems 1--4 develop the core
exactness test and potential-function reconstruction (Modules 7.1--7.3).
Problems 5--7 require finding a special integrating factor $\mu(x)$ or $\mu(y)$
to repair a non-exact equation (Modules 7.4--7.5). Problem 8 applies the
conservative-field interpretation and the Fundamental Theorem for Line
Integrals (Module 7.6). Problems 9--10 treat thermodynamic state functions,
the exact versus inexact distinction, and entropy via the $1/T$ integrating
factor (Module 7.7). Complete solutions follow in Part III.

\begin{enumerate}[leftmargin=2em, itemsep=8pt]

    \item For the equation $(2x + 3y)\,dx + (3x + 2y)\,dy = 0$, compute $M_y$
          and $N_x$, confirm the equation is exact, and reconstruct the
          potential function $F(x,y) = C$.

    \item Verify that $(2xy - \sec^2\!x)\,dx + (x^2 + 2y)\,dy = 0$ is exact,
          then determine $F(x,y)$ by integrating $M$ with respect to $x$ and
          solving for the integration term $g(y)$.

    \item Show that $(e^x\sin y + 2y)\,dx + (e^x\cos y + 2x - 2)\,dy = 0$
          is exact and find the implicit solution $F(x,y) = C$.

    \item Given $\bigl(\tfrac{y}{x} + 6x\bigr)\,dx + (\ln x - 2)\,dy = 0$
          for $x > 0$, test for exactness, reconstruct $F(x,y)$ including the
          full evaluation of $g(y)$, and then solve the initial-value problem
          with $y(1) = 4$.

    \item The equation $(3xy + y^2)\,dx + (x^2 + xy)\,dy = 0$ is not exact.
          Determine whether an integrating factor of the form $\mu(x)$ exists,
          find it, and solve the resulting exact equation.

    \item Show that $(2y^2 + 3x)\,dx + 2xy\,dy = 0$ fails the exactness test,
          find an integrating factor depending on $x$ alone, and produce the
          solution $F(x,y) = C$.

    \item For $y\,dx + (2x - ye^y)\,dy = 0$, demonstrate that no integrating
          factor $\mu(x)$ works but an integrating factor $\mu(y)$ does;
          determine $\mu(y)$ and solve.

    \item Consider the vector field
          $\mathbf{F}(x,y) = \langle 2xy + y^3,\; x^2 + 3xy^2 \rangle$.
          Show that $\mathbf{F}$ is conservative, find a scalar potential
          $\varphi$ with $\nabla\varphi = \mathbf{F}$, and use it to evaluate
          $\displaystyle\int_C \mathbf{F}\cdot d\mathbf{r}$ along any path
          from $(0,0)$ to $(2,1)$.

    \item For an ideal gas the inexact heat differential is
          $\delta Q = C_V\,dT + \tfrac{nRT}{V}\,dV$.
          Show that $\delta Q$ is not exact, then verify that multiplying by
          the integrating factor $1/T$ produces an exact differential
          $dS = \delta Q/T$, and reconstruct the state function $S(T,V)$.

    \item \textbf{(Capstone)} Given $dU = C_V\,dT$ (internal energy) and
          $\delta Q = dU + P\,dV$ with $P = nRT/V$ for an ideal gas, determine
          which of $dU$ and $\delta Q$ is an exact (state) differential by the
          $M_y = N_x$ criterion applied to each. Confirm that $1/T$ converts
          the inexact form into the exact entropy differential $dS$, and
          exhibit the recovered $S(T,V)$ up to an additive constant.

\end{enumerate}

% ===========================================================================
% PART III: COMPLETE SOLUTIONS
% Structural \clearpage before Part III in the master Reference Guide.
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part III --- Complete Solutions}
\section*{Part III --- Complete Solutions}

% -------------------------------------------------------------------
\noindent\textbf{Solution 1.}\quad
Identify $M = 2x + 3y$ and $N = 3x + 2y$. Compute the cross-partial
derivatives:
\[
  \textcolor{workpurple}{M_y = \frac{\partial}{\partial y}(2x + 3y) = 3,
  \qquad
  N_x = \frac{\partial}{\partial x}(3x + 2y) = 3.}
\]
Since $M_y = N_x = 3$, the equation is \textbf{exact}.
Integrate $M$ with respect to $x$, holding $y$ fixed:
\[
  \textcolor{workpurple}{F(x,y) = \int(2x + 3y)\,dx = x^2 + 3xy + g(y).}
\]
Differentiate $F$ with respect to $y$ and match to $N$:
\[
  \textcolor{workpurple}{F_y = 3x + g'(y) = N = 3x + 2y
  \quad\Longrightarrow\quad g'(y) = 2y.}
\]
Integrate $g'(y)$:
\[
  \textcolor{workpurple}{g(y) = \int 2y\,dy = y^2.}
\]
The general solution is
\[
  \textcolor{workpurple}{F(x,y) = x^2 + 3xy + y^2 = C.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 2.}\quad
Identify $M = 2xy - \sec^2\!x$ and $N = x^2 + 2y$. Compute the
cross-partial derivatives:
\[
  \textcolor{workpurple}{M_y = \frac{\partial}{\partial y}(2xy - \sec^2\!x) = 2x,
  \qquad
  N_x = \frac{\partial}{\partial x}(x^2 + 2y) = 2x.}
\]
Since $M_y = N_x = 2x$, the equation is \textbf{exact}.
Integrate $M$ with respect to $x$, using $\int\sec^2\!x\,dx = \tan x$:
\[
  \textcolor{workpurple}{F(x,y)
  = \int(2xy - \sec^2\!x)\,dx
  = x^2 y - \tan x + g(y).}
\]
Differentiate $F$ with respect to $y$ and match to $N$:
\[
  \textcolor{workpurple}{F_y = x^2 + g'(y) = x^2 + 2y
  \quad\Longrightarrow\quad g'(y) = 2y.}
\]
Integrate $g'(y)$:
\[
  \textcolor{workpurple}{g(y) = y^2.}
\]
The general solution is
\[
  \textcolor{workpurple}{F(x,y) = x^2 y - \tan x + y^2 = C.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 3.}\quad
Identify $M = e^x\sin y + 2y$ and $N = e^x\cos y + 2x - 2$.
Compute the cross-partial derivatives:
\[
  \textcolor{workpurple}{M_y
  = \frac{\partial}{\partial y}(e^x\sin y + 2y)
  = e^x\cos y + 2.}
\]
\[
  \textcolor{workpurple}{N_x
  = \frac{\partial}{\partial x}(e^x\cos y + 2x - 2)
  = e^x\cos y + 2.}
\]
Since $M_y = N_x = e^x\cos y + 2$, the equation is \textbf{exact}.
Integrate $M$ with respect to $x$:
\[
  \textcolor{workpurple}{F(x,y)
  = \int(e^x\sin y + 2y)\,dx
  = e^x\sin y + 2xy + g(y).}
\]
Differentiate $F$ with respect to $y$ and match to $N$:
\[
  \textcolor{workpurple}{F_y = e^x\cos y + 2x + g'(y)
  = e^x\cos y + 2x - 2
  \quad\Longrightarrow\quad g'(y) = -2.}
\]
Integrate $g'(y)$:
\[
  \textcolor{workpurple}{g(y) = -2y.}
\]
The general solution is
\[
  \textcolor{workpurple}{F(x,y) = e^x\sin y + 2xy - 2y = C.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 4.}\quad
Identify $M = y/x + 6x$ and $N = \ln x - 2$ (with $x>0$).
Compute the cross-partial derivatives:
\[
  \textcolor{workpurple}{M_y
  = \frac{\partial}{\partial y}\!\left(\frac{y}{x} + 6x\right)
  = \frac{1}{x},
  \qquad
  N_x
  = \frac{\partial}{\partial x}(\ln x - 2)
  = \frac{1}{x}.}
\]
Since $M_y = N_x = 1/x$, the equation is \textbf{exact}.
Integrate $M$ with respect to $x$:
\[
  \textcolor{workpurple}{F(x,y)
  = \int\!\left(\frac{y}{x} + 6x\right)dx
  = y\ln x + 3x^2 + g(y).}
\]
Differentiate $F$ with respect to $y$ and match to $N$:
\[
  \textcolor{workpurple}{F_y = \ln x + g'(y) = \ln x - 2
  \quad\Longrightarrow\quad g'(y) = -2.}
\]
Integrate $g'(y)$:
\[
  \textcolor{workpurple}{g(y) = -2y.}
\]
The general solution is $y\ln x + 3x^2 - 2y = C$.
Apply the initial condition $y(1) = 4$; since $\ln 1 = 0$:
\[
  \textcolor{workpurple}{(4)\ln 1 + 3(1)^2 - 2(4)
  = 0 + 3 - 8 = -5 = C.}
\]
The particular implicit solution is $y\ln x + 3x^2 - 2y = -5$.
Collecting $y$ terms and solving explicitly:
\[
  \textcolor{workpurple}{y(\ln x - 2) = -5 - 3x^2
  \quad\Longrightarrow\quad
  y = \frac{-5 - 3x^2}{\ln x - 2} = \frac{3x^2 + 5}{2 - \ln x}.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 5.}\quad
Identify $M = 3xy + y^2$ and $N = x^2 + xy$.
Compute the cross-partial derivatives:
\[
  \textcolor{workpurple}{M_y
  = \frac{\partial}{\partial y}(3xy + y^2)
  = 3x + 2y,
  \qquad
  N_x
  = \frac{\partial}{\partial x}(x^2 + xy)
  = 2x + y.}
\]
Since $M_y = 3x + 2y \neq 2x + y = N_x$, the equation is \textbf{not exact}.
Compute the exactness defect:
\[
  \textcolor{workpurple}{\Delta = M_y - N_x = (3x + 2y) - (2x + y) = x + y.}
\]
Test for $\mu(x)$ by forming $\Delta/N$:
\[
  \textcolor{workpurple}{\frac{\Delta}{N}
  = \frac{x + y}{x^2 + xy}
  = \frac{x + y}{x(x + y)}
  = \frac{1}{x}.}
\]
This is a function of $x$ alone, so $\mu(x)$ exists:
\[
  \textcolor{workpurple}{\mu(x)
  = e^{\int (1/x)\,dx}
  = e^{\ln x}
  = x.}
\]
Multiply the ODE by $\mu = x$; the new coefficients are
$M^* = x(3xy + y^2) = 3x^2 y + xy^2$ and
$N^* = x(x^2 + xy) = x^3 + x^2 y$.
Re-verify exactness:
\[
  \textcolor{workpurple}{(M^*)_y
  = \frac{\partial}{\partial y}(3x^2 y + xy^2)
  = 3x^2 + 2xy,
  \qquad
  (N^*)_x
  = \frac{\partial}{\partial x}(x^3 + x^2 y)
  = 3x^2 + 2xy.}
\]
Equal. Now exact. Integrate $M^*$ with respect to $x$:
\[
  \textcolor{workpurple}{F(x,y)
  = \int(3x^2 y + xy^2)\,dx
  = x^3 y + \tfrac{1}{2}x^2 y^2 + g(y).}
\]
Differentiate $F$ with respect to $y$ and match to $N^*$:
\[
  \textcolor{workpurple}{F_y = x^3 + x^2 y + g'(y) = x^3 + x^2 y
  \quad\Longrightarrow\quad g'(y) = 0,\quad g(y) = 0.}
\]
The general solution is
\[
  \textcolor{workpurple}{F(x,y) = x^3 y + \tfrac{1}{2}x^2 y^2 = C.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 6.}\quad
Identify $M = 2y^2 + 3x$ and $N = 2xy$.
Compute the cross-partial derivatives:
\[
  \textcolor{workpurple}{M_y
  = \frac{\partial}{\partial y}(2y^2 + 3x)
  = 4y,
  \qquad
  N_x
  = \frac{\partial}{\partial x}(2xy)
  = 2y.}
\]
Since $M_y = 4y \neq 2y = N_x$, the equation is \textbf{not exact}.
Compute the exactness defect:
\[
  \textcolor{workpurple}{\Delta = M_y - N_x = 4y - 2y = 2y.}
\]
Test for $\mu(x)$ by forming $\Delta/N$:
\[
  \textcolor{workpurple}{\frac{\Delta}{N}
  = \frac{2y}{2xy}
  = \frac{1}{x}.}
\]
Pure in $x$. The integrating factor is:
\[
  \textcolor{workpurple}{\mu(x)
  = e^{\int (1/x)\,dx}
  = e^{\ln x}
  = x.}
\]
Multiply the ODE by $\mu = x$; the new coefficients are
$M^* = x(2y^2 + 3x) = 2xy^2 + 3x^2$ and $N^* = x\cdot 2xy = 2x^2 y$.
Re-verify exactness:
\[
  \textcolor{workpurple}{(M^*)_y
  = \frac{\partial}{\partial y}(2xy^2 + 3x^2)
  = 4xy,
  \qquad
  (N^*)_x
  = \frac{\partial}{\partial x}(2x^2 y)
  = 4xy.}
\]
Equal. Now exact. Integrate $M^*$ with respect to $x$:
\[
  \textcolor{workpurple}{F(x,y)
  = \int(2xy^2 + 3x^2)\,dx
  = x^2 y^2 + x^3 + g(y).}
\]
Differentiate $F$ with respect to $y$ and match to $N^*$:
\[
  \textcolor{workpurple}{F_y = 2x^2 y + g'(y) = 2x^2 y
  \quad\Longrightarrow\quad g'(y) = 0,\quad g(y) = 0.}
\]
The general solution is
\[
  \textcolor{workpurple}{F(x,y) = x^2 y^2 + x^3 = C.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 7.}\quad
Identify $M = y$ and $N = 2x - ye^y$.
Compute the cross-partial derivatives:
\[
  \textcolor{workpurple}{M_y
  = \frac{\partial}{\partial y}(y)
  = 1,
  \qquad
  N_x
  = \frac{\partial}{\partial x}(2x - ye^y)
  = 2.}
\]
Since $M_y = 1 \neq 2 = N_x$, the equation is \textbf{not exact}.
Compute $\Delta = M_y - N_x = 1 - 2 = -1$.

\noindent\textit{Test $\mu(x)$:} Form $\Delta/N$:
\[
  \textcolor{workpurple}{\frac{\Delta}{N}
  = \frac{-1}{2x - ye^y}.}
\]
This contains $y$, so it is not a function of $x$ alone. \textbf{No $\mu(x)$
exists.}

\noindent\textit{Test $\mu(y)$:} Form $-\Delta/M = (N_x - M_y)/M$:
\[
  \textcolor{workpurple}{\frac{N_x - M_y}{M}
  = \frac{2 - 1}{y}
  = \frac{1}{y}.}
\]
Pure in $y$. The integrating factor is:
\[
  \textcolor{workpurple}{\mu(y)
  = e^{\int (1/y)\,dy}
  = e^{\ln y}
  = y.}
\]
Multiply the ODE by $\mu = y$; the new coefficients are
$M^* = y\cdot y = y^2$ and $N^* = y(2x - ye^y) = 2xy - y^2 e^y$.
Re-verify exactness:
\[
  \textcolor{workpurple}{(M^*)_y
  = \frac{\partial}{\partial y}(y^2)
  = 2y,
  \qquad
  (N^*)_x
  = \frac{\partial}{\partial x}(2xy - y^2 e^y)
  = 2y.}
\]
Equal. Now exact. Integrate $M^*$ with respect to $x$:
\[
  \textcolor{workpurple}{F(x,y)
  = \int y^2\,dx
  = xy^2 + g(y).}
\]
Differentiate $F$ with respect to $y$ and match to $N^*$:
\[
  \textcolor{workpurple}{F_y = 2xy + g'(y) = 2xy - y^2 e^y
  \quad\Longrightarrow\quad g'(y) = -y^2 e^y.}
\]
Evaluate $g(y) = -\int y^2 e^y\,dy$ by integration by parts twice.
First application, $u = y^2$ and $dv = e^y\,dy$:
\[
  \textcolor{workpurple}{\int y^2 e^y\,dy
  = y^2 e^y - 2\int y e^y\,dy.}
\]
Second application for $\int y e^y\,dy$, with $u = y$ and $dv = e^y\,dy$:
\[
  \textcolor{workpurple}{\int y e^y\,dy
  = ye^y - \int e^y\,dy
  = ye^y - e^y
  = e^y(y - 1).}
\]
Combining:
\[
  \textcolor{workpurple}{\int y^2 e^y\,dy
  = y^2 e^y - 2e^y(y-1)
  = e^y(y^2 - 2y + 2).}
\]
Therefore:
\[
  \textcolor{workpurple}{g(y) = -e^y(y^2 - 2y + 2).}
\]
The general solution is
\[
  \textcolor{workpurple}{F(x,y) = xy^2 - e^y(y^2 - 2y + 2) = C.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 8.}\quad
Let $P = 2xy + y^3$ and $Q = x^2 + 3xy^2$.
Compute the cross-partial derivatives:
\[
  \textcolor{workpurple}{P_y
  = \frac{\partial}{\partial y}(2xy + y^3)
  = 2x + 3y^2,
  \qquad
  Q_x
  = \frac{\partial}{\partial x}(x^2 + 3xy^2)
  = 2x + 3y^2.}
\]
Since $P_y = Q_x = 2x + 3y^2$ on all of $\mathbb{R}^2$ (a simply connected
domain), $\mathbf{F}$ is \textbf{conservative}.
Find the scalar potential $\varphi$ by integrating $P$ with respect to $x$:
\[
  \textcolor{workpurple}{\varphi(x,y)
  = \int(2xy + y^3)\,dx
  = x^2 y + xy^3 + g(y).}
\]
Differentiate $\varphi$ with respect to $y$ and match to $Q$:
\[
  \textcolor{workpurple}{\varphi_y
  = x^2 + 3xy^2 + g'(y)
  = x^2 + 3xy^2
  \quad\Longrightarrow\quad g'(y) = 0,\quad g(y) = 0.}
\]
Therefore:
\[
  \textcolor{workpurple}{\varphi(x,y) = x^2 y + xy^3.}
\]
By the Fundamental Theorem for Line Integrals, the integral is path
independent:
\[
  \textcolor{workpurple}{\int_C \mathbf{F}\cdot d\mathbf{r}
  = \varphi(2,1) - \varphi(0,0).}
\]
Evaluate each endpoint:
\[
  \textcolor{workpurple}{\varphi(2,1)
  = (2)^2(1) + (2)(1)^3
  = 4 + 2
  = 6,
  \qquad
  \varphi(0,0) = 0.}
\]
Therefore:
\[
  \textcolor{workpurple}{\int_C \mathbf{F}\cdot d\mathbf{r}
  = 6 - 0 = 6.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 9.}\quad
Treat $T$ and $V$ as the two independent variables. The differential is
$\delta Q = C_V\,dT + \tfrac{nRT}{V}\,dV$, so
$M = C_V$ (coefficient of $dT$) and $N = nRT/V$ (coefficient of $dV$).

Compute the cross-partial derivatives:
\[
  \textcolor{workpurple}{\frac{\partial M}{\partial V}
  = \frac{\partial C_V}{\partial V}
  = 0,
  \qquad
  \frac{\partial N}{\partial T}
  = \frac{\partial}{\partial T}\!\left(\frac{nRT}{V}\right)
  = \frac{nR}{V}.}
\]
Since $0 \neq nR/V$, the cross partials are unequal: $\delta Q$ is
\textbf{not exact} and heat is not a state function.

Multiply by the integrating factor $1/T$ to define $dS = \delta Q/T$:
\[
  \textcolor{workpurple}{dS
  = \frac{\delta Q}{T}
  = \frac{C_V}{T}\,dT + \frac{nR}{V}\,dV.}
\]
Now $M^* = C_V/T$ and $N^* = nR/V$. Re-verify exactness:
\[
  \textcolor{workpurple}{\frac{\partial M^*}{\partial V}
  = \frac{\partial}{\partial V}\!\left(\frac{C_V}{T}\right)
  = 0,
  \qquad
  \frac{\partial N^*}{\partial T}
  = \frac{\partial}{\partial T}\!\left(\frac{nR}{V}\right)
  = 0.}
\]
Both cross partials equal zero; $dS$ is \textbf{exact}, confirming $1/T$ is a
valid integrating factor and $S$ is a state function.

Reconstruct $S$ by integrating $M^* = C_V/T$ with respect to $T$
($C_V$ constant):
\[
  \textcolor{workpurple}{S(T,V)
  = \int \frac{C_V}{T}\,dT + g(V)
  = C_V\ln T + g(V).}
\]
Differentiate $S$ with respect to $V$ and match to $N^*$:
\[
  \textcolor{workpurple}{\frac{\partial S}{\partial V}
  = g'(V)
  = \frac{nR}{V}
  \quad\Longrightarrow\quad
  g(V) = nR\ln V.}
\]
The entropy state function is:
\[
  \textcolor{workpurple}{S(T,V) = C_V\ln T + nR\ln V + \text{const.}}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 10 (Capstone).}\quad
The internal energy of an ideal gas depends only on temperature (Joule's
theorem), so the genuine state-function differential is
$dU = C_V\,dT + 0\cdot dV$. With $M = C_V$ (coefficient of $dT$) and
$N = 0$ (coefficient of $dV$), test exactness:
\[
  \textcolor{workpurple}{\frac{\partial M}{\partial V}
  = \frac{\partial C_V}{\partial V}
  = 0,
  \qquad
  \frac{\partial N}{\partial T}
  = \frac{\partial (0)}{\partial T}
  = 0.}
\]
The cross partials are equal: $dU = C_V\,dT$ is \textbf{exact}. Integrate to
confirm the potential:
\[
  \textcolor{workpurple}{U(T) = \int C_V\,dT = C_V T + \text{const.}}
\]
$U$ is a genuine state function depending only on $T$.

Now build $\delta Q$ using $dU = C_V\,dT$ and $P\,dV = (nRT/V)\,dV$:
\[
  \textcolor{workpurple}{\delta Q = dU + P\,dV
  = C_V\,dT + \frac{nRT}{V}\,dV.}
\]
With $M = C_V$ and $N = nRT/V$, test exactness:
\[
  \textcolor{workpurple}{\frac{\partial M}{\partial V}
  = 0,
  \qquad
  \frac{\partial N}{\partial T}
  = \frac{nR}{V}.}
\]
Since $0 \neq nR/V$, $\delta Q$ is \textbf{not exact}: heat is not a state
function. Of the two, $dU$ is the exact (state) differential and $\delta Q$
is the inexact one.

Apply the integrating factor $1/T$ to $\delta Q$:
\[
  \textcolor{workpurple}{dS
  = \frac{\delta Q}{T}
  = \frac{C_V}{T}\,dT + \frac{nR}{V}\,dV.}
\]
Test exactness of $dS$ with $M^* = C_V/T$ and $N^* = nR/V$:
\[
  \textcolor{workpurple}{\frac{\partial M^*}{\partial V}
  = 0,
  \qquad
  \frac{\partial N^*}{\partial T}
  = 0.}
\]
Equal. $dS$ is \textbf{exact}. Integrate $M^*$ with respect to $T$:
\[
  \textcolor{workpurple}{S(T,V)
  = \int\frac{C_V}{T}\,dT + g(V)
  = C_V\ln T + g(V).}
\]
Match $\partial S/\partial V$ to $N^*$:
\[
  \textcolor{workpurple}{g'(V)
  = \frac{nR}{V}
  \quad\Longrightarrow\quad
  g(V) = nR\ln V.}
\]
The entropy state function recovered from $\delta Q/T$ is:
\[
  \textcolor{workpurple}{S(T,V)
  = C_V\ln T + nR\ln V + \text{const.}}
\]

\vspace{0.8em}

\begin{conceptbox}[Unit 7 Key Takeaway]
Every exact equation $M\,dx+N\,dy=0$ reduces to finding the potential
\textcolor{accentorange}{$F(x,y)=C$} via the integration loop
\textcolor{accentorange}{$F=\int M\,dx+g(y)$} with $g'(y)$ pinned by matching
$F_y=N$. For non-exact equations, compute the defect
\textcolor{accentorange}{$\Delta=M_y-N_x$} and apply
\textcolor{accentorange}{$\mu(x)=e^{\int\Delta/N\,dx}$} when $\Delta/N$ is
pure in $x$, or \textcolor{accentorange}{$\mu(y)=e^{-\int\Delta/M\,dy}$} when
$-\Delta/M$ is pure in $y$; always re-verify exactness after multiplying. The
conservative-field identity \textcolor{accentorange}{$M_y=N_x\iff\mathbf{F}=\nabla\varphi$}
makes line integrals path-independent:
\textcolor{accentorange}{$\int_C\mathbf{F}\cdot d\mathbf{r}=\varphi(B)-\varphi(A)$}.
In thermodynamics, \textcolor{accentorange}{$dU$} is exact while
\textcolor{accentorange}{$\delta Q$} is not; the integrating factor
\textcolor{accentorange}{$1/T$} converts $\delta Q$ into the exact entropy
differential \textcolor{accentorange}{$dS=C_V\ln T+nR\ln V+\text{const}$}.
\end{conceptbox}

\end{document}
