% Unit 5: Numerical Methods
% Master Unit Reference Guide (Cheat Sheet, Problems, and Solutions)
% Prepared by Staples Education
% Content informed by the local curriculum modules (5.1 to 5.6) and the
% QUIZ_DATA micro-practice items for Unit 5.
% Shared style harness pulled from the unified preamble input.

\input{ode/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 5: Numerical Methods}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Master Unit Reference Guide: Cheat Sheet, Problems, and Solutions}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Table of contents (master guide)
% ---------------------------------------------------------------------------
\tableofcontents

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 5 at a Glance: Core Sub-Topics]
When an initial value problem $\frac{dy}{dt}=f(t,y),\ y(t_0)=y_0$ cannot be
solved in closed form, we march the solution forward in small steps of size $h$.
Every method in this unit is a recipe for turning the slope field $f$ into the
next point; they differ only in how many slopes they sample per step and how they
weight them.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Euler's Method} --- one tangent-line step
          $y_{n+1}=y_n+hf(t_n,y_n)$; first order, global error $O(h)$.
    \item \textbf{Improved and Modified Euler} --- two-slope averaging and the
          midpoint step; both second order, $O(h^{2})$.
    \item \textbf{Heun's Method} --- the predictor--corrector reading of improved
          Euler as a trapezoidal average of endpoint slopes.
    \item \textbf{The RK4 Method} --- four weighted stage slopes
          $\tfrac{h}{6}(k_1+2k_2+2k_3+k_4)$; fourth order, $O(h^{4})$.
    \item \textbf{The Runge--Kutta Family} --- the general weighted-stage template
          that contains Euler, Heun, and RK4 as members.
    \item \textbf{Adaptive Step Size and Comparison} --- estimating local error to
          resize $h$, and ranking the methods by order of accuracy.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{5}   % subsections auto-number as the unit modules 5.1, 5.2, ...

\subsection{Euler's Method}

Euler's method is the simplest numerical solver: from the current point it walks
a single step along the tangent line whose slope is $f(t_n,y_n)$. It is the
foundation every richer method refines.

\begin{formulabox}[The Euler Update]
\[ y_{n+1} = y_n + h\,f(t_n,y_n), \qquad t_{n+1} = t_n + h. \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item Each step follows the \textbf{tangent line} with slope $f(t_n,y_n)$.
    \item \textbf{Local} truncation error is $O(h^{2})$ (the dropped
          $\tfrac{h^{2}}{2}y''$ term); \textbf{global} error is $O(h)$, so Euler
          is a \textbf{first-order} method.
    \item Halving $h$ roughly \textbf{halves} the global error.
\end{itemize}
\end{formulabox}

\begin{examplebox}[Worked Example 5.1: One Euler step for $y'=y$]
With $y(0)=1$ and $h=0.1$, the slope at the start is $f(0,1)=1$:
\[ \textcolor{workpurple}{y_1 = 1 + 0.1\cdot 1 = 1.1, \qquad t_1 = 0.1.} \]
The exact value is $e^{0.1}\approx 1.10517$, so a single first-order step is
already low by about $0.005$.
\end{examplebox}

\subsection{Improved and Modified Euler Methods}

Both upgrades sample a \emph{second} slope to capture the curvature that plain
Euler ignores, lifting the order from one to two. The improved method averages
the slopes at the two endpoints; the modified (midpoint) method reads one slope
at the interval's center.

\begin{formulabox}[Improved Euler (Endpoint Average) and Modified Euler (Midpoint)]
\textbf{Improved Euler:}
\[ k_1 = f(t_n,y_n), \qquad k_2 = f\!\left(t_n+h,\; y_n+h k_1\right), \]
\[ y_{n+1} = y_n + \frac{h}{2}\left(k_1 + k_2\right). \]
\textbf{Modified Euler (midpoint):}
\[ k_1 = f(t_n,y_n), \qquad k_2 = f\!\left(t_n+\tfrac{h}{2},\; y_n+\tfrac{h}{2}k_1\right), \]
\[ y_{n+1} = y_n + h\,k_2. \]
Both are \textbf{second order}, with global error $O(h^{2})$: halving $h$ cuts the
error by about a factor of four.
\end{formulabox}

\begin{examplebox}[Worked Example 5.2: Improved Euler for $y'=y$]
With $y(0)=1$, $h=0.1$: the predictor slope is $k_1=1$ and the corrected slope is
$k_2=f(0.1,\,1.1)=1.1$. Averaging,
\[ \textcolor{workpurple}{y_1 = 1 + \frac{0.1}{2}\left(1 + 1.1\right) = 1 + 0.05\cdot 2.1 = 1.105.} \]
This already matches $e^{0.1}\approx 1.10517$ to three decimals --- far better than
the plain Euler $1.1$.
\end{examplebox}

\subsection{Heun's Method}

Heun's method is improved Euler told as a \emph{predictor--corrector}: take a
provisional Euler step to predict the endpoint, then correct with the
\emph{average} of the slopes at the two ends. That average is exactly the
trapezoidal rule applied to $\int f\,dt$.

\begin{methodbox}[Heun's Predictor--Corrector]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item \textbf{Predict} the endpoint with a plain Euler step:
          \[ \tilde{y}_{n+1} = y_n + h\,f(t_n,y_n). \]
    \item \textbf{Correct} with the trapezoidal average of the two slopes:
          \[ y_{n+1} = y_n + \frac{h}{2}\Big[\,f(t_n,y_n) + f(t_{n+1},\tilde{y}_{n+1})\,\Big]. \]
\end{enumerate}
Averaging the two endpoint slopes recovers the curvature term plain Euler drops,
making Heun a \textbf{second-order} method --- identical to improved Euler.
\end{methodbox}

\begin{warningbox}[Common Pitfall: Unequal vs.\ Equal Slope Weights]
The improved Euler / Heun corrector weights the two slopes \emph{equally},
$\tfrac{h}{2}(k_1+k_2)$. Do not borrow RK4's $\tfrac{h}{6}(k_1+2k_2+2k_3+k_4)$
weights for a two-slope method, and do not forget the predictor uses the
\emph{full} step $h$ while the midpoint method uses a \emph{half} step
$\tfrac{h}{2}$.
\end{warningbox}

\subsection{The RK4 Method}

The classical fourth-order Runge--Kutta method samples four slopes per step --- at
the start, twice at the midpoint, and once at the end --- and blends them with
Simpson-like weights $1,2,2,1$. It is the workhorse of practical ODE solving.

\begin{formulabox}[The Classical RK4 Step]
\[ k_1 = f(t_n,\,y_n), \]
\[ k_2 = f\!\left(t_n+\tfrac{h}{2},\; y_n+\tfrac{h}{2}k_1\right), \qquad
   k_3 = f\!\left(t_n+\tfrac{h}{2},\; y_n+\tfrac{h}{2}k_2\right), \]
\[ k_4 = f\!\left(t_n+h,\; y_n+h k_3\right), \]
\[ y_{n+1} = y_n + \frac{h}{6}\left(k_1 + 2k_2 + 2k_3 + k_4\right). \]
RK4 is \textbf{fourth order}, with global error $O(h^{4})$: halving $h$ shrinks the
error by about a factor of $\mathbf{16}$.
\end{formulabox}

\begin{examplebox}[Worked Example 5.4: One RK4 step for $y'=y$]
With $y(0)=1$, $h=0.1$:
\[ \textcolor{workpurple}{k_1 = 1,\quad k_2 = 1.05,\quad k_3 = 1.0525,\quad k_4 = 1.10525.} \]
\[ \textcolor{workpurple}{y_1 = 1 + \frac{0.1}{6}\big(1 + 2(1.05) + 2(1.0525) + 1.10525\big) = 1.105171.} \]
Against $e^{0.1}=1.1051709\ldots$ the error is below $10^{-6}$ in a single step.
\end{examplebox}

\subsection{The Runge--Kutta Family}

Euler, the midpoint method, Heun, and RK4 are all instances of one template: take
a weighted average of slopes sampled at chosen points across the step. The
\emph{number of stages} and the \emph{weights} fix the order of accuracy.

\begin{formulabox}[The General Explicit Runge--Kutta Step]
For $s$ stages with weights $b_i$, nodes $c_i$, and coupling coefficients
$a_{ij}$:
\[ k_i = f\!\left(t_n + c_i h,\; y_n + h\sum_{j<i} a_{ij}\,k_j\right), \qquad
   y_{n+1} = y_n + h\sum_{i=1}^{s} b_i\,k_i. \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item Consistency requires $\sum_i b_i = 1$ (the weights sum to one).
    \item Euler is the $1$-stage member; improved Euler, the midpoint method, and
          Heun are $2$-stage second-order members; RK4 is the classic $4$-stage
          fourth-order member.
\end{itemize}
\end{formulabox}

\subsection{Adaptive Step Size and Method Comparison}

Rather than fix $h$ in advance, adaptive solvers \emph{estimate} the local error
each step and resize $h$ to hold it under a tolerance --- small steps through
sharp turns, large steps across calm stretches.

\begin{methodbox}[Estimating Local Error to Resize $h$]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item Advance the step two ways of \emph{different order} (e.g.\ an embedded
          RK4/RK5 pair, or one full step vs.\ two half steps).
    \item The difference of the two estimates approximates the \textbf{local
          error}.
    \item If the error exceeds the tolerance, \textbf{shrink} $h$ and retry; if it
          is comfortably under, \textbf{grow} $h$ for the next step.
\end{enumerate}
\end{methodbox}

\begin{warningbox}[Common Pitfall: Confusing Order with Step Count]
A method's \emph{order} $p$ is how fast error falls as $h\to 0$ (global error
$O(h^{p})$), not how many steps it takes. RK4 with a large $h$ can beat Euler with
a tiny $h$ at equal cost. And smaller $h$ is not free: too small, and accumulated
roundoff begins to swamp the truncation gain.
\end{warningbox}

\begin{conceptbox}[Unit 5 Key Takeaway]
Every solver here marches $y_{n+1}=y_n+h\,(\text{weighted slope})$; they differ
only in the slope. \textcolor{accentorange}{Euler} samples one slope (first order,
$O(h)$); \textcolor{accentorange}{improved/modified Euler and Heun} average two
(second order, $O(h^{2})$); \textcolor{accentorange}{RK4} blends four with weights
$\textcolor{accentorange}{1,2,2,1}$ (fourth order, $O(h^{4})$). All sit inside the
\textcolor{accentorange}{Runge--Kutta family}, and \textcolor{accentorange}{adaptive
stepping} tunes $h$ from an on-the-fly local-error estimate.
\end{conceptbox}

% ===========================================================================
% PART II: REVIEW GUIDE (PRACTICE QUESTIONS)
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part II --- Review Guide: Practice Problems}
\section*{Part II --- Review Guide: Practice Problems}

Work the following ten problems in order. They climb from a single Euler step,
through the two-slope second-order methods (improved Euler, modified Euler, and
Heun), into the four-stage RK4 step, and close with a method comparison and an
adaptive-stepping capstone. Throughout, take $\frac{dy}{dt}=f(t,y)$ with step
size $h$. Solutions follow in Part III.

\begin{enumerate}[leftmargin=2em]
    \item Write the Euler update formula, then take one Euler step for
          $\dfrac{dy}{dt} = t + y$ with $y(0) = 1$ and $h = 0.2$. Report $y_1$
          and $t_1$.

    \item Continue from Problem 1 and take a \emph{second} Euler step to find
          $y_2$ at $t_2 = 0.4$.

    \item State the local and global truncation error orders of Euler's method,
          and the factor by which the global error changes when $h$ is halved.

    \item Take one \emph{improved Euler} step for $\dfrac{dy}{dt} = t + y$ with
          $y(0) = 1$ and $h = 0.2$. Show $k_1$, $k_2$, and $y_1$.

    \item Take one \emph{modified Euler} (midpoint) step for the same problem,
          $\dfrac{dy}{dt} = t + y$, $y(0)=1$, $h = 0.2$. Show $k_1$, $k_2$, and
          $y_1$.

    \item Write Heun's predictor--corrector formulas and explain in one line why
          the method is second order.

    \item For a general $f(t,y)$, write the four RK4 stage slopes
          $k_1,k_2,k_3,k_4$ and the final RK4 update for $y_{n+1}$.

    \item Take one RK4 step for $\dfrac{dy}{dt} = y$ with $y(0) = 1$ and
          $h = 0.1$. Compute $y_1$ and compare it to $e^{0.1}$.

    \item Rank Euler, improved/modified Euler, Heun, and RK4 by order of
          accuracy, and give the factor by which each method's global error
          drops when $h$ is halved.

    \item Explain how an adaptive solver uses two estimates of \emph{different}
          order to decide whether to shrink or grow the step size $h$.

    \item \textbf{(Capstone)} Take one RK4 step for $\dfrac{dy}{dt} = t + y$ with
          $y(0) = 1$ and $h = 0.2$. Compute $y_1$, then compare it against the
          Euler result (Problem 1) and the improved Euler result (Problem 4)
          using the exact solution $y = 2e^{t} - t - 1$.
\end{enumerate}

% ===========================================================================
% PART III: SOLUTIONS
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part III --- Complete Solutions}
\section*{Part III --- Complete Solutions}

\noindent\textbf{Solution 1.}\quad The Euler update is
$y_{n+1} = y_n + h\,f(t_n,y_n)$. With $f(t,y)=t+y$, the starting slope is
$f(0,1) = 0 + 1 = 1$:
\[ \textcolor{workpurple}{y_1 = 1 + 0.2\cdot 1 = 1.2, \qquad t_1 = 0.2.} \]

\medskip
\noindent\textbf{Solution 2.}\quad From $(t_1,y_1) = (0.2,\,1.2)$ the slope is
$f(0.2,1.2) = 0.2 + 1.2 = 1.4$:
\[ \textcolor{workpurple}{y_2 = 1.2 + 0.2\cdot 1.4 = 1.48, \qquad t_2 = 0.4.} \]

\medskip
\noindent\textbf{Solution 3.}\quad A single Euler step drops the
$\tfrac{h^{2}}{2}y''$ term, so the \textbf{local} truncation error is $O(h^{2})$.
Accumulated over $\sim 1/h$ steps, the \textbf{global} error is $O(h)$, making
Euler a first-order method. Halving $h$ therefore
\[ \textcolor{workpurple}{\text{cuts the global error by about a factor of } 2.} \]

\medskip
\noindent\textbf{Solution 4.}\quad Improved Euler uses
$k_1 = f(t_n,y_n)$ and $k_2 = f(t_n+h,\,y_n+hk_1)$, then averages:
\[ \textcolor{workpurple}{k_1 = f(0,1) = 1, \qquad k_2 = f(0.2,\;1 + 0.2\cdot 1) = f(0.2,\,1.2) = 1.4.} \]
\[ \textcolor{workpurple}{y_1 = 1 + \frac{0.2}{2}\left(1 + 1.4\right) = 1 + 0.1\cdot 2.4 = 1.24.} \]

\medskip
\noindent\textbf{Solution 5.}\quad Modified Euler reads the slope at the
midpoint after a half-step prediction:
\[ \textcolor{workpurple}{k_1 = f(0,1) = 1, \qquad k_2 = f\!\left(0.1,\; 1 + 0.1\cdot 1\right) = f(0.1,\,1.1) = 1.2.} \]
\[ \textcolor{workpurple}{y_1 = 1 + h\,k_2 = 1 + 0.2\cdot 1.2 = 1.24.} \]
Here both second-order methods agree to the digits shown.

\medskip
\noindent\textbf{Solution 6.}\quad Heun predicts with a full Euler step and
corrects with the trapezoidal average of the endpoint slopes:
\[ \textcolor{workpurple}{\tilde{y}_{n+1} = y_n + h f(t_n,y_n), \qquad
   y_{n+1} = y_n + \frac{h}{2}\big[f(t_n,y_n) + f(t_{n+1},\tilde{y}_{n+1})\big].} \]
Averaging the two endpoint slopes restores the leading curvature term that plain
Euler discards, so the global error is $O(h^{2})$ --- second order.

\medskip
\noindent\textbf{Solution 7.}\quad The classical RK4 stages and update are
\[ \textcolor{workpurple}{k_1 = f(t_n,y_n),} \]
\[ \textcolor{workpurple}{k_2 = f\!\left(t_n+\tfrac{h}{2},\, y_n+\tfrac{h}{2}k_1\right),
   \qquad k_3 = f\!\left(t_n+\tfrac{h}{2},\, y_n+\tfrac{h}{2}k_2\right),} \]
\[ \textcolor{workpurple}{k_4 = f\!\left(t_n+h,\, y_n+h k_3\right),} \]
\[ \textcolor{workpurple}{y_{n+1} = y_n + \frac{h}{6}\left(k_1 + 2k_2 + 2k_3 + k_4\right).} \]

\medskip
\noindent\textbf{Solution 8.}\quad For $f(t,y)=y$ with $y(0)=1$, $h=0.1$:
\[ \textcolor{workpurple}{k_1 = 1, \quad k_2 = 1 + 0.05\cdot 1 = 1.05, \quad
   k_3 = 1 + 0.05\cdot 1.05 = 1.0525,} \]
\[ \textcolor{workpurple}{k_4 = 1 + 0.1\cdot 1.0525 = 1.10525.} \]
\[ \textcolor{workpurple}{y_1 = 1 + \frac{0.1}{6}\big(1 + 2(1.05) + 2(1.0525) + 1.10525\big) = 1 + \frac{0.1}{6}(6.31025) = 1.105171.} \]
Against $e^{0.1} = 1.1051709\ldots$ the error is below $10^{-6}$.

\medskip
\noindent\textbf{Solution 9.}\quad Ranked by order of accuracy:
\[
\textcolor{workpurple}{
\begin{array}{lll}
\text{Euler} & \text{1st order } O(h) & \text{halving } h \Rightarrow \div 2\\
\text{Improved / Modified Euler} & \text{2nd order } O(h^{2}) & \text{halving } h \Rightarrow \div 4\\
\text{Heun} & \text{2nd order } O(h^{2}) & \text{halving } h \Rightarrow \div 4\\
\text{RK4} & \text{4th order } O(h^{4}) & \text{halving } h \Rightarrow \div 16
\end{array}}
\]

\medskip
\noindent\textbf{Solution 10.}\quad An adaptive solver advances each step two
ways of \emph{different} order --- for example an embedded RK4/RK5 pair, or one
full step compared against two half steps. The difference between the two
estimates approximates the local error. If that estimate \textbf{exceeds} the
tolerance, the solver \textbf{shrinks} $h$ and retries the step; if it sits
comfortably \textbf{under} the tolerance, the solver \textbf{grows} $h$ for the
next step, so the work concentrates where the solution changes fastest.

\medskip
\noindent\textbf{Solution (Capstone).}\quad RK4 for $f(t,y)=t+y$, $y(0)=1$,
$h=0.2$:
\[ \textcolor{workpurple}{k_1 = f(0,1) = 1, \qquad k_2 = f(0.1,\;1+0.1\cdot 1) = f(0.1,\,1.1) = 1.2,} \]
\[ \textcolor{workpurple}{k_3 = f(0.1,\;1+0.1\cdot 1.2) = f(0.1,\,1.12) = 1.22, \qquad k_4 = f(0.2,\;1+0.2\cdot 1.22) = f(0.2,\,1.244) = 1.444.} \]
\[ \textcolor{workpurple}{y_1 = 1 + \frac{0.2}{6}\big(1 + 2(1.2) + 2(1.22) + 1.444\big) = 1 + \frac{0.2}{6}(7.284) = 1.242805.} \]
The exact solution gives $y(0.2) = 2e^{0.2} - 0.2 - 1 = 1.242806$, so RK4 is
correct to six figures. Compare the one-step estimates:
\[ \textcolor{workpurple}{\text{Euler } 1.2 \;(\text{err }\approx 4.3\times 10^{-2}), \quad
   \text{Improved } 1.24 \;(\text{err }\approx 2.8\times 10^{-3}), \quad
   \text{RK4 } 1.242805 \;(\text{err }\approx 10^{-6}).} \]

\vspace{0.8em}
\begin{conceptbox}[Unit 5 Key Takeaway]
Accuracy is bought with slopes per step. \textcolor{accentorange}{Euler} uses one
(first order); \textcolor{accentorange}{improved/modified Euler and Heun} average
two (second order); \textcolor{accentorange}{RK4} weights four as
$\textcolor{accentorange}{\tfrac{h}{6}(k_1+2k_2+2k_3+k_4)}$ (fourth order). On
$y'=t+y$ over a single $h=0.2$ step, the errors fall from $\sim 10^{-2}$ to
$\sim 10^{-6}$ across the ladder --- and \textcolor{accentorange}{adaptive stepping}
spends that accuracy budget only where the solution demands it.
\end{conceptbox}

\end{document}
