% Unit 6: Multivariable Calculus Foundations
% Master Unit Reference Guide (Cheat Sheet, Problems, and Solutions)
% Prepared by Staples Education
% Content informed by the local curriculum modules (6.1 to 6.10) and the
% QUIZ_DATA micro-practice items for Unit 6.
% Shared style harness pulled from the unified preamble input.

\input{app/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 6: Multivariable Calculus Foundations}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Master Unit Reference Guide: Cheat Sheet, Problems, and Solutions}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Table of contents (master guide)
% ---------------------------------------------------------------------------
\tableofcontents

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 6 at a Glance: Core Sub-Topics]
Differential equations in more than one variable need the language of
multivariable calculus. This unit builds that bridge: we learn to differentiate
one variable at a time, to package those partials into a gradient, to chain
through intermediate variables, to locate and classify extrema, and to read the
geometry of level curves --- the same geometry that decides when a differential
$M\,dx+N\,dy$ is \emph{exact}.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Partial Derivatives} --- differentiate in one variable while
          holding the others constant; $f_x,\ f_y$.
    \item \textbf{Higher-Order and Mixed Partials} --- Clairaut's theorem:
          $f_{xy}=f_{yx}$ when the second partials are continuous.
    \item \textbf{Gradient and Directional Derivatives} --- $\nabla f$ collects
          the partials; $D_{\mathbf u}f=\nabla f\cdot\mathbf u$ measures change
          along a unit direction.
    \item \textbf{The Multivariable Chain Rule} --- sum one product term per path
          from the output down to the independent variable.
    \item \textbf{Critical Points and Optimization} --- set $\nabla f=\mathbf 0$,
          then classify with the second-derivative (Hessian) test.
    \item \textbf{Level Curves and Exactness} --- $f(x,y)=c$, $\nabla f$ normal to
          the curve, and the exact-differential test $\partial_y M=\partial_x N$.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{6}   % subsections auto-number as the unit modules 6.1, 6.2, ...

\subsection{Partial Derivatives}

A partial derivative differentiates a multivariable function in \emph{one}
variable while treating every other variable as a constant symbol. It measures
the instantaneous rate of change of $f$ along a single coordinate axis.

\begin{formulabox}[Definition of the Partial Derivatives]
\[ f_x = \frac{\partial f}{\partial x} = \lim_{h\to 0}\frac{f(x+h,y)-f(x,y)}{h},
   \qquad
   f_y = \frac{\partial f}{\partial y} = \lim_{h\to 0}\frac{f(x,y+h)-f(x,y)}{h}. \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item To compute $f_x$, \textbf{hold $y$ constant} and differentiate in $x$
          with the ordinary rules; any pure-$y$ term is a constant and dies.
    \item Holding a variable \emph{constant} is not the same as
          \emph{setting it to zero}: the symbol stays in the answer.
\end{itemize}
\end{formulabox}

\begin{examplebox}[Worked Example 6.1: First partials of $f=x^2y+y^3$]
Treat $y$ as constant for $f_x$, then $x$ as constant for $f_y$:
\[ \textcolor{workpurple}{f_x = 2xy + 0 = 2xy, \qquad f_y = x^2 + 3y^2.} \]
The $y^3$ term vanishes under $\partial_x$, while the $x^2$ term contributes its
coefficient $x^2$ under $\partial_y$.
\end{examplebox}

\subsection{Higher-Order and Mixed Partials}

Each first partial can itself be differentiated again, giving four second-order
partials: two \emph{pure} ($f_{xx},f_{yy}$) and two \emph{mixed}
($f_{xy},f_{yx}$). Clairaut's theorem says the mixed pair agrees whenever the
function is smooth enough.

\begin{formulabox}[Clairaut's Theorem on Mixed Partials]
If the second partial derivatives of $f$ are \textbf{continuous} near a point,
then the mixed partials are equal there:
\[ f_{xy} = \frac{\partial^2 f}{\partial y\,\partial x}
          = \frac{\partial^2 f}{\partial x\,\partial y} = f_{yx}. \]
The order of differentiation does not matter --- a function of two variables has
\textbf{four} second partials but only \textbf{three} distinct values.
\end{formulabox}

\begin{examplebox}[Worked Example 6.2: Mixed partials of $f=x^2y^3$]
Differentiate in $x$ first, then in $y$ --- and the reverse:
\[ \textcolor{workpurple}{f_x = 2xy^3 \;\Rightarrow\; f_{xy} = 6xy^2,
   \qquad f_y = 3x^2y^2 \;\Rightarrow\; f_{yx} = 6xy^2.} \]
Both routes land on $6xy^2$, exactly as Clairaut guarantees.
\end{examplebox}

\subsection{The Gradient and Directional Derivatives}

The gradient packages the partials into a single vector that points in the
direction of steepest increase. Dotting it with a \emph{unit} direction gives
the rate of change along that direction --- the directional derivative.

\begin{formulabox}[Gradient, Directional Derivative, and Steepest Ascent]
\[ \nabla f = \left(\frac{\partial f}{\partial x},\ \frac{\partial f}{\partial y}\right),
   \qquad
   D_{\mathbf u} f = \nabla f \cdot \mathbf u \quad(\mathbf u \text{ a unit vector}). \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item The \textbf{maximum} rate of increase at a point is $\lvert\nabla f\rvert$,
          attained in the direction $\mathbf u = \nabla f/\lvert\nabla f\rvert$.
    \item $\nabla f$ is \textbf{perpendicular} to the level curve through the point.
\end{itemize}
\end{formulabox}

\begin{warningbox}[Common Pitfall: The Direction Must Be a Unit Vector]
$D_{\mathbf u}f=\nabla f\cdot\mathbf u$ is only valid when $\mathbf u$ is
\textbf{normalized}. Given a raw direction $\mathbf v$, divide first:
$\mathbf u=\mathbf v/\lvert\mathbf v\rvert$. Skipping the normalization rescales
the answer by $\lvert\mathbf v\rvert$ and silently corrupts every directional
derivative and max-rate computation that follows.
\end{warningbox}

\begin{examplebox}[Worked Example 6.3: Directional derivative of $f=x^2+xy+y^2$]
At $(1,2)$, $\nabla f=(2x+y,\,x+2y)=(4,5)$. Along $\mathbf v=(3,4)$, normalize to
$\mathbf u=(\tfrac35,\tfrac45)$:
\[ \textcolor{workpurple}{D_{\mathbf u}f = (4,5)\cdot\left(\tfrac35,\tfrac45\right)
   = \tfrac{12}{5}+\tfrac{20}{5} = \tfrac{32}{5} = 6.4.} \]
The steepest ascent there points along $(4,5)$ with rate
$\lvert\nabla f\rvert=\sqrt{41}\approx 6.40$.
\end{examplebox}

\subsection{The Multivariable Chain Rule}

When the inputs of $f$ are themselves functions, differentiate by summing one
product term for every path from the output down to the variable you are
differentiating with respect to. A small \emph{tree diagram} bookkeeps the paths.

\begin{methodbox}[Chain Rule Trees: Counting the Paths]
\textbf{One independent variable} $z=f(x(t),y(t))$ --- two paths to $t$:
\[ \frac{dz}{dt} = \frac{\partial f}{\partial x}\frac{dx}{dt}
                 + \frac{\partial f}{\partial y}\frac{dy}{dt}
                 = \nabla f \cdot (x',y'). \]
\textbf{Two independent variables} $z=f(x(s,t),y(s,t))$ --- two paths to each:
\[ \frac{\partial z}{\partial s} = f_x\,x_s + f_y\,y_s,
   \qquad
   \frac{\partial z}{\partial t} = f_x\,x_t + f_y\,y_t. \]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item Draw a branch from $z$ to each intermediate variable ($x,y$).
    \item Extend a branch from each intermediate variable to the target ($t$).
    \item \textbf{Multiply} along each branch, then \textbf{add} the branches.
\end{enumerate}
\end{methodbox}

\begin{examplebox}[Worked Example 6.4: Chain rule for $z=x^2y$, $x=t^2$, $y=t^3$]
With $z_x=2xy$, $z_y=x^2$, $x'=2t$, $y'=3t^2$:
\[ \textcolor{workpurple}{\frac{dz}{dt} = (2xy)(2t) + (x^2)(3t^2)
   = (2t^5)(2t) + (t^4)(3t^2) = 4t^6 + 3t^6 = 7t^6.} \]
The direct substitution $z=t^4\cdot t^3=t^7$ confirms $dz/dt=7t^6$.
\end{examplebox}

\subsection{Critical Points and Optimization}

Interior extrema of a smooth $f(x,y)$ occur where the gradient vanishes. The
\emph{second-derivative test} then uses the sign of the Hessian discriminant to
sort each critical point into a minimum, maximum, or saddle.

\begin{methodbox}[Locating and Classifying Critical Points]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item \textbf{Solve} $\nabla f=\mathbf 0$, i.e.\ $f_x=0$ and $f_y=0$
          simultaneously, to find all critical points.
    \item \textbf{Form} the discriminant
          $D = f_{xx}f_{yy} - \left(f_{xy}\right)^2$ at each point.
    \item \textbf{Classify} with $D$ and $f_{xx}$ (Step in the box below).
\end{enumerate}
\end{methodbox}

\begin{formulabox}[The Second-Derivative (Hessian) Test]
\[ D = f_{xx}\,f_{yy} - \left(f_{xy}\right)^2. \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item $D>0,\ f_{xx}>0$: \textbf{local minimum}.
    \item $D>0,\ f_{xx}<0$: \textbf{local maximum}.
    \item $D<0$: \textbf{saddle point} (no extremum).
    \item $D=0$: the test is \textbf{inconclusive}.
\end{itemize}
\end{formulabox}

\begin{warningbox}[Common Pitfall: Reading the Sign Boundary Wrong]
The sign of $f_{xx}$ only decides min vs.\ max \emph{after} $D>0$ confirms an
extremum. When $D<0$ the point is a \textbf{saddle} regardless of $f_{xx}$, and
when $D=0$ the test says \emph{nothing} --- do not declare a saddle by default.
A single mishandled sign flips a minimum into a maximum.
\end{warningbox}

\subsection{Level Curves, Gradients, and Exactness}

Level curves $f(x,y)=c$ slice a surface into contour lines; the gradient cuts
across them at right angles. This same gradient structure is exactly what makes a
first-order differential \emph{exact} --- the payoff that carries into Unit 7.

\begin{formulabox}[Level Curves and the Exact-Differential Test]
A \textbf{level curve} is the set $\{(x,y):f(x,y)=c\}$; $\nabla f$ is normal to
it. A differential $M(x,y)\,dx+N(x,y)\,dy$ is \textbf{exact} on a simply
connected region exactly when
\[ \frac{\partial M}{\partial y} = \frac{\partial N}{\partial x}, \]
in which case it equals $df$ for a potential $f$ with $f_x=M,\ f_y=N$ --- and the
field $\mathbf F=(M,N)=\nabla f$ is \textbf{conservative}.
\end{formulabox}

\begin{warningbox}[Common Pitfall: Path Dependence and the Wrong Cross Partials]
The exactness/conservative test compares the \emph{cross} partials
$\partial_y M$ and $\partial_x N$ --- not $\partial_x M$ and $\partial_y N$. When
they differ the field is path-dependent: $\int_C\mathbf F\cdot d\mathbf r$
depends on the route, so there is \emph{no} single potential to read off.
\end{warningbox}

% ===========================================================================
% PART II: REVIEW GUIDE (PRACTICE QUESTIONS)
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part II --- Review Guide: Practice Problems}
\section*{Part II --- Review Guide: Practice Problems}

Work the following ten problems in order. They climb from first and mixed partial
derivatives, through the gradient and directional derivatives, into the
multivariable chain rule, and close with a full optimization sweep --- finding
and classifying critical points with the second-derivative test. Solutions follow
in Part III.

\begin{enumerate}[leftmargin=2em]
    \item For $f(x,y) = x^2 y + y^3$, compute both first partials $f_x$ and
          $f_y$.

    \item For $f(x,y) = x^2 y^3$, compute the mixed partials $f_{xy}$ and
          $f_{yx}$ and confirm Clairaut's theorem.

    \item For $f(x,y) = x^2 + xy + y^2$, find the gradient $\nabla f$ and
          evaluate it at the point $(1,2)$.

    \item Using the gradient from Problem 3, compute the directional derivative of
          $f$ at $(1,2)$ in the direction of $\mathbf v = (3,4)$. Remember to
          normalize.

    \item At the same point $(1,2)$, state the direction of steepest ascent of
          $f=x^2+xy+y^2$ and the maximum rate of increase there.

    \item For $z = x^2 + y^2$ with $x = \cos t$ and $y = \sin t$, use the chain
          rule to find $\dfrac{dz}{dt}$, and explain the result geometrically.

    \item For $z = x^2 y$ with $x = s + t$ and $y = st$, use the chain-rule tree
          to find $\dfrac{\partial z}{\partial s}$.

    \item Find and classify the single critical point of
          $f(x,y) = x^2 + y^2 - 4x - 6y + 13$, and give the extreme value.

    \item Find and classify the critical point of $f(x,y) = x^2 - y^2$ using the
          second-derivative test.

    \item \textbf{(Capstone)} Find \emph{all} critical points of
          $f(x,y) = x^3 - 3x + y^2$ and classify each with the second-derivative
          test, reporting any extreme values.
\end{enumerate}

% ===========================================================================
% PART III: SOLUTIONS
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part III --- Complete Solutions}
\section*{Part III --- Complete Solutions}

\noindent\textbf{Solution 1.}\quad Hold $y$ constant for $f_x$, then $x$ constant
for $f_y$. The pure-$y$ term $y^3$ vanishes under $\partial_x$:
\[ \textcolor{workpurple}{f_x = 2xy, \qquad f_y = x^2 + 3y^2.} \]

\medskip
\noindent\textbf{Solution 2.}\quad Differentiate in $x$ first, then $y$ --- and
in the reverse order:
\[ \textcolor{workpurple}{f_x = 2xy^3 \;\Rightarrow\; f_{xy} = 6xy^2,
   \qquad f_y = 3x^2y^2 \;\Rightarrow\; f_{yx} = 6xy^2.} \]
Both equal $6xy^2$, so $f_{xy}=f_{yx}$ --- Clairaut's theorem holds because the
second partials are continuous everywhere.

\medskip
\noindent\textbf{Solution 3.}\quad Differentiate componentwise:
\[ \textcolor{workpurple}{\nabla f = (2x + y,\; x + 2y),
   \qquad \nabla f(1,2) = (2+2,\; 1+4) = (4,5).} \]

\medskip
\noindent\textbf{Solution 4.}\quad The raw direction $\mathbf v=(3,4)$ has
$\lvert\mathbf v\rvert=\sqrt{9+16}=5$, so the unit direction is
$\mathbf u=(\tfrac35,\tfrac45)$. Then
\[ \textcolor{workpurple}{D_{\mathbf u}f = \nabla f(1,2)\cdot\mathbf u
   = (4,5)\cdot\left(\tfrac35,\tfrac45\right)
   = \tfrac{12}{5} + \tfrac{20}{5} = \tfrac{32}{5} = 6.4.} \]

\medskip
\noindent\textbf{Solution 5.}\quad Steepest ascent points along the gradient
itself, $\nabla f(1,2)=(4,5)$, i.e.\ the unit direction
$(4,5)/\sqrt{41}$. The maximum rate of increase is the gradient's magnitude:
\[ \textcolor{workpurple}{\lvert\nabla f(1,2)\rvert = \sqrt{4^2+5^2}
   = \sqrt{41} \approx 6.40.} \]

\medskip
\noindent\textbf{Solution 6.}\quad With $z_x=2x$, $z_y=2y$, $x'=-\sin t$,
$y'=\cos t$:
\[ \textcolor{workpurple}{\frac{dz}{dt} = (2x)(-\sin t) + (2y)(\cos t)
   = 2\cos t(-\sin t) + 2\sin t(\cos t) = 0.} \]
Geometrically the path $(\cos t,\sin t)$ rides the unit circle, where
$z=x^2+y^2=1$ is constant --- so its rate of change is zero, and the velocity is
tangent to the level curve, perpendicular to $\nabla z$.

\medskip
\noindent\textbf{Solution 7.}\quad The tree has two branches to $s$:
$\dfrac{\partial z}{\partial s}=z_x\,x_s+z_y\,y_s$ with $z_x=2xy$, $z_y=x^2$,
$x_s=1$, $y_s=t$:
\[ \textcolor{workpurple}{\frac{\partial z}{\partial s}
   = (2xy)(1) + (x^2)(t)
   = 2(s+t)(st) + (s+t)^2\,t.} \]
Factoring $t(s+t)$ gives
\[ \textcolor{workpurple}{\frac{\partial z}{\partial s}
   = t(s+t)\big[\,2s + (s+t)\,\big] = t(s+t)(3s+t).} \]

\medskip
\noindent\textbf{Solution 8.}\quad Set the gradient to zero:
\[ \textcolor{workpurple}{f_x = 2x - 4 = 0 \Rightarrow x = 2,
   \qquad f_y = 2y - 6 = 0 \Rightarrow y = 3,} \]
so the only critical point is $(2,3)$. The second partials are $f_{xx}=2$,
$f_{yy}=2$, $f_{xy}=0$, giving
\[ \textcolor{workpurple}{D = f_{xx}f_{yy} - (f_{xy})^2 = (2)(2) - 0 = 4 > 0,
   \quad f_{xx} = 2 > 0,} \]
a \textbf{local minimum}. The value is
$\textcolor{workpurple}{f(2,3) = 4 + 9 - 8 - 18 + 13 = 0}$.

\medskip
\noindent\textbf{Solution 9.}\quad Here $f_x=2x=0$ and $f_y=-2y=0$ give the lone
critical point $(0,0)$. With $f_{xx}=2$, $f_{yy}=-2$, $f_{xy}=0$:
\[ \textcolor{workpurple}{D = (2)(-2) - 0 = -4 < 0,} \]
so $(0,0)$ is a \textbf{saddle point} --- $f$ rises along $x$ and falls along
$y$, and the sign of $f_{xx}$ is irrelevant once $D<0$.

\medskip
\noindent\textbf{Solution (Capstone).}\quad Set both partials to zero:
\[ \textcolor{workpurple}{f_x = 3x^2 - 3 = 0 \Rightarrow x = \pm 1,
   \qquad f_y = 2y = 0 \Rightarrow y = 0,} \]
giving critical points $(1,0)$ and $(-1,0)$. The second partials are
$f_{xx}=6x$, $f_{yy}=2$, $f_{xy}=0$, so $D=12x$.
\[ \textcolor{workpurple}{(1,0):\ D = 12 > 0,\ f_{xx} = 6 > 0
   \Rightarrow \text{local minimum},\quad f(1,0) = 1 - 3 = -2.} \]
\[ \textcolor{workpurple}{(-1,0):\ D = -12 < 0
   \Rightarrow \text{saddle point}.} \]
So $f$ has a local minimum value $-2$ at $(1,0)$ and a saddle at $(-1,0)$; the
cubic term means there is no global extremum.

\vspace{0.8em}
\begin{conceptbox}[Unit 6 Key Takeaway]
The workflow is uniform: \textcolor{accentorange}{partials} build the
\textcolor{accentorange}{gradient $\nabla f$}; the gradient drives directional
derivatives $\textcolor{accentorange}{D_{\mathbf u}f=\nabla f\cdot\mathbf u}$
(always on a \emph{unit} $\mathbf u$) and points along steepest ascent
$\textcolor{accentorange}{\lvert\nabla f\rvert}$; the
\textcolor{accentorange}{chain rule} sums one term per path; and
\textcolor{accentorange}{optimization} solves $\nabla f=\mathbf 0$ then sorts each
point with $\textcolor{accentorange}{D=f_{xx}f_{yy}-f_{xy}^{\,2}}$ --- $D>0$ an
extremum (sign of $f_{xx}$ decides min vs.\ max), $D<0$ a saddle.
\end{conceptbox}

\end{document}
