% Unit 4: Autonomous Equations, Equilibrium Solutions, and Stability
% Master Unit Reference Guide (Cheat Sheet, Problems, and Solutions)
% Prepared by Staples Education
% Content informed by the local curriculum modules (4.1 to 4.6) and the
% QUIZ_DATA micro-practice items for Unit 4.
% Shared style harness pulled from the unified preamble input.

\input{app/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 4: Autonomous Equations, Equilibrium Solutions, and Stability}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Master Unit Reference Guide: Cheat Sheet, Problems, and Solutions}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Table of contents (master guide)
% ---------------------------------------------------------------------------
\tableofcontents

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 4 at a Glance: Core Sub-Topics]
This unit studies first-order equations \emph{qualitatively}: when the right
side depends on $y$ alone, the long-run behavior can be read off without ever
solving. It then closes with the Bernoulli family, the one nonlinear form that a
clever substitution turns linear.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Equilibrium Solutions and Stability} --- constant solutions
          $f(c)=0$, and the sign of $f'(c)$ that makes them attract or repel.
    \item \textbf{Autonomous Equations and the Phase Line} --- $y'=f(y)$ as a
          one-dimensional flow, with arrows set by the sign of $f$.
    \item \textbf{The Logistic Equation} --- $P'=rP(1-P/K)$, its two equilibria,
          and the carrying capacity $K$.
    \item \textbf{Bernoulli Differential Equations} --- the form
          $y'+P(x)y=Q(x)y^{n}$ and the linearizing substitution $v=y^{1-n}$.
    \item \textbf{Solving Bernoulli Equations} --- the full divide, substitute,
          solve-linear, and back-substitute workflow.
    \item \textbf{Applications of Bernoulli Equations} --- population models and
          the fluid-energy form along a streamline.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{4}   % subsections auto-number as the unit modules 4.1, 4.2, ...

\subsection{Equilibrium Solutions and Stability}

An \emph{equilibrium} (steady state) of $\frac{dy}{dt} = f(y)$ is a constant
solution $y = c$. A constant has zero derivative, so the rate must vanish there:
$f(c) = 0$. Whether nearby solutions return to $c$ or flee from it is decided by
the sign of the slope $f'(c)$.

\begin{formulabox}[Equilibria and the Linearized Stability Test]
$y = c$ is an equilibrium of $\dfrac{dy}{dt}=f(y)$ exactly when
\[ f(c) = 0. \]
Classify it by the sign of $f'(c)$:
\[
\begin{array}{ll}
f'(c) < 0 & \text{stable (a \textbf{sink}) --- solutions return}\\
f'(c) > 0 & \text{unstable (a \textbf{source}) --- solutions flee}\\
f'(c) = 0 & \text{test fails; check the sign of } f \text{ on each side}
\end{array}
\]
When $f$ touches zero \emph{without changing sign}, the point is
\textbf{semi-stable}: it attracts from one side and repels from the other.
\end{formulabox}

\begin{examplebox}[Worked Example 4.1: Classify the equilibria of $\frac{dy}{dt}=y-y^{3}$]
\textbf{Step 1.} Set $f(y)=0$ and factor:
\[ \textcolor{workpurple}{y - y^{3} = y(1-y^{2}) = y(1-y)(1+y) = 0 \;\Longrightarrow\; y = -1,\,0,\,1.} \]
\textbf{Step 2.} Test each root with $f'(y) = 1 - 3y^{2}$:
\[ \textcolor{workpurple}{f'(0) = 1 > 0\ (\text{unstable}), \qquad f'(\pm 1) = -2 < 0\ (\text{stable}).} \]
So $y=\pm 1$ are sinks and $y=0$ is a source.
\end{examplebox}

\subsection{Autonomous Equations and the Phase Line}

An equation is \emph{autonomous} when its right side depends on $y$ alone, with
no explicit $t$: $\frac{dy}{dt} = f(y)$. Because nothing depends on the clock,
any time-shift $y(t-c)$ of a solution is again a solution --- solution curves are
horizontal translates of one another. All the behavior collapses onto a single
\emph{phase line}: a $y$-axis with arrows.

\begin{formulabox}[Reading the Phase Line]
Mark the equilibria ($f=0$) on the $y$-axis. Between them the sign of $f$ sets
the arrow:
\[ f(y) > 0 \Rightarrow \text{arrow up},\qquad f(y) < 0 \Rightarrow \text{arrow down}. \]
\[
\begin{array}{ll}
\text{arrows in from both sides} & \text{stable sink}\\
\text{arrows out from both sides} & \text{unstable source}\\
\text{one in, one out} & \text{semi-stable}
\end{array}
\]
\end{formulabox}

\begin{examplebox}[Worked Example 4.2: Phase line for $\frac{dy}{dt}=y^{2}-4$]
\textbf{Step 1.} Equilibria where the rate vanishes:
\[ \textcolor{workpurple}{y^{2} - 4 = 0 \;\Longrightarrow\; y = -2,\,2.} \]
\textbf{Step 2.} Test the rate sign in each interval:
\[ \textcolor{workpurple}{y<-2:\ (+)\uparrow, \qquad -2<y<2:\ (-)\downarrow, \qquad y>2:\ (+)\uparrow.} \]
Arrows converge on $-2$ (stable) and diverge from $2$ (unstable).
\end{examplebox}

\subsection{The Logistic Equation}

The logistic model corrects exponential growth with a crowding term, so growth
slows as the population nears its ceiling. For small $P$ it behaves like
exponential growth $P' \approx rP$; near the ceiling the growth shuts off.

\begin{formulabox}[The Logistic Equation and Its Equilibria]
\[ \frac{dP}{dt} = rP\!\left(1 - \frac{P}{K}\right), \qquad r>0. \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item Equilibria: $P = 0$ (\textbf{unstable}) and $P = K$ (\textbf{stable}).
    \item $K$ is the \textbf{carrying capacity} --- the population level the
          solution approaches.
    \item Growth rate is greatest at the midpoint $P = \tfrac{K}{2}$, where the
          downward parabola in $P$ peaks.
\end{itemize}
\end{formulabox}

\begin{warningbox}[Common Pitfall: Carrying-Capacity Traps]
$K$ is a \emph{population level}, not a growth rate --- the intrinsic rate is the
constant $r$. The peak growth sits at $\tfrac{K}{2}$, not at $K$ (where growth is
zero). And do not drop the $P=0$ equilibrium: with $r>0$ it is \emph{unstable},
while $P=K$ is the \emph{stable} ceiling --- reversing these two is the classic
sign error.
\end{warningbox}

\begin{examplebox}[Worked Example 4.3: Evaluate the logistic rate]
With $r = 0.5$, $K = 100$, find $\frac{dP}{dt}$ at $P = 25$:
\[ \textcolor{workpurple}{\frac{dP}{dt} = 0.5 \cdot 25\!\left(1 - \frac{25}{100}\right) = 12.5 \cdot 0.75 = 9.375.} \]
The population is below $\tfrac{K}{2}=50$, so it is still accelerating toward the
fastest-growth point.
\end{examplebox}

\subsection{Bernoulli Differential Equations}

A \emph{Bernoulli equation} looks like a first-order linear equation but carries
a power of $y$ on the right. That single nonlinear term is exactly what the
substitution $v = y^{1-n}$ removes.

\begin{formulabox}[Standard Form and Linearizing Substitution]
\[ y' + P(x)\,y = Q(x)\,y^{n}. \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item For $n = 0$ or $n = 1$ the equation is already linear --- no
          substitution needed.
    \item For any other $n$, set
          \[ v = y^{1-n}, \qquad \frac{dv}{dx} = (1-n)\,y^{-n}\,\frac{dy}{dx}, \]
          which turns the equation into a \textbf{first-order linear} equation
          in $v$.
\end{itemize}
\end{formulabox}

\begin{methodbox}[Why $v=y^{1-n}$ Works]
Dividing $y' + P y = Q y^{n}$ by $y^{n}$ gives $y^{-n}y' + P\,y^{1-n} = Q$. The
group $y^{1-n}$ is exactly $v$, and $y^{-n}y'$ is $\frac{1}{1-n}\,v'$, so the
whole equation becomes
\[ \frac{1}{1-n}\,v' + P\,v = Q \quad\Longrightarrow\quad v' + (1-n)P\,v = (1-n)Q, \]
a clean linear equation ready for an integrating factor.
\end{methodbox}

\subsection{Solving Bernoulli Equations, Worked Examples}

The solution is a fixed four-step pipeline: clear the power, substitute, solve
the linear equation with an integrating factor, then undo the substitution.

\begin{methodbox}[The Bernoulli Solution Algorithm]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item \textbf{Divide} every term by $y^{n}$.
    \item \textbf{Substitute} $v = y^{1-n}$, $v' = (1-n)y^{-n}y'$ to get
          $v' + (1-n)P\,v = (1-n)Q$.
    \item \textbf{Solve} the linear equation with $\mu = e^{\int (1-n)P\,dx}$.
    \item \textbf{Back-substitute} $v = y^{1-n}$ and solve for $y$.
\end{enumerate}
\end{methodbox}

\begin{examplebox}[Worked Example 4.5: Reduce $y' + y = xy^{3}$]
\textbf{Step 1.} Here $n = 3$, so $v = y^{1-3} = y^{-2}$ and
$v' = -2y^{-3}y'$.
\[ \textcolor{workpurple}{\text{Divide by } y^{3}:\quad y^{-3}y' + y^{-2} = x.} \]
\textbf{Step 2.} Substitute, using $y^{-3}y' = -\tfrac{1}{2}v'$:
\[ \textcolor{workpurple}{-\tfrac{1}{2}v' + v = x \;\Longrightarrow\; v' - 2v = -2x,} \]
a first-order linear equation in $v$ ready for $\mu = e^{-2x}$.
\end{examplebox}

\subsection{Applications of Bernoulli Equations}

The name attaches to two distinct ideas. The Bernoulli \emph{differential
equation} models constrained growth and resistive dynamics; the Bernoulli
\emph{principle} in fluids is an energy balance along a streamline.

\begin{formulabox}[The Fluid Bernoulli Equation]
Along a streamline, mechanical energy per unit volume is conserved:
\[ P + \tfrac{1}{2}\rho v^{2} + \rho g h = \text{constant}. \]
The three terms are pressure, kinetic, and potential energy density. With the
total fixed, a faster flow ($v\uparrow$) forces a lower pressure ($P\downarrow$);
because the kinetic term scales with $v^{2}$, doubling the speed quadruples it.
\end{formulabox}

\begin{warningbox}[Common Pitfall: Two Different ``Bernoulli'' Equations]
The Bernoulli \emph{ODE} $y'+P(x)y = Q(x)y^{n}$ and the fluid Bernoulli
\emph{principle} share only a name. Do not import the substitution $v=y^{1-n}$
into a fluids problem, or treat $\tfrac{1}{2}\rho v^{2}$ as something to
linearize --- one is a solution method, the other a conservation law.
\end{warningbox}

% ===========================================================================
% PART II: REVIEW GUIDE (PRACTICE QUESTIONS)
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part II --- Review Guide: Practice Problems}
\section*{Part II --- Review Guide: Practice Problems}

Work the following ten problems in order. They climb from locating equilibria,
through stability tests and phase lines, into the logistic model, and finish with
the Bernoulli substitution $v = y^{1-n}$ and a full solve. Solutions follow in
Part III.

\begin{enumerate}[leftmargin=2em]
    \item Find all equilibrium solutions of $\dfrac{dy}{dt} = y(y-3)$.

    \item Classify the equilibrium $y = 2$ of $\dfrac{dy}{dt} = y - 2$ using the
          sign of $f'$.

    \item For $\dfrac{dy}{dt} = y - y^{3}$, find every equilibrium and classify
          each with the linearized test.

    \item Classify the equilibrium $y = 1$ of $\dfrac{dy}{dt} = (y-1)^{2}$. Why
          does the $f'$ test alone not settle it?

    \item Build the phase line for $\dfrac{dy}{dt} = y^{2} - 4$: mark the
          equilibria, set the arrows, and classify each.

    \item For the logistic equation $\dfrac{dP}{dt} = rP\!\left(1-\dfrac{P}{K}\right)$
          with $r > 0$, find both equilibria, classify them, and give the
          population of fastest growth.

    \item With $r = 0.5$ and $K = 100$, compute $\dfrac{dP}{dt}$ at $P = 25$ and
          state the population at which the growth rate is greatest.

    \item For $y' + y = xy^{3}$, identify $n$, write the substitution $v$, and
          compute $\dfrac{dv}{dx}$ in terms of $y'$.

    \item Reduce $y' + \dfrac{1}{x}\,y = x\,y^{2}$ to a first-order \emph{linear}
          equation in $v = y^{1-n}$ (do not solve it yet).

    \item \textbf{(Capstone)} Solve the Bernoulli equation $y' + y = y^{2}$
          completely: reduce it with $v = y^{-1}$, solve the linear equation, and
          back-substitute for $y$.
\end{enumerate}

% ===========================================================================
% PART III: SOLUTIONS
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part III --- Complete Solutions}
\section*{Part III --- Complete Solutions}

\noindent\textbf{Solution 1.}\quad Equilibria are the roots of $f(y)=0$:
\[ \textcolor{workpurple}{y(y-3) = 0 \;\Longrightarrow\; y = 0 \text{ and } y = 3.} \]

\medskip
\noindent\textbf{Solution 2.}\quad Here $f(y) = y - 2$, so $f'(y) = 1$:
\[ \textcolor{workpurple}{f'(2) = 1 > 0 \;\Longrightarrow\; \text{unstable (a source).}} \]
A positive slope of $f$ at the equilibrium repels nearby solutions.

\medskip
\noindent\textbf{Solution 3.}\quad Factor for the equilibria, then apply
$f'(y) = 1 - 3y^{2}$:
\[ \textcolor{workpurple}{y - y^{3} = y(1-y)(1+y) = 0 \;\Longrightarrow\; y = -1,\,0,\,1.} \]
\[ \textcolor{workpurple}{f'(0) = 1 > 0\ (\text{unstable}), \qquad f'(\pm 1) = 1 - 3 = -2 < 0\ (\text{stable}).} \]
So $y = \pm 1$ are sinks and $y = 0$ is a source.

\medskip
\noindent\textbf{Solution 4.}\quad The only equilibrium is $y = 1$. The
linearized test is inconclusive because
\[ \textcolor{workpurple}{f'(y) = 2(y-1) \;\Longrightarrow\; f'(1) = 0.} \]
Fall back on the sign of $f(y) = (y-1)^{2}$, which is $\ge 0$ everywhere and
\emph{never} changes sign across $y = 1$:
\[ \textcolor{workpurple}{y<1:\ f>0\ (\uparrow), \qquad y>1:\ f>0\ (\uparrow).} \]
Solutions rise toward $1$ from below but away from $1$ above, so $y = 1$ is
\textbf{semi-stable}.

\medskip
\noindent\textbf{Solution 5.}\quad Equilibria where the rate vanishes:
\[ \textcolor{workpurple}{y^{2} - 4 = 0 \;\Longrightarrow\; y = -2,\,2.} \]
Test the rate sign on each interval:
\[ \textcolor{workpurple}{y<-2:\ (+)\,\uparrow, \qquad -2<y<2:\ (-)\,\downarrow, \qquad y>2:\ (+)\,\uparrow.} \]
Arrows converge on $-2$ (\textbf{stable}) and diverge from $2$
(\textbf{unstable}).

\medskip
\noindent\textbf{Solution 6.}\quad Set the product to zero:
\[ \textcolor{workpurple}{rP\!\left(1-\frac{P}{K}\right) = 0 \;\Longrightarrow\; P = 0 \text{ and } P = K.} \]
Classify with $f'(P) = r\!\left(1 - \dfrac{2P}{K}\right)$:
\[ \textcolor{workpurple}{f'(0) = r > 0\ (\text{unstable}), \qquad f'(K) = r(1-2) = -r < 0\ (\text{stable}).} \]
The growth rate, a downward parabola in $P$ with roots $0$ and $K$, peaks at the
midpoint:
\[ \textcolor{workpurple}{P = \frac{K}{2}.} \]

\medskip
\noindent\textbf{Solution 7.}\quad Substitute into $rP\!\left(1-\frac{P}{K}\right)$:
\[ \textcolor{workpurple}{\frac{dP}{dt} = 0.5 \cdot 25\!\left(1 - \frac{25}{100}\right) = 12.5 \cdot 0.75 = 9.375.} \]
The greatest growth occurs at
$\textcolor{workpurple}{P = \frac{K}{2} = 50}$.

\medskip
\noindent\textbf{Solution 8.}\quad The right side is $y^{3}$, so $n = 3$ and the
linearizing exponent is $1 - n = -2$:
\[ \textcolor{workpurple}{v = y^{1-n} = y^{-2}, \qquad \frac{dv}{dx} = (1-n)y^{-n}\,\frac{dy}{dx} = -2\,y^{-3}\,y'.} \]

\medskip
\noindent\textbf{Solution 9.}\quad Here $n = 2$, so $v = y^{-1}$ and
$v' = -y^{-2}y'$. Divide every term by $y^{2}$:
\[ \textcolor{workpurple}{y^{-2}y' + \frac{1}{x}\,y^{-1} = x.} \]
Replace $y^{-1} = v$ and $y^{-2}y' = -v'$:
\[ \textcolor{workpurple}{-v' + \frac{1}{x}\,v = x \;\Longrightarrow\; v' - \frac{1}{x}\,v = -x,} \]
a first-order linear equation in $v$.

\medskip
\noindent\textbf{Solution 10.}\quad With $n = 2$, set $v = y^{-1}$,
$v' = -y^{-2}y'$. Divide $y' + y = y^{2}$ by $y^{2}$:
\[ \textcolor{workpurple}{y^{-2}y' + y^{-1} = 1.} \]
Substitute $y^{-1} = v$, $y^{-2}y' = -v'$:
\[ \textcolor{workpurple}{-v' + v = 1 \;\Longrightarrow\; v' - v = -1.} \]
Integrating factor $\mu = e^{-x}$:
\[ \textcolor{workpurple}{\left(e^{-x}v\right)' = -e^{-x} \;\Longrightarrow\; e^{-x}v = e^{-x} + C.} \]
So $\textcolor{workpurple}{v = 1 + Ce^{x}}$, and back-substituting $v = y^{-1}$:
\[ \textcolor{workpurple}{y = \frac{1}{v} = \frac{1}{1 + Ce^{x}}.} \]

\vspace{0.8em}
\begin{conceptbox}[Unit 4 Key Takeaway]
Stability is settled by two signs: \textcolor{accentorange}{$f(c)=0$} locates an
equilibrium and \textcolor{accentorange}{the sign of $f'(c)$} (or, when that is
zero, the sign of $f$ on each side) classifies it. The
\textcolor{accentorange}{logistic equation} packages this with a stable carrying
capacity $K$ and fastest growth at $\tfrac{K}{2}$. Every Bernoulli equation
yields to \textcolor{accentorange}{$v = y^{1-n}$}, which trades the lone power of
$y$ for a clean linear solve.
\end{conceptbox}

\end{document}
