% Unit 9: Second-Order Linear Equations, Theory and Structure
% Master Unit Reference Guide (Cheat Sheet, Problems, and Solutions)
% Prepared by Staples Education
% Content informed by the local curriculum modules (9.1 to 9.8) and the
% QUIZ_DATA micro-practice items for Unit 9.
% Shared style harness pulled from the unified preamble input.

\input{app/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 9: Second-Order Linear Equations}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Master Unit Reference Guide: Cheat Sheet, Problems, and Solutions}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Table of contents (master guide)
% ---------------------------------------------------------------------------
\tableofcontents

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 9 at a Glance: Core Sub-Topics]
A second-order linear equation reads
$a(x)y'' + b(x)y' + c(x)y = g(x)$; this unit builds the homogeneous theory and the
constant-coefficient solution machine.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Existence, uniqueness, superposition} --- two side conditions
          pin a unique solution; linear combinations of solutions are solutions.
    \item \textbf{The Wronskian} --- $W(y_1,y_2)=y_1y_2'-y_2y_1'$ tests linear
          independence and certifies a fundamental set.
    \item \textbf{Abel's identity} --- the Wronskian itself solves a first-order
          ODE, so it is either never zero or identically zero.
    \item \textbf{Reduction of order} --- one known solution $y_1$ manufactures a
          second, independent $y_2$.
    \item \textbf{Constant-coefficient roots} --- the characteristic equation
          $ar^2+br+c=0$ splits into real-distinct, repeated, and complex cases.
    \item \textbf{Cauchy--Euler and higher order} --- the $x^m$ trick and the
          degree-$n$ characteristic polynomial.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{9}   % subsections auto-number as the unit modules 9.1, 9.2, ...

\subsection{Introduction to Second-Order Equations}

The general linear second-order equation is
$a(x)y'' + b(x)y' + c(x)y = g(x)$. It is \textbf{homogeneous} when $g(x)=0$ and
\textbf{nonhomogeneous} otherwise; its general solution carries \emph{two}
arbitrary constants.

\begin{formulabox}[Standard and Normalized Forms]
\[ a(x)y'' + b(x)y' + c(x)y = g(x), \qquad
   y'' + p(x)y' + q(x)y = r(x), \]
with $p=b/a$, $q=c/a$, $r=g/a$ where $a(x)\neq 0$. Linearity means $y$, $y'$, $y''$
each appear to the first power and are never multiplied together.
\end{formulabox}

\subsection{Existence, Uniqueness, Superposition, and Independence}

On any interval where $p$, $q$, and $r$ are continuous, the initial value problem
with $y(x_0)=y_0$, $y'(x_0)=y_0'$ has a \emph{unique} solution.

\begin{formulabox}[Superposition Principle]
If $y_1$ and $y_2$ each solve the homogeneous equation
$y''+p\,y'+q\,y=0$, then for \emph{any} constants $c_1,c_2$,
\[ y = c_1 y_1 + c_2 y_2 \]
is also a solution. The general solution is this two-parameter family precisely
when $y_1,y_2$ are linearly independent.
\end{formulabox}

\subsection{Linear Independence and the Wronskian}

Two solutions form a \textbf{fundamental set} exactly when they are linearly
independent, and the Wronskian is the determinant that tests it.

\begin{formulabox}[The Wronskian]
\[ W(y_1,y_2)(x) = \begin{vmatrix} y_1 & y_2 \\ y_1' & y_2' \end{vmatrix}
   = y_1 y_2' - y_2 y_1'. \]
If $W(x_0)\neq 0$ at any point, $y_1,y_2$ are \textbf{linearly independent} on the
interval and $\{y_1,y_2\}$ is a fundamental set.
\end{formulabox}

\subsection{Abel's Identity}

Abel's identity says the Wronskian of two solutions of
$y''+p(x)y'+q(x)y=0$ itself satisfies a first-order ODE, so its sign can never
change.

\begin{methodbox}[Abel's Formula]
For $y'' + p(x)\,y' + q(x)\,y = 0$,
\[ W(x) = W(x_0)\,\exp\!\left(-\int_{x_0}^{x} p(t)\,dt\right). \]
Because the exponential is never zero, $W$ is \textbf{either identically zero}
(dependent solutions) \textbf{or never zero} (independent) --- there is no middle
ground.
\end{methodbox}

\subsection{Reduction of Order}

Given one nonzero solution $y_1$ of the homogeneous equation, reduction of order
substitutes $y_2 = v(x)\,y_1$ and solves for $v$, producing a second independent
solution.

\begin{methodbox}[Reduction of Order Formula]
For $y'' + p(x)\,y' + q(x)\,y = 0$ with a known solution $y_1$,
\[ y_2 = y_1(x)\int \frac{e^{-\int p(x)\,dx}}{y_1(x)^2}\,dx. \]
The substitution $y_2=v y_1$ collapses the second-order equation to a
\emph{first-order} equation in $v'$, which the integral above solves directly.
\end{methodbox}

\subsection{Constant Coefficient Homogeneous Equations}

For $ay'' + by' + cy = 0$ with constants $a,b,c$, try $y=e^{rx}$ to get the
\textbf{characteristic equation} $ar^2 + br + c = 0$. The discriminant
$b^2-4ac$ selects one of three cases.

\begin{formulabox}[The Three Root Cases]
\[ ar^2 + br + c = 0, \qquad r = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}. \]
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Real distinct} ($b^2-4ac>0$), roots $r_1\neq r_2$:
          \[ y = c_1 e^{r_1 x} + c_2 e^{r_2 x}. \]
    \item \textbf{Repeated} ($b^2-4ac=0$), one root $r$:
          \[ y = (c_1 + c_2 x)\,e^{r x}. \]
    \item \textbf{Complex conjugate} ($b^2-4ac<0$), $r=\alpha\pm\beta i$:
          \[ y = e^{\alpha x}\big(c_1\cos\beta x + c_2\sin\beta x\big). \]
\end{itemize}
\end{formulabox}

\begin{warningbox}[Common Pitfall: The Repeated-Root Boundary]
The repeated case ($b^2-4ac=0$) is the boundary between real and complex roots.
Do \emph{not} write $y=(c_1+c_2)e^{rx}$ --- that is a single constant; the second
solution carries the factor $x$, giving $y=(c_1+c_2 x)e^{rx}$. For complex roots,
$\alpha$ is the \emph{real} part (in the exponential) and $\beta$ the
\emph{imaginary} part (inside the sine and cosine); swapping them is the classic
sign trap.
\end{warningbox}

\subsection{The Cauchy--Euler Equation}

The Cauchy--Euler (equidimensional) equation $ax^2y'' + bxy' + cy = 0$ has powers
of $x$ matching derivative orders. Try $y=x^m$.

\begin{formulabox}[Cauchy--Euler Root Cases]
Substituting $y=x^m$ gives the indicial equation $am(m-1)+bm+c=0$:
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Distinct} $m_1\neq m_2$: \quad $y = c_1 x^{m_1} + c_2 x^{m_2}$.
    \item \textbf{Repeated} $m$: \quad $y = (c_1 + c_2\ln x)\,x^{m}$.
    \item \textbf{Complex} $m=\alpha\pm\beta i$:
          \[ y = x^{\alpha}\big(c_1\cos(\beta\ln x) + c_2\sin(\beta\ln x)\big). \]
\end{itemize}
\end{formulabox}

\subsection{Higher-Order Linear Equations}

An $n$th-order constant-coefficient equation has a degree-$n$ characteristic
polynomial; each real root contributes an exponential, and a root of
\textbf{multiplicity} $k$ contributes $k$ independent terms.

\begin{formulabox}[Repeated Roots at Higher Order]
A root $r$ of multiplicity $k$ contributes
\[ e^{r x},\; x e^{r x},\; x^2 e^{r x},\; \ldots,\; x^{k-1} e^{r x}, \]
and a complex pair $\alpha\pm\beta i$ of multiplicity $k$ contributes the analogous
$x^j e^{\alpha x}\cos\beta x$ and $x^j e^{\alpha x}\sin\beta x$ for $j=0,\ldots,k-1$.
\end{formulabox}

\begin{conceptbox}[Unit 9 Key Takeaway]
The general solution of a homogeneous second-order linear equation is
$\textcolor{accentorange}{c_1 y_1 + c_2 y_2}$ for any
\textcolor{accentorange}{linearly independent} pair, which the
\textcolor{accentorange}{Wronskian} certifies and \textcolor{accentorange}{Abel's
identity} guards. For constant coefficients the
\textcolor{accentorange}{characteristic roots} dictate the form ---
\textcolor{accentorange}{real-distinct}, \textcolor{accentorange}{repeated}
(carry the $x$), or \textcolor{accentorange}{complex} ($e^{\alpha x}$ times sines
and cosines) --- and \textcolor{accentorange}{reduction of order} recovers a second
solution from a first.
\end{conceptbox}

% ===========================================================================
% PART II: REVIEW GUIDE (PRACTICE QUESTIONS)
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part II --- Review Guide: Practice Problems}
\section*{Part II --- Review Guide: Practice Problems}

Work the following ten problems in order. They climb through the three
characteristic-root cases --- real distinct, repeated, and complex conjugate ---
then exercise the Wronskian, Abel's identity, reduction of order, the
Cauchy--Euler equation, and a higher-order capstone. Solutions follow in
Part III.

\begin{enumerate}[leftmargin=2em]
    \item \textbf{(Real distinct roots)} Solve $y'' - 5y' + 6y = 0$. Write the
          characteristic equation, factor it, and give the general solution.

    \item \textbf{(Repeated root)} Solve $y'' - 6y' + 9y = 0$. Identify the
          double root and explain why the factor $x$ appears in the second
          solution.

    \item \textbf{(Complex conjugate roots)} Solve $y'' + 4y' + 13y = 0$.
          Report the complex roots $\alpha\pm\beta i$ and the real general
          solution.

    \item \textbf{(Real-root IVP)} Solve the initial value problem
          $y'' - y' - 2y = 0$, $y(0) = 1$, $y'(0) = 0$.

    \item \textbf{(Wronskian)} Compute the Wronskian of $y_1 = e^{2x}$ and
          $y_2 = e^{3x}$, and state what it certifies.

    \item \textbf{(Abel's identity)} For $y'' + 3y' + 2y = 0$, use Abel's formula
          to find the Wronskian up to a constant, then verify it directly from the
          fundamental solutions.

    \item \textbf{(Reduction of order)} The equation $x^2 y'' - 3x y' + 4y = 0$
          has the solution $y_1 = x^2$. Use reduction of order to find a second,
          independent solution $y_2$.

    \item \textbf{(Cauchy--Euler)} Solve $x^2 y'' + 2x y' - 6y = 0$ for $x>0$.

    \item \textbf{(Repeated-root IVP)} Solve $y'' - 4y' + 4y = 0$,
          $y(0) = 1$, $y'(0) = 1$.

    \item \textbf{(Capstone, higher order)} Solve $y''' - 3y'' + 3y' - y = 0$.
          Identify the multiplicity of the characteristic root and write the
          general solution.
\end{enumerate}

% ===========================================================================
% PART III: SOLUTIONS
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part III --- Complete Solutions}
\section*{Part III --- Complete Solutions}

\noindent\textbf{Solution 1.}\quad Try $y=e^{rx}$ to get the characteristic
equation, then factor:
\[ \textcolor{workpurple}{r^2 - 5r + 6 = (r-2)(r-3) = 0 \;\Rightarrow\; r_1 = 2,\; r_2 = 3.} \]
The roots are real and distinct, so
\[ \textcolor{workpurple}{y = c_1 e^{2x} + c_2 e^{3x}.} \]

\medskip
\noindent\textbf{Solution 2.}\quad The characteristic equation is a perfect
square:
\[ \textcolor{workpurple}{r^2 - 6r + 9 = (r-3)^2 = 0 \;\Rightarrow\; r = 3 \text{ (double)}.} \]
A double root supplies only one exponential $e^{3x}$; reduction of order furnishes
the second solution $x e^{3x}$, so
\[ \textcolor{workpurple}{y = (c_1 + c_2 x)\,e^{3x}.} \]

\medskip
\noindent\textbf{Solution 3.}\quad Complete the discriminant:
\[ \textcolor{workpurple}{r = \frac{-4 \pm \sqrt{16 - 52}}{2} = \frac{-4 \pm \sqrt{-36}}{2} = -2 \pm 3i.} \]
With $\alpha=-2$ and $\beta=3$,
\[ \textcolor{workpurple}{y = e^{-2x}\big(c_1\cos 3x + c_2\sin 3x\big).} \]

\medskip
\noindent\textbf{Solution 4.}\quad Characteristic equation
$r^2 - r - 2 = (r-2)(r+1)=0$ gives $r=2,-1$, so
$y = c_1 e^{2x} + c_2 e^{-x}$. Apply the initial conditions:
\[ \textcolor{workpurple}{y(0) = c_1 + c_2 = 1, \qquad y'(0) = 2c_1 - c_2 = 0.} \]
Adding the equations, $3c_1 = 1$, so
\[ \textcolor{workpurple}{c_1 = \tfrac{1}{3},\quad c_2 = \tfrac{2}{3}, \qquad
   y = \tfrac{1}{3} e^{2x} + \tfrac{2}{3} e^{-x}.} \]

\medskip
\noindent\textbf{Solution 5.}\quad With $y_1=e^{2x}$, $y_2=e^{3x}$:
\[ \textcolor{workpurple}{W = y_1 y_2' - y_2 y_1' = e^{2x}\cdot 3e^{3x} - e^{3x}\cdot 2e^{2x} = (3-2)e^{5x} = e^{5x}.} \]
Since $e^{5x}\neq 0$ for all $x$, the two solutions are \textbf{linearly
independent} and $\{e^{2x},e^{3x}\}$ is a fundamental set.

\medskip
\noindent\textbf{Solution 6.}\quad Here $p(x)=3$, so Abel's formula gives
\[ \textcolor{workpurple}{W(x) = C\,\exp\!\left(-\int 3\,dx\right) = C\,e^{-3x}.} \]
Check directly: the roots of $r^2+3r+2=(r+1)(r+2)=0$ are $r=-1,-2$, so
$y_1=e^{-x}$, $y_2=e^{-2x}$ and
\[ \textcolor{workpurple}{W = e^{-x}(-2e^{-2x}) - e^{-2x}(-e^{-x}) = (-2+1)e^{-3x} = -e^{-3x},} \]
which matches $C e^{-3x}$ with $C=-1$.

\medskip
\noindent\textbf{Solution 7.}\quad Normalize to $y'' - \tfrac{3}{x}y' + \tfrac{4}{x^2}y = 0$,
so $p(x) = -3/x$ and $e^{-\int p\,dx} = e^{\int (3/x)\,dx} = e^{3\ln x} = x^3$. With
$y_1=x^2$,
\[ \textcolor{workpurple}{y_2 = y_1\int \frac{e^{-\int p\,dx}}{y_1^2}\,dx
   = x^2\int \frac{x^3}{x^4}\,dx = x^2\int \frac{1}{x}\,dx = x^2\ln x.} \]
The second independent solution is $y_2 = x^2\ln x$ (the repeated-root
Cauchy--Euler form).

\medskip
\noindent\textbf{Solution 8.}\quad Try $y=x^m$ in $x^2y''+2xy'-6y=0$. The indicial
equation is
\[ \textcolor{workpurple}{m(m-1) + 2m - 6 = m^2 + m - 6 = (m+3)(m-2) = 0 \;\Rightarrow\; m = 2,\,-3.} \]
The roots are real and distinct, so for $x>0$
\[ \textcolor{workpurple}{y = c_1 x^2 + c_2 x^{-3}.} \]

\medskip
\noindent\textbf{Solution 9.}\quad The characteristic equation
$r^2 - 4r + 4 = (r-2)^2 = 0$ has the double root $r=2$, so
$y = (c_1 + c_2 x)e^{2x}$. Differentiate:
$y' = \big(c_2 + 2(c_1 + c_2 x)\big)e^{2x}$. Apply the conditions:
\[ \textcolor{workpurple}{y(0) = c_1 = 1, \qquad y'(0) = c_2 + 2c_1 = 1 \;\Rightarrow\; c_2 = -1.} \]
Therefore
\[ \textcolor{workpurple}{y = (1 - x)\,e^{2x}.} \]

\medskip
\noindent\textbf{Solution (Capstone).}\quad The characteristic equation is a
perfect cube:
\[ \textcolor{workpurple}{r^3 - 3r^2 + 3r - 1 = (r-1)^3 = 0 \;\Rightarrow\; r = 1 \text{ (multiplicity 3)}.} \]
A triple root contributes $e^{x}$, $x e^{x}$, and $x^2 e^{x}$, so
\[ \textcolor{workpurple}{y = (c_1 + c_2 x + c_3 x^2)\,e^{x}.} \]

\vspace{0.8em}
\begin{conceptbox}[Unit 9 Key Takeaway]
Every constant-coefficient solution is read straight off the
\textcolor{accentorange}{characteristic roots}: real-distinct roots give
\textcolor{accentorange}{$c_1 e^{r_1 x} + c_2 e^{r_2 x}$}, a repeated root carries
the factor $x$ as \textcolor{accentorange}{$(c_1 + c_2 x)e^{rx}$}, and a complex
pair gives \textcolor{accentorange}{$e^{\alpha x}(c_1\cos\beta x + c_2\sin\beta x)$}.
The \textcolor{accentorange}{Wronskian} confirms independence,
\textcolor{accentorange}{Abel's identity} predicts it, and
\textcolor{accentorange}{reduction of order} recovers the missing partner ---
exactly the $x^2\ln x$ that the repeated Cauchy--Euler root demands.
\end{conceptbox}

\end{document}
