% Unit 5: Numerical Methods
% Condensed Cheat Sheet (Part I only)
% Prepared by Staples Education
% Content informed by the local curriculum modules (5.1 to 5.6) and the
% QUIZ_DATA micro-practice items for Unit 5.
% Shared style harness pulled from the unified preamble input.

\input{app/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 5: Numerical Methods}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Condensed Cheat Sheet}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 5 at a Glance: Core Sub-Topics]
When an initial value problem $\frac{dy}{dt}=f(t,y),\ y(t_0)=y_0$ cannot be
solved in closed form, we march the solution forward in small steps of size $h$.
Every method in this unit is a recipe for turning the slope field $f$ into the
next point; they differ only in how many slopes they sample per step and how they
weight them.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Euler's Method} --- one tangent-line step
          $y_{n+1}=y_n+hf(t_n,y_n)$; first order, global error $O(h)$.
    \item \textbf{Improved and Modified Euler} --- two-slope averaging and the
          midpoint step; both second order, $O(h^{2})$.
    \item \textbf{Heun's Method} --- the predictor--corrector reading of improved
          Euler as a trapezoidal average of endpoint slopes.
    \item \textbf{The RK4 Method} --- four weighted stage slopes
          $\tfrac{h}{6}(k_1+2k_2+2k_3+k_4)$; fourth order, $O(h^{4})$.
    \item \textbf{The Runge--Kutta Family} --- the general weighted-stage template
          that contains Euler, Heun, and RK4 as members.
    \item \textbf{Adaptive Step Size and Comparison} --- estimating local error to
          resize $h$, and ranking the methods by order of accuracy.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{5}   % subsections auto-number as the unit modules 5.1, 5.2, ...

\subsection{Euler's Method}

Euler's method is the simplest numerical solver: from the current point it walks
a single step along the tangent line whose slope is $f(t_n,y_n)$. It is the
foundation every richer method refines.

\begin{formulabox}[The Euler Update]
\[ y_{n+1} = y_n + h\,f(t_n,y_n), \qquad t_{n+1} = t_n + h. \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item Each step follows the \textbf{tangent line} with slope $f(t_n,y_n)$.
    \item \textbf{Local} truncation error is $O(h^{2})$ (the dropped
          $\tfrac{h^{2}}{2}y''$ term); \textbf{global} error is $O(h)$, so Euler
          is a \textbf{first-order} method.
    \item Halving $h$ roughly \textbf{halves} the global error.
\end{itemize}
\end{formulabox}

\begin{examplebox}[Worked Example 5.1: One Euler step for $y'=y$]
With $y(0)=1$ and $h=0.1$, the slope at the start is $f(0,1)=1$:
\[ \textcolor{workpurple}{y_1 = 1 + 0.1\cdot 1 = 1.1, \qquad t_1 = 0.1.} \]
The exact value is $e^{0.1}\approx 1.10517$, so a single first-order step is
already low by about $0.005$.
\end{examplebox}

\subsection{Improved and Modified Euler Methods}

Both upgrades sample a \emph{second} slope to capture the curvature that plain
Euler ignores, lifting the order from one to two. The improved method averages
the slopes at the two endpoints; the modified (midpoint) method reads one slope
at the interval's center.

\begin{formulabox}[Improved Euler (Endpoint Average) and Modified Euler (Midpoint)]
\textbf{Improved Euler:}
\[ k_1 = f(t_n,y_n), \qquad k_2 = f\!\left(t_n+h,\; y_n+h k_1\right), \]
\[ y_{n+1} = y_n + \frac{h}{2}\left(k_1 + k_2\right). \]
\textbf{Modified Euler (midpoint):}
\[ k_1 = f(t_n,y_n), \qquad k_2 = f\!\left(t_n+\tfrac{h}{2},\; y_n+\tfrac{h}{2}k_1\right), \]
\[ y_{n+1} = y_n + h\,k_2. \]
Both are \textbf{second order}, with global error $O(h^{2})$: halving $h$ cuts the
error by about a factor of four.
\end{formulabox}

\begin{examplebox}[Worked Example 5.2: Improved Euler for $y'=y$]
With $y(0)=1$, $h=0.1$: the predictor slope is $k_1=1$ and the corrected slope is
$k_2=f(0.1,\,1.1)=1.1$. Averaging,
\[ \textcolor{workpurple}{y_1 = 1 + \frac{0.1}{2}\left(1 + 1.1\right) = 1 + 0.05\cdot 2.1 = 1.105.} \]
This already matches $e^{0.1}\approx 1.10517$ to three decimals --- far better than
the plain Euler $1.1$.
\end{examplebox}

\subsection{Heun's Method}

Heun's method is improved Euler told as a \emph{predictor--corrector}: take a
provisional Euler step to predict the endpoint, then correct with the
\emph{average} of the slopes at the two ends. That average is exactly the
trapezoidal rule applied to $\int f\,dt$.

\begin{methodbox}[Heun's Predictor--Corrector]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item \textbf{Predict} the endpoint with a plain Euler step:
          \[ \tilde{y}_{n+1} = y_n + h\,f(t_n,y_n). \]
    \item \textbf{Correct} with the trapezoidal average of the two slopes:
          \[ y_{n+1} = y_n + \frac{h}{2}\Big[\,f(t_n,y_n) + f(t_{n+1},\tilde{y}_{n+1})\,\Big]. \]
\end{enumerate}
Averaging the two endpoint slopes recovers the curvature term plain Euler drops,
making Heun a \textbf{second-order} method --- identical to improved Euler.
\end{methodbox}

\begin{warningbox}[Common Pitfall: Unequal vs.\ Equal Slope Weights]
The improved Euler / Heun corrector weights the two slopes \emph{equally},
$\tfrac{h}{2}(k_1+k_2)$. Do not borrow RK4's $\tfrac{h}{6}(k_1+2k_2+2k_3+k_4)$
weights for a two-slope method, and do not forget the predictor uses the
\emph{full} step $h$ while the midpoint method uses a \emph{half} step
$\tfrac{h}{2}$.
\end{warningbox}

\subsection{The RK4 Method}

The classical fourth-order Runge--Kutta method samples four slopes per step --- at
the start, twice at the midpoint, and once at the end --- and blends them with
Simpson-like weights $1,2,2,1$. It is the workhorse of practical ODE solving.

\begin{formulabox}[The Classical RK4 Step]
\[ k_1 = f(t_n,\,y_n), \]
\[ k_2 = f\!\left(t_n+\tfrac{h}{2},\; y_n+\tfrac{h}{2}k_1\right), \qquad
   k_3 = f\!\left(t_n+\tfrac{h}{2},\; y_n+\tfrac{h}{2}k_2\right), \]
\[ k_4 = f\!\left(t_n+h,\; y_n+h k_3\right), \]
\[ y_{n+1} = y_n + \frac{h}{6}\left(k_1 + 2k_2 + 2k_3 + k_4\right). \]
RK4 is \textbf{fourth order}, with global error $O(h^{4})$: halving $h$ shrinks the
error by about a factor of $\mathbf{16}$.
\end{formulabox}

\begin{examplebox}[Worked Example 5.4: One RK4 step for $y'=y$]
With $y(0)=1$, $h=0.1$:
\[ \textcolor{workpurple}{k_1 = 1,\quad k_2 = 1.05,\quad k_3 = 1.0525,\quad k_4 = 1.10525.} \]
\[ \textcolor{workpurple}{y_1 = 1 + \frac{0.1}{6}\big(1 + 2(1.05) + 2(1.0525) + 1.10525\big) = 1.105171.} \]
Against $e^{0.1}=1.1051709\ldots$ the error is below $10^{-6}$ in a single step.
\end{examplebox}

\subsection{The Runge--Kutta Family}

Euler, the midpoint method, Heun, and RK4 are all instances of one template: take
a weighted average of slopes sampled at chosen points across the step. The
\emph{number of stages} and the \emph{weights} fix the order of accuracy.

\begin{formulabox}[The General Explicit Runge--Kutta Step]
For $s$ stages with weights $b_i$, nodes $c_i$, and coupling coefficients
$a_{ij}$:
\[ k_i = f\!\left(t_n + c_i h,\; y_n + h\sum_{j<i} a_{ij}\,k_j\right), \qquad
   y_{n+1} = y_n + h\sum_{i=1}^{s} b_i\,k_i. \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item Consistency requires $\sum_i b_i = 1$ (the weights sum to one).
    \item Euler is the $1$-stage member; improved Euler, the midpoint method, and
          Heun are $2$-stage second-order members; RK4 is the classic $4$-stage
          fourth-order member.
\end{itemize}
\end{formulabox}

\subsection{Adaptive Step Size and Method Comparison}

Rather than fix $h$ in advance, adaptive solvers \emph{estimate} the local error
each step and resize $h$ to hold it under a tolerance --- small steps through
sharp turns, large steps across calm stretches.

\begin{methodbox}[Estimating Local Error to Resize $h$]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item Advance the step two ways of \emph{different order} (e.g.\ an embedded
          RK4/RK5 pair, or one full step vs.\ two half steps).
    \item The difference of the two estimates approximates the \textbf{local
          error}.
    \item If the error exceeds the tolerance, \textbf{shrink} $h$ and retry; if it
          is comfortably under, \textbf{grow} $h$ for the next step.
\end{enumerate}
\end{methodbox}

\begin{warningbox}[Common Pitfall: Confusing Order with Step Count]
A method's \emph{order} $p$ is how fast error falls as $h\to 0$ (global error
$O(h^{p})$), not how many steps it takes. RK4 with a large $h$ can beat Euler with
a tiny $h$ at equal cost. And smaller $h$ is not free: too small, and accumulated
roundoff begins to swamp the truncation gain.
\end{warningbox}

\begin{conceptbox}[Unit 5 Key Takeaway]
Every solver here marches $y_{n+1}=y_n+h\,(\text{weighted slope})$; they differ
only in the slope. \textcolor{accentorange}{Euler} samples one slope (first order,
$O(h)$); \textcolor{accentorange}{improved/modified Euler and Heun} average two
(second order, $O(h^{2})$); \textcolor{accentorange}{RK4} blends four with weights
$\textcolor{accentorange}{1,2,2,1}$ (fourth order, $O(h^{4})$). All sit inside the
\textcolor{accentorange}{Runge--Kutta family}, and \textcolor{accentorange}{adaptive
stepping} tunes $h$ from an on-the-fly local-error estimate.
\end{conceptbox}

\end{document}
