% Unit 7: First-Order Exactness and Methods
% Condensed Cheat Sheet (Part I only)
% Prepared by Staples Education
% Content informed by the local curriculum modules (7.1 to 7.7) read from
% curriculum-data.js and the 10-problem master practice set for Unit 7.
% Shared style harness pulled from the unified preamble input.

\input{app/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 7: First-Order Exactness and Methods}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Condensed Cheat Sheet}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box (all 7 curriculum modules from curriculum-data.js)
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 7 at a Glance: Core Sub-Topics]
A first-order ODE written as $M\,dx + N\,dy = 0$ is \emph{exact} when its
left side is already the total differential $dF$ of some scalar potential $F$,
making the general solution simply $F(x,y)=C$. This unit develops the complete
toolkit: test the cross-partials to detect exactness, reconstruct the potential
via the integration loop, repair non-exact equations with special integrating
factors, connect the theory to conservative vector fields, and read the physics
through thermodynamic state functions.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{7.1 Exact Differential Equations} --- the criterion $M_y = N_x$
          and the implicit general solution $F(x,y)=C$.
    \item \textbf{7.2 Exact Equations Intuition} --- Clairaut's theorem:
          $F_{xy}=F_{yx}$ is why $M_y=N_x$ guarantees a potential.
    \item \textbf{7.3 Exact Equations Motivation and Theory} --- the integration
          loop: $F = \int M\,dx + g(y)$, fix $g'(y)$ by matching $F_y = N$.
    \item \textbf{7.4 Non-Exact Equations and Special Integrating Factors} ---
          integrating factor $\mu(x)$ when $\Delta/N$ is pure in $x$; $\mu(y)$
          when $-\Delta/M$ is pure in $y$ (where $\Delta = M_y - N_x$).
    \item \textbf{7.5 Almost Exact Differential Equations} --- decision tree:
          test both $\mu(x)$ and $\mu(y)$ routes systematically, then execute
          the repair and re-verify exactness before reconstructing $F$.
    \item \textbf{7.6 Exact Equations and Conservative Vector Fields} ---
          $M_y=N_x$ iff $\mathbf F=\langle M,N\rangle$ is conservative;
          path independence via the Fundamental Theorem for Line Integrals.
    \item \textbf{7.7 Applications in Thermodynamics} --- $dU$ is an exact
          state differential; $\delta Q$ is inexact; the integrating factor
          $1/T$ produces exact entropy $dS = \delta Q/T$.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{7}   % subsections auto-number as 7.1, 7.2, ...

% ---------------------------------------------------------------------------
\subsection{Exact Differential Equations}
% ---------------------------------------------------------------------------

A first-order ODE $M(x,y)\,dx + N(x,y)\,dy = 0$ is \emph{exact} on a simply
connected region when its left side equals the total differential $dF$ of some
potential function $F(x,y)$. The key theorem reduces exactness to a single
algebraic test on the cross-partial derivatives of $M$ and $N$.

\begin{formulabox}[Exactness Criterion and Implicit Solution]
The ODE $M\,dx + N\,dy = 0$ (with $M,N$ having continuous first partials) is
\textbf{exact} on a simply connected region if and only if
\[
  \frac{\partial M}{\partial y} = \frac{\partial N}{\partial x}.
\]
When exact, a potential $F(x,y)$ exists satisfying $F_x = M$ and $F_y = N$,
and the \textbf{general solution} is
\[
  F(x,y) = C.
\]
\end{formulabox}

\begin{warningbox}[Common Pitfall: Swapped Cross Partials]
Always differentiate $M$ with respect to \textbf{$y$} and $N$ with respect to
\textbf{$x$}. The test is $M_y = N_x$, never $M_x$ versus $N_y$. Computing the
same-variable partials tells you nothing about exactness and will produce a
false positive or false negative on every equation you test.
\end{warningbox}

% ---------------------------------------------------------------------------
\subsection{Exact Equations Intuition}
% ---------------------------------------------------------------------------

The cross-partial criterion is not an arbitrary rule --- it is Clairaut's
theorem in disguise. If $M = F_x$ and $N = F_y$ for some $F$, then
$M_y = F_{xy}$ and $N_x = F_{yx}$, and Clairaut guarantees the two mixed
partials agree whenever $F$ has continuous second partials. Exactness is
precisely the statement that $M$ and $N$ are partials of a common function.

\begin{formulabox}[Clairaut's Theorem and the Exact-Differential Connection]
On a simply connected region with $M,N$ continuously differentiable,
\[
  M\,dx + N\,dy = dF
  \iff
  \frac{\partial M}{\partial y}
  = \frac{\partial^{2}F}{\partial y\,\partial x}
  = \frac{\partial^{2}F}{\partial x\,\partial y}
  = \frac{\partial N}{\partial x}.
\]
The equality of mixed partials is both \textbf{necessary} (if $dF$ exists then
the cross partials must agree) and \textbf{sufficient} (on a simply connected
domain, equal cross partials guarantee a global potential).
\end{formulabox}

% ---------------------------------------------------------------------------
\subsection{Exact Equations Motivation and Theory}
% ---------------------------------------------------------------------------

Once exactness is confirmed, the potential $F$ is found by a two-step
integration loop. Integrating $M$ with respect to $x$ introduces an arbitrary
function $g(y)$ in place of a constant; that function is then pinned down by
the condition $F_y = N$.

\begin{methodbox}[Potential Function Reconstruction: The Integration Loop]
Given $M\,dx + N\,dy = 0$ confirmed exact by $M_y = N_x$:
\begin{enumerate}[leftmargin=1.6em, itemsep=3pt]
    \item \textbf{Integrate $M$ in $x$}, holding $y$ fixed:
          \[
            F(x,y) = \int M(x,y)\,dx + g(y).
          \]
    \item \textbf{Differentiate $F$ in $y$} and set the result equal to $N$:
          \[
            \frac{\partial F}{\partial y} = N(x,y)
            \quad\Longrightarrow\quad
            g'(y) = N(x,y) - \frac{\partial}{\partial y}\int M\,dx.
          \]
          (All $x$-terms must cancel; if they do not, re-check the exactness
          test.)
    \item \textbf{Integrate $g'(y)$} in $y$ to obtain $g(y)$.
    \item Write the \textbf{general solution} $F(x,y) = C$.
\end{enumerate}
\end{methodbox}

\begin{warningbox}[Common Pitfall: $g'(y)$ Must Be Free of $x$]
After forming $g'(y) = N - \partial_y\!\int M\,dx$, every $x$-term must
cancel. If $x$ survives in $g'(y)$, the equation was not exact (or an
arithmetic error occurred). Integrating an $x$-dependent expression as if it
were a function of $y$ alone produces a wrong answer without any LaTeX or
algebra error to signal the problem.
\end{warningbox}

% ---------------------------------------------------------------------------
\subsection{Non-Exact Equations and Special Integrating Factors}
% ---------------------------------------------------------------------------

When $M_y \neq N_x$, no potential $F$ exists for the equation as written.
Multiplying both sides by a non-zero \emph{integrating factor} $\mu$ can
change the cross partials of $\mu M$ and $\mu N$ until they match. Two
special forms of $\mu$ are algebraically testable and cover the majority of
non-exact equations encountered in practice.

\begin{formulabox}[Integrating Factors $\mu(x)$ and $\mu(y)$]
Let $\Delta = M_y - N_x$ (the \textbf{exactness defect}).
\begin{itemize}[leftmargin=1.4em, itemsep=4pt]
    \item \textbf{Test for $\mu(x)$:} Compute $\dfrac{\Delta}{N}$.
          If the result depends on $x$ \emph{only}, set $h(x)=\Delta/N$ and
          \[
            \mu(x) = \exp\!\left(\int h(x)\,dx\right).
          \]
    \item \textbf{Test for $\mu(y)$:} Compute $\dfrac{N_x - M_y}{M}
          = \dfrac{-\Delta}{M}$.
          If the result depends on $y$ \emph{only}, set $k(y)=-\Delta/M$ and
          \[
            \mu(y) = \exp\!\left(\int k(y)\,dy\right).
          \]
\end{itemize}
After multiplying the ODE through by $\mu$, solve the resulting exact equation
via the integration loop (Section~7.3).
\end{formulabox}

\begin{warningbox}[Common Pitfall: Always Re-Verify Exactness After Multiplying]
The formulas for $\mu(x)$ and $\mu(y)$ are derived by assuming the repaired
equation will be exact. After multiplying by $\mu$, explicitly re-compute the
cross partials $(\mu M)_y$ and $(\mu N)_x$ and confirm they are equal before
proceeding. A mis-simplified defect ratio can give a wrong $\mu$ that does
\emph{not} restore exactness.
\end{warningbox}

% ---------------------------------------------------------------------------
\subsection{Almost Exact Differential Equations}
% ---------------------------------------------------------------------------

An \emph{almost exact} equation is non-exact but possesses an integrating
factor of one of the two standard forms ($\mu(x)$ or $\mu(y)$). The key skill
is recognizing which test succeeds and executing the full repair-and-solve
workflow systematically.

\begin{methodbox}[Integrating-Factor Decision Tree]
Given $M\,dx + N\,dy = 0$ with $M_y \neq N_x$:
\begin{enumerate}[leftmargin=1.6em, itemsep=3pt]
    \item Compute the defect $\Delta = M_y - N_x$.
    \item \textbf{Try $\mu(x)$:} Form $\Delta/N$. Does all $y$ cancel?
          If yes, $\mu(x) = e^{\int \Delta/N\,dx}$.
    \item \textbf{Try $\mu(y)$:} Form $-\Delta/M$. Does all $x$ cancel?
          If yes, $\mu(y) = e^{\int(-\Delta/M)\,dy}$.
    \item Multiply the ODE by $\mu$, re-verify $(\mu M)_y = (\mu N)_x$, then
          use the integration loop to find $F$ and write $F = C$.
\end{enumerate}
\end{methodbox}

\begin{formulabox}[Why the Integrating Factor ODE Works]
Requiring $(\mu M)_y = (\mu N)_x$ for $\mu = \mu(x)$ yields
\[
  \mu\,M_y = \mu'\,N + \mu\,N_x
  \quad\Longrightarrow\quad
  \frac{\mu'}{\mu} = \frac{M_y - N_x}{N} = \frac{\Delta}{N},
\]
a separable ODE in $\mu(x)$ whose solution is $\mu = e^{\int \Delta/N\,dx}$.
The analogous calculation with $\mu = \mu(y)$ gives $\mu'/\mu = -\Delta/M$.
\end{formulabox}

% ---------------------------------------------------------------------------
\subsection{Exact Equations and Conservative Vector Fields}
% ---------------------------------------------------------------------------

Every exact equation $M\,dx + N\,dy = 0$ can be read as the zero-work
condition $\mathbf{F}\cdot d\mathbf{r}=0$ for the vector field
$\mathbf{F}=\langle M,N\rangle$. Exactness is equivalent to $\mathbf{F}$
being \emph{conservative}, and the potential $F$ of the ODE is the scalar
potential $\varphi$ with $\nabla\varphi=\mathbf{F}$.

\begin{formulabox}[Conservative Fields, Scalar Potentials, and Path Independence]
A $C^1$ vector field $\mathbf{F}=\langle M,N\rangle$ on a simply connected
region is \textbf{conservative} if and only if
\[
  \frac{\partial M}{\partial y} = \frac{\partial N}{\partial x}.
\]
When conservative, there exists a scalar potential $\varphi$ with
$\nabla\varphi = \mathbf{F}$, found by the same integration loop as the ODE
potential. Line integrals are then \textbf{path independent}:
\[
  \int_C \mathbf{F}\cdot d\mathbf{r}
  = \varphi\!\left(B\right) - \varphi\!\left(A\right)
\]
for any smooth curve $C$ from point $A$ to point $B$ (Fundamental Theorem for
Line Integrals). Equivalently, $\oint_C \mathbf{F}\cdot d\mathbf{r}=0$ for
every closed curve.
\end{formulabox}

\begin{warningbox}[Common Pitfall: The Domain Must Be Simply Connected]
The equivalence $M_y = N_x \iff$ conservative holds only on a
\textbf{simply connected} domain (a domain with no holes). On a
multiply-connected domain such as $\mathbb{R}^2\setminus\{(0,0)\}$, the field
$\mathbf{F}=\langle -y,\,x\rangle/(x^2+y^2)$ satisfies $M_y=N_x$ everywhere
it is defined yet is \emph{not} conservative: its integral around the unit
circle equals $2\pi\neq 0$.
\end{warningbox}

\vspace{0.6em}

\begin{conceptbox}[Unit 7 Key Takeaway]
Every exact equation $M\,dx+N\,dy=0$ is secretly $dF=0$; its general solution
is \textcolor{accentorange}{$F(x,y)=C$} recovered by the integration loop
\textcolor{accentorange}{$F=\int M\,dx+g(y)$} with $g'(y)$ fixed by matching
$F_y=N$. When $M_y\neq N_x$, form the defect \textcolor{accentorange}{$\Delta=M_y-N_x$}
and repair with \textcolor{accentorange}{$\mu(x)=e^{\int\Delta/N\,dx}$} if
$\Delta/N$ is pure in $x$, or \textcolor{accentorange}{$\mu(y)=e^{-\int\Delta/M\,dy}$}
if $-\Delta/M$ is pure in $y$. Exactness is conservatism:
\textcolor{accentorange}{$M_y=N_x\iff\mathbf{F}=\nabla\varphi$}, giving
path independence $\int_C\mathbf{F}\cdot d\mathbf{r}=\varphi(B)-\varphi(A)$,
and the integrating factor $1/T$ converts inexact thermodynamic heat
$\delta Q$ into the exact entropy differential $dS$.
\end{conceptbox}

\end{document}
