% Unit 6: Multivariable Calculus Foundations
% Condensed Cheat Sheet (Part I only)
% Prepared by Staples Education
% Content informed by the local curriculum modules (6.1 to 6.10) and the
% QUIZ_DATA micro-practice items for Unit 6.
% Shared style harness pulled from the unified preamble input.

\input{app/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 6: Multivariable Calculus Foundations}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Condensed Cheat Sheet}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 6 at a Glance: Core Sub-Topics]
Differential equations in more than one variable need the language of
multivariable calculus. This unit builds that bridge: we learn to differentiate
one variable at a time, to package those partials into a gradient, to chain
through intermediate variables, to locate and classify extrema, and to read the
geometry of level curves --- the same geometry that decides when a differential
$M\,dx+N\,dy$ is \emph{exact}.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Partial Derivatives} --- differentiate in one variable while
          holding the others constant; $f_x,\ f_y$.
    \item \textbf{Higher-Order and Mixed Partials} --- Clairaut's theorem:
          $f_{xy}=f_{yx}$ when the second partials are continuous.
    \item \textbf{Gradient and Directional Derivatives} --- $\nabla f$ collects
          the partials; $D_{\mathbf u}f=\nabla f\cdot\mathbf u$ measures change
          along a unit direction.
    \item \textbf{The Multivariable Chain Rule} --- sum one product term per path
          from the output down to the independent variable.
    \item \textbf{Critical Points and Optimization} --- set $\nabla f=\mathbf 0$,
          then classify with the second-derivative (Hessian) test.
    \item \textbf{Level Curves and Exactness} --- $f(x,y)=c$, $\nabla f$ normal to
          the curve, and the exact-differential test $\partial_y M=\partial_x N$.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{6}   % subsections auto-number as the unit modules 6.1, 6.2, ...

\subsection{Partial Derivatives}

A partial derivative differentiates a multivariable function in \emph{one}
variable while treating every other variable as a constant symbol. It measures
the instantaneous rate of change of $f$ along a single coordinate axis.

\begin{formulabox}[Definition of the Partial Derivatives]
\[ f_x = \frac{\partial f}{\partial x} = \lim_{h\to 0}\frac{f(x+h,y)-f(x,y)}{h},
   \qquad
   f_y = \frac{\partial f}{\partial y} = \lim_{h\to 0}\frac{f(x,y+h)-f(x,y)}{h}. \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item To compute $f_x$, \textbf{hold $y$ constant} and differentiate in $x$
          with the ordinary rules; any pure-$y$ term is a constant and dies.
    \item Holding a variable \emph{constant} is not the same as
          \emph{setting it to zero}: the symbol stays in the answer.
\end{itemize}
\end{formulabox}

\begin{examplebox}[Worked Example 6.1: First partials of $f=x^2y+y^3$]
Treat $y$ as constant for $f_x$, then $x$ as constant for $f_y$:
\[ \textcolor{workpurple}{f_x = 2xy + 0 = 2xy, \qquad f_y = x^2 + 3y^2.} \]
The $y^3$ term vanishes under $\partial_x$, while the $x^2$ term contributes its
coefficient $x^2$ under $\partial_y$.
\end{examplebox}

\subsection{Higher-Order and Mixed Partials}

Each first partial can itself be differentiated again, giving four second-order
partials: two \emph{pure} ($f_{xx},f_{yy}$) and two \emph{mixed}
($f_{xy},f_{yx}$). Clairaut's theorem says the mixed pair agrees whenever the
function is smooth enough.

\begin{formulabox}[Clairaut's Theorem on Mixed Partials]
If the second partial derivatives of $f$ are \textbf{continuous} near a point,
then the mixed partials are equal there:
\[ f_{xy} = \frac{\partial^2 f}{\partial y\,\partial x}
          = \frac{\partial^2 f}{\partial x\,\partial y} = f_{yx}. \]
The order of differentiation does not matter --- a function of two variables has
\textbf{four} second partials but only \textbf{three} distinct values.
\end{formulabox}

\begin{examplebox}[Worked Example 6.2: Mixed partials of $f=x^2y^3$]
Differentiate in $x$ first, then in $y$ --- and the reverse:
\[ \textcolor{workpurple}{f_x = 2xy^3 \;\Rightarrow\; f_{xy} = 6xy^2,
   \qquad f_y = 3x^2y^2 \;\Rightarrow\; f_{yx} = 6xy^2.} \]
Both routes land on $6xy^2$, exactly as Clairaut guarantees.
\end{examplebox}

\subsection{The Gradient and Directional Derivatives}

The gradient packages the partials into a single vector that points in the
direction of steepest increase. Dotting it with a \emph{unit} direction gives
the rate of change along that direction --- the directional derivative.

\begin{formulabox}[Gradient, Directional Derivative, and Steepest Ascent]
\[ \nabla f = \left(\frac{\partial f}{\partial x},\ \frac{\partial f}{\partial y}\right),
   \qquad
   D_{\mathbf u} f = \nabla f \cdot \mathbf u \quad(\mathbf u \text{ a unit vector}). \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item The \textbf{maximum} rate of increase at a point is $\lvert\nabla f\rvert$,
          attained in the direction $\mathbf u = \nabla f/\lvert\nabla f\rvert$.
    \item $\nabla f$ is \textbf{perpendicular} to the level curve through the point.
\end{itemize}
\end{formulabox}

\begin{warningbox}[Common Pitfall: The Direction Must Be a Unit Vector]
$D_{\mathbf u}f=\nabla f\cdot\mathbf u$ is only valid when $\mathbf u$ is
\textbf{normalized}. Given a raw direction $\mathbf v$, divide first:
$\mathbf u=\mathbf v/\lvert\mathbf v\rvert$. Skipping the normalization rescales
the answer by $\lvert\mathbf v\rvert$ and silently corrupts every directional
derivative and max-rate computation that follows.
\end{warningbox}

\begin{examplebox}[Worked Example 6.3: Directional derivative of $f=x^2+xy+y^2$]
At $(1,2)$, $\nabla f=(2x+y,\,x+2y)=(4,5)$. Along $\mathbf v=(3,4)$, normalize to
$\mathbf u=(\tfrac35,\tfrac45)$:
\[ \textcolor{workpurple}{D_{\mathbf u}f = (4,5)\cdot\left(\tfrac35,\tfrac45\right)
   = \tfrac{12}{5}+\tfrac{20}{5} = \tfrac{32}{5} = 6.4.} \]
The steepest ascent there points along $(4,5)$ with rate
$\lvert\nabla f\rvert=\sqrt{41}\approx 6.40$.
\end{examplebox}

\subsection{The Multivariable Chain Rule}

When the inputs of $f$ are themselves functions, differentiate by summing one
product term for every path from the output down to the variable you are
differentiating with respect to. A small \emph{tree diagram} bookkeeps the paths.

\begin{methodbox}[Chain Rule Trees: Counting the Paths]
\textbf{One independent variable} $z=f(x(t),y(t))$ --- two paths to $t$:
\[ \frac{dz}{dt} = \frac{\partial f}{\partial x}\frac{dx}{dt}
                 + \frac{\partial f}{\partial y}\frac{dy}{dt}
                 = \nabla f \cdot (x',y'). \]
\textbf{Two independent variables} $z=f(x(s,t),y(s,t))$ --- two paths to each:
\[ \frac{\partial z}{\partial s} = f_x\,x_s + f_y\,y_s,
   \qquad
   \frac{\partial z}{\partial t} = f_x\,x_t + f_y\,y_t. \]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item Draw a branch from $z$ to each intermediate variable ($x,y$).
    \item Extend a branch from each intermediate variable to the target ($t$).
    \item \textbf{Multiply} along each branch, then \textbf{add} the branches.
\end{enumerate}
\end{methodbox}

\begin{examplebox}[Worked Example 6.4: Chain rule for $z=x^2y$, $x=t^2$, $y=t^3$]
With $z_x=2xy$, $z_y=x^2$, $x'=2t$, $y'=3t^2$:
\[ \textcolor{workpurple}{\frac{dz}{dt} = (2xy)(2t) + (x^2)(3t^2)
   = (2t^5)(2t) + (t^4)(3t^2) = 4t^6 + 3t^6 = 7t^6.} \]
The direct substitution $z=t^4\cdot t^3=t^7$ confirms $dz/dt=7t^6$.
\end{examplebox}

\subsection{Critical Points and Optimization}

Interior extrema of a smooth $f(x,y)$ occur where the gradient vanishes. The
\emph{second-derivative test} then uses the sign of the Hessian discriminant to
sort each critical point into a minimum, maximum, or saddle.

\begin{methodbox}[Locating and Classifying Critical Points]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item \textbf{Solve} $\nabla f=\mathbf 0$, i.e.\ $f_x=0$ and $f_y=0$
          simultaneously, to find all critical points.
    \item \textbf{Form} the discriminant
          $D = f_{xx}f_{yy} - \left(f_{xy}\right)^2$ at each point.
    \item \textbf{Classify} with $D$ and $f_{xx}$ (Step in the box below).
\end{enumerate}
\end{methodbox}

\begin{formulabox}[The Second-Derivative (Hessian) Test]
\[ D = f_{xx}\,f_{yy} - \left(f_{xy}\right)^2. \]
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item $D>0,\ f_{xx}>0$: \textbf{local minimum}.
    \item $D>0,\ f_{xx}<0$: \textbf{local maximum}.
    \item $D<0$: \textbf{saddle point} (no extremum).
    \item $D=0$: the test is \textbf{inconclusive}.
\end{itemize}
\end{formulabox}

\begin{warningbox}[Common Pitfall: Reading the Sign Boundary Wrong]
The sign of $f_{xx}$ only decides min vs.\ max \emph{after} $D>0$ confirms an
extremum. When $D<0$ the point is a \textbf{saddle} regardless of $f_{xx}$, and
when $D=0$ the test says \emph{nothing} --- do not declare a saddle by default.
A single mishandled sign flips a minimum into a maximum.
\end{warningbox}

\subsection{Level Curves, Gradients, and Exactness}

Level curves $f(x,y)=c$ slice a surface into contour lines; the gradient cuts
across them at right angles. This same gradient structure is exactly what makes a
first-order differential \emph{exact} --- the payoff that carries into Unit 7.

\begin{formulabox}[Level Curves and the Exact-Differential Test]
A \textbf{level curve} is the set $\{(x,y):f(x,y)=c\}$; $\nabla f$ is normal to
it. A differential $M(x,y)\,dx+N(x,y)\,dy$ is \textbf{exact} on a simply
connected region exactly when
\[ \frac{\partial M}{\partial y} = \frac{\partial N}{\partial x}, \]
in which case it equals $df$ for a potential $f$ with $f_x=M,\ f_y=N$ --- and the
field $\mathbf F=(M,N)=\nabla f$ is \textbf{conservative}.
\end{formulabox}

\begin{warningbox}[Common Pitfall: Path Dependence and the Wrong Cross Partials]
The exactness/conservative test compares the \emph{cross} partials
$\partial_y M$ and $\partial_x N$ --- not $\partial_x M$ and $\partial_y N$. When
they differ the field is path-dependent: $\int_C\mathbf F\cdot d\mathbf r$
depends on the route, so there is \emph{no} single potential to read off.
\end{warningbox}

\begin{conceptbox}[Unit 6 Key Takeaway]
Multivariable calculus is the toolkit the rest of the course runs on. The
\textcolor{accentorange}{partials} $f_x,f_y$ collect into the
\textcolor{accentorange}{gradient $\nabla f$}, which both gives directional
derivatives $\textcolor{accentorange}{D_{\mathbf u}f=\nabla f\cdot\mathbf u}$ and
points along steepest ascent. The \textcolor{accentorange}{chain rule} sums one
term per path; \textcolor{accentorange}{optimization} sets $\nabla f=\mathbf 0$
and classifies with $\textcolor{accentorange}{D=f_{xx}f_{yy}-f_{xy}^{\,2}}$; and
the exactness test $\textcolor{accentorange}{\partial_y M=\partial_x N}$ is the
gradient idea in disguise, the gateway to exact equations.
\end{conceptbox}

\end{document}
