% Unit 8: Special Equations and Classical Models
% Master Unit Reference Guide (Cheat Sheet, Problems, and Solutions)
% Prepared by Staples Education
% Content informed by the local curriculum modules (8.1 to 8.7) read from
% curriculum-data.js and the 10-problem master practice set for Unit 8.
% Shared style harness pulled from the unified preamble input.

\input{app/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 8: Special Equations and Classical Models}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Master Unit Reference Guide: Cheat Sheet, Problems, and Solutions}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Table of contents (master guide only)
% ---------------------------------------------------------------------------
\tableofcontents

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box (all 7 curriculum modules from curriculum-data.js)
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 8 at a Glance: Core Sub-Topics]
Unit 8 moves from the algebraic theory of first-order ODEs into the classical
applied models that drive physics, biology, and economics. It opens with two
\emph{special equation types}---the Clairaut equation (singular-solution
envelopes) and pursuit/tractrix curves---before building three major applied
models: the Navier--Stokes-derived plane Poiseuille velocity profile, the
supply-and-demand price-adjustment ODE, and the Lotka--Volterra predator-prey
and competition systems. The unit closes with a structured survey of first-order
solution methods.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{8.1 The Clairaut Equation and Singular Solutions} ---
          general solution $y = Cx + f(C)$; singular (envelope) solution from
          the system $\{y = xp + f(p),\; 0 = x + f'(p)\}$.
    \item \textbf{8.2 Pursuit Curves and the Tractrix} ---
          tractrix DE $dy/dx = -\sqrt{a^2 - x^2}/x$; general pursuit
          $x\,y'' = k\sqrt{1+(y')^2}$ with speed ratio $k = v_m/v_d$.
    \item \textbf{8.3 Fluid Dynamics and Viscosity} ---
          Newtonian shear $\tau = \mu\,du/dy$; plane Poiseuille flow
          $\mu u'' = -G$ with no-slip boundary conditions.
    \item \textbf{8.4 Economic Models, Supply, Demand, and Price} ---
          price ODE $dp/dt = k(D - S)$; equilibrium $p^* = (a+c)/(b+d)$;
          convergence for $k(b+d) > 0$.
    \item \textbf{8.5 The Lotka--Volterra Predator-Prey Model} ---
          coexistence equilibrium $(c/d,\, a/b)$; conserved quantity
          $H = dx - c\ln x + by - a\ln y$.
    \item \textbf{8.6 Competition Models} ---
          two-species carrying-capacity system; nullcline inequalities for
          stable coexistence versus competitive exclusion.
    \item \textbf{8.7 A Survey of Solution Methods} ---
          decision tree: separable, linear (integrating factor), Bernoulli,
          exact; classify by structure before solving.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{8}   % subsections auto-number as 8.1, 8.2, ...

% ---------------------------------------------------------------------------
\subsection{The Clairaut Equation and Singular Solutions}
% ---------------------------------------------------------------------------

A \emph{Clairaut equation} has the special form $y = xy' + f(y')$.
Setting $p = y'$ and differentiating both sides with respect to $x$ yields
$(x + f'(p))\,dp/dx = 0$, splitting into two branches whose solutions are
the general family of straight lines and the singular envelope.

\begin{formulabox}[Clairaut Equation: General and Singular Solutions]
For $y = xp + f(p)$ with $p = y'$:
\[
  \textbf{General solution:}\quad y = Cx + f(C) \quad (p = C = \text{const}),
\]
\[
  \textbf{Singular solution (envelope):}\quad
  \begin{cases}
    y = xp + f(p) \\
    0 = x + f'(p)
  \end{cases}
\]
Eliminate $p$ from this system to obtain the envelope: a curve tangent to every
line in the general family.
The singular solution satisfies the ODE but \emph{cannot} be obtained from the
general solution by any choice of $C$.
\end{formulabox}

\begin{methodbox}[Envelope Elimination Procedure]
\begin{enumerate}[leftmargin=1.6em, itemsep=3pt]
  \item Write the equation as $F(x,y,C) = y - Cx - f(C) = 0$.
  \item Differentiate with respect to $C$: $\partial F/\partial C = -x - f'(C) = 0$,
        giving $x = -f'(C)$.
  \item Solve for $C = C(x)$ and substitute back into $F = 0$ to obtain the
        envelope curve $y = y(x)$.
  \item Verify: the envelope must satisfy the original ODE directly.
\end{enumerate}
\end{methodbox}

\begin{warningbox}[Common Pitfall: Singular Solution Is Not a Particular Solution]
The singular solution is \emph{not} a member of the general family $y = Cx + f(C)$
for any finite $C$. Derive it exclusively via the envelope elimination system;
attempting to recover it by substituting a value of $C$ will always fail.
\end{warningbox}

% ---------------------------------------------------------------------------
\subsection{Pursuit Curves and the Tractrix}
% ---------------------------------------------------------------------------

A \emph{pursuit curve} traces the path of a pursuer who always aims directly at
a moving target. The tractrix is the classical limiting case in which the
pursuer is dragged at constant distance $a$ from a target walking along a line.

\begin{formulabox}[Tractrix Differential Equation and Solution]
Target on the $y$-axis, pursuer starts at $(a,0)$, equal speeds, string length $a$:
\[
  \frac{dy}{dx} = -\frac{\sqrt{a^2 - x^2}}{x}, \qquad 0 < x \le a,\quad y(a) = 0.
\]
Solution (via $x = a\sin\theta$):
\[
  y(x) = a\ln\!\left(\frac{a + \sqrt{a^2 - x^2}}{x}\right) - \sqrt{a^2 - x^2}.
\]
\end{formulabox}

\begin{formulabox}[General Pursuit Equation]
Dog at origin pursues owner on line $x = c$ at speed $v_m$; dog speed $v_d > v_m$;
$k = v_m/v_d < 1$:
\[
  x\,y'' = k\sqrt{1+(y')^2}, \qquad y(c) = 0,\quad y'(c) = 0.
\]
Solution ($k \ne 1$):
\[
  y(x) = \frac{1}{2}\!\left[\frac{x^{k+1}}{(k+1)c^k}
  - \frac{c^k x^{1-k}}{1-k}\right] + \frac{ck}{1-k^2}.
\]
Capture height: $y(0) = ck/(1-k^2)$.
\end{formulabox}

\begin{warningbox}[Common Pitfall: Direction of the Derivative Sign]
In the tractrix, $dy/dx$ is \emph{negative} because $x$ decreases as the pursuer
is dragged toward the $y$-axis. Always verify the geometric sign from the
tangent-segment picture before writing the ODE.
\end{warningbox}

% ---------------------------------------------------------------------------
\subsection{Fluid Dynamics and Viscosity}
% ---------------------------------------------------------------------------

Newtonian viscosity: $\tau = \mu\,du/dy$. Plane Poiseuille flow reduces the
momentum balance to a second-order constant-coefficient ODE with a parabolic solution.

\begin{formulabox}[Plane Poiseuille Velocity Profile]
$\mu\,u''(y) = -G$, $u(0) = u(h) = 0$:
\[
  u(y) = \frac{G}{2\mu}\,y(h - y), \qquad
  y_{\max} = \frac{h}{2}, \qquad
  u_{\max} = \frac{Gh^2}{8\mu}.
\]
Shear stress: $\tau(y) = \tfrac{G}{2}(h - 2y)$, linear, zero at the centerline.
\end{formulabox}

\begin{formulabox}[2D Incompressible Continuity and Stream Function]
$\partial u/\partial x + \partial v/\partial y = 0$.
Given $u$: integrate $\partial v/\partial y = -\partial u/\partial x$ in $y$.
Stream function: $u = \partial\psi/\partial y$, $v = -\partial\psi/\partial x$;
streamlines are $\psi(x,y) = C$.
\end{formulabox}

\begin{warningbox}[Common Pitfall: Both No-Slip Conditions Are Required]
The Poiseuille ODE produces a quadratic with two free constants. Both
$u(0) = 0$ and $u(h) = 0$ must be applied; using only one gives a wrong family.
\end{warningbox}

% ---------------------------------------------------------------------------
\subsection{Economic Models, Supply, Demand, and Price}
% ---------------------------------------------------------------------------

Price adjusts at a rate proportional to excess demand. With linear demand and
supply the ODE is first-order linear, and the equilibrium price is globally stable.

\begin{formulabox}[Price Adjustment ODE and Equilibrium]
$D = a - bp$, $S = -c + dp$ ($a,b,c,d > 0$):
\[
  \frac{dp}{dt} = k\bigl[(a+c) - (b+d)p\bigr],
  \qquad
  p(t) = p^* + (p_0 - p^*)\,e^{-k(b+d)t},
  \qquad
  p^* = \frac{a+c}{b+d}.
\]
Convergence: $k(b+d) > 0$ always.
\end{formulabox}

\begin{methodbox}[Solving the Price ODE via Integrating Factor]
\begin{enumerate}[leftmargin=1.6em, itemsep=3pt]
  \item Write $dp/dt + k(b+d)p = k(a+c)$.
  \item $\mu(t) = e^{k(b+d)t}$; multiply and integrate.
  \item Apply $p(0) = p_0$: $C = p_0 - p^*$.
\end{enumerate}
\end{methodbox}

% ---------------------------------------------------------------------------
\subsection{The Lotka--Volterra Predator-Prey Model}
% ---------------------------------------------------------------------------

The Lotka--Volterra system produces closed phase-plane orbits guaranteed by
the first integral $H(x,y)$.

\begin{formulabox}[Lotka--Volterra System and Conserved Quantity]
$\dot{x} = ax - bxy$, $\dot{y} = -cy + dxy$ ($a,b,c,d > 0$):
\[
  E_0 = (0,0), \qquad E^* = (c/d,\; a/b).
\]
\[
  H(x,y) = dx - c\ln x + by - a\ln y = \text{const} \quad
  (\text{minimum at } E^*).
\]
\end{formulabox}

\begin{methodbox}[Deriving the Conserved Quantity]
\begin{enumerate}[leftmargin=1.6em, itemsep=3pt]
  \item Form $dy/dx = \dot{y}/\dot{x} = y(-c + dx)/[x(a - by)]$.
  \item Separate: $(a/y - b)\,dy = (-c/x + d)\,dx$.
  \item Integrate: $a\ln y - by = -c\ln x + dx + \text{const}$.
  \item Rearrange: $H = dx - c\ln x + by - a\ln y$.
  \item Verify $\nabla H = 0$ at $E^*$ and Hessian positive definite.
\end{enumerate}
\end{methodbox}

\begin{warningbox}[Common Pitfall: The Coexistence Equilibrium Is a Center]
$(c/d, a/b)$ is a \emph{center} (closed orbits), not a stable spiral.
Perturbations do not damp out in the pure Lotka--Volterra model; adding
saturation converts the center to a stable spiral.
\end{warningbox}

% ---------------------------------------------------------------------------
\subsection{Competition Models}
% ---------------------------------------------------------------------------

Two-species competition: the outcome depends on the relative positions of the
$x$- and $y$-nullclines.

\begin{formulabox}[Competition System and Interior Equilibrium]
$\dot{x} = r_1 x(1 - (x + \alpha y)/K_1)$,\;
$\dot{y} = r_2 y(1 - (y + \beta x)/K_2)$.
\[
  E_3 = (x^*, y^*), \quad
  x^* = \frac{K_1 - \alpha K_2}{1 - \alpha\beta}, \quad
  y^* = \frac{K_2 - \beta K_1}{1 - \alpha\beta}.
\]
Nullclines: $x + \alpha y = K_1$ and $\beta x + y = K_2$.
\end{formulabox}

\begin{methodbox}[Coexistence vs.\ Exclusion: Nullcline Intercept Test]
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
  \item \textbf{Stable coexistence} ($\alpha\beta < 1$):
        $K_1 < K_2/\beta$ and $K_2 < K_1/\alpha$.
  \item \textbf{Bistable exclusion} ($\alpha\beta > 1$):
        $K_1 > K_2/\beta$ and $K_2 > K_1/\alpha$.
  \item \textbf{$x$ wins}: $K_1 > K_2/\beta$ and $K_2 < K_1/\alpha$.
  \item \textbf{$y$ wins}: $K_1 < K_2/\beta$ and $K_2 > K_1/\alpha$.
\end{itemize}
\end{methodbox}

\begin{warningbox}[Common Pitfall: Interior Equilibrium Feasibility]
$E_3$ is biologically meaningful only when both $x^* > 0$ and $y^* > 0$.
Check the signs of numerators and denominator before drawing conclusions.
\end{warningbox}

\vspace{0.6em}

\begin{conceptbox}[Unit 8 Key Takeaway --- Part I]
Clairaut equations split into a \textcolor{accentorange}{general line family
$y = Cx + f(C)$} and a \textcolor{accentorange}{singular envelope} from
$x + f'(p) = 0$. The Poiseuille profile
\textcolor{accentorange}{$u = (G/2\mu)\,y(h-y)$} and price path
\textcolor{accentorange}{$p(t) = p^* + (p_0 - p^*)e^{-k(b+d)t}$} follow from
linear BVPs. Lotka--Volterra orbits are level sets of
\textcolor{accentorange}{$H = dx - c\ln x + by - a\ln y$}; competition outcomes
follow from \textcolor{accentorange}{$\alpha\beta < 1$} (coexistence) versus
\textcolor{accentorange}{$\alpha\beta > 1$} (exclusion).
\end{conceptbox}

% ===========================================================================
% PART II: REVIEW GUIDE (PRACTICE QUESTIONS)
% Master guide: \clearpage before Part II so it begins on a fresh page.
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part II --- Review Guide: Practice Problems}
\section*{Part II --- Review Guide: Practice Problems}

Work the following ten problems in order. Problems 1--2 develop the Clairaut
equation: general straight-line family and singular envelope elimination
(Module 8.1). Problems 3--4 derive and solve the tractrix ODE and the general
pursuit equation with a speed-ratio parameter (Module 8.2). Problem 5 solves
the plane Poiseuille velocity profile with no-slip boundary conditions, and
Problem 6 applies 2D incompressible continuity to find the stream function
(Module 8.3). Problem 7 solves the linear price-adjustment ODE and identifies
the equilibrium (Module 8.4). Problem 8 analyzes the full Lotka--Volterra
predator-prey system including the conserved quantity $H$ (Module 8.5).
Problem 9 classifies coexistence versus competitive exclusion in a two-species
competition model via nullcline inequalities (Module 8.6). Problem 10 closes
with a four-part classification and solution survey (Module 8.7).

\begin{enumerate}[leftmargin=2em, itemsep=8pt]

    \item Find the general solution and the singular solution of the Clairaut
          equation $y = xy' + (y')^2$. Verify that the singular solution is
          the envelope of the general straight-line family.

    \item For the Clairaut equation $y = xy' - \ln(y')$ (with $y' > 0$),
          find the one-parameter family of straight-line solutions and the
          singular solution. Confirm that the singular solution satisfies the
          ODE directly.

    \item A target moves from the origin up the positive $y$-axis. A pursuer
          starts at $(a,0)$ and is always connected to the target by a taut
          string of length $a$, with both moving at the same speed. Derive the
          tractrix differential equation
          \[
            \frac{dy}{dx} = -\frac{\sqrt{a^2 - x^2}}{x},
            \qquad y(a) = 0,
          \]
          from the constant-length tangent-line geometry, then solve it for
          $y(x)$ using the substitution $x = a\sin\theta$.

    \item A dog at the origin always runs directly toward its owner, who walks
          along $x = c$ at speed $v_m$. The dog runs at speed $v_d > v_m$ with
          $k = v_m/v_d$. Derive $x\,y'' = k\sqrt{1+(y')^2}$, $y(c) = 0$,
          $y'(c) = 0$. Set $p = y'$, solve the separable equation in $p$,
          integrate to find $y(x)$, and state the finite capture height $y(0)$.

    \item Solve $\mu\,u''(y) = -G$, $u(0) = u(h) = 0$ for the velocity profile
          $u(y)$. State the location and value of the maximum velocity.

    \item Given $u = x^2 - y^2$ for a 2D incompressible steady flow, find the
          most general $v(x,y)$ consistent with $\partial u/\partial x +
          \partial v/\partial y = 0$. For $g(x) = 0$, determine the stream
          function $\psi(x,y)$ and write the streamline equation.

    \item Price obeys $dp/dt = k(D - S)$ with $D = a - bp$, $S = -c + dp$
          ($a,b,c,d,k > 0$). Write the ODE, solve via integrating factor with
          $p(0) = p_0$, find $p^*$, and state the convergence condition.

    \item For $\dot{x} = ax - bxy$, $\dot{y} = -cy + dxy$ ($a,b,c,d > 0$):
          interpret each parameter; find both equilibria; derive the conserved
          quantity $H(x,y) = dx - c\ln x + by - a\ln y$; verify $H$ is
          minimized at $(c/d, a/b)$.

    \item For the competition system with parameters $r_1, r_2, K_1, K_2,
          \alpha, \beta$: find all four equilibria; write the nullclines;
          state the inequalities governing stable coexistence, bistable
          exclusion, and single-species dominance; derive $(x^*, y^*)$ explicitly.

    \item Classify and solve completely:
          (a) $dy/dx = x(1+y^2)/(1+x^2)$;\;
          (b) $dy/dx + (2/x)y = x^3$;\;
          (c) $dy/dx + y = xy^3$;\;
          (d) $(2xy + \cos x)\,dx + (x^2 - \sin y)\,dy = 0$.

\end{enumerate}

\vspace{0.4em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART III: COMPLETE SOLUTIONS
% Master guide: \clearpage before Part III so solutions begin on a fresh page.
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part III --- Complete Solutions}
\section*{Part III --- Complete Solutions}

% -------------------------------------------------------------------
\noindent\textbf{Solution 1.}\quad
Set $p = y'$. The Clairaut equation is $y = xp + p^2$.

Differentiate both sides with respect to $x$:
\[
  \textcolor{workpurple}{p = p + x\,\frac{dp}{dx} + 2p\,\frac{dp}{dx}
  = p + (x + 2p)\,\frac{dp}{dx}.}
\]
Cancel $p$:
\[
  \textcolor{workpurple}{(x + 2p)\,\frac{dp}{dx} = 0.}
\]

\textit{Case A:} $dp/dx = 0 \Rightarrow p = C$.
\textbf{General solution:} $y = Cx + C^2$.

\textit{Case B:} $x + 2p = 0 \Rightarrow p = -x/2$. Substitute:
\[
  \textcolor{workpurple}{y = x\!\left(-\frac{x}{2}\right)
  + \left(-\frac{x}{2}\right)^2
  = -\frac{x^2}{2} + \frac{x^2}{4} = -\frac{x^2}{4}.}
\]

Envelope check: $F = y - Cx - C^2$; $\partial F/\partial C = -x - 2C = 0$,
so $C = -x/2$. Substituting:
\[
  \textcolor{workpurple}{y + \frac{x^2}{2} - \frac{x^2}{4}
  = y + \frac{x^2}{4} = 0 \;\Longrightarrow\; y = -\frac{x^2}{4}.}
\]
Direct check: $y' = -x/2$, $(y')^2 = x^2/4$,
$xy' + (y')^2 = -x^2/2 + x^2/4 = -x^2/4 = y$. \checkmark

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 2.}\quad
Set $p = y' > 0$. The equation is $y = xp - \ln p$.

Differentiate:
\[
  \textcolor{workpurple}{p = p + \!\left(x - \frac{1}{p}\right)\frac{dp}{dx}
  \quad\Longrightarrow\quad
  \left(x - \frac{1}{p}\right)\frac{dp}{dx} = 0.}
\]

\textit{Case A:} $p = C$.
\textbf{General solution:} $y = Cx - \ln C$.

\textit{Case B:} $p = 1/x$. Substitute:
\[
  \textcolor{workpurple}{y = x\cdot\frac{1}{x} - \ln\!\frac{1}{x} = 1 + \ln x.}
\]
\textbf{Singular solution:} $y = 1 + \ln x$.

Envelope: $F = y - Cx + \ln C$; $\partial F/\partial C = -x + 1/C = 0$,
so $C = 1/x$. Then $y - 1 - \ln x = 0$. \checkmark

Direct check: $y' = 1/x$; $xy' - \ln y' = 1 + \ln x = y$. \checkmark

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 3.}\quad
At $(x,y)$ the string has horizontal leg $x$ and hypotenuse $a$, so vertical
leg $= \sqrt{a^2 - x^2}$. Since $x$ decreases as the pursuer is dragged:
\[
  \textcolor{workpurple}{\frac{dy}{dx} = -\frac{\sqrt{a^2 - x^2}}{x}.}
\]

Substitute $x = a\sin\theta$, $dx = a\cos\theta\,d\theta$,
$\sqrt{a^2 - x^2} = a\cos\theta$:
\[
  \textcolor{workpurple}{\int\frac{\sqrt{a^2-x^2}}{x}\,dx
  = a\int\frac{\cos^2\theta}{\sin\theta}\,d\theta
  = a\int(\csc\theta - \sin\theta)\,d\theta
  = a\bigl[\ln|\csc\theta - \cot\theta| + \cos\theta\bigr].}
\]
Back-substitute ($\csc\theta = a/x$, $\cot\theta = \sqrt{a^2-x^2}/x$,
$\cos\theta = \sqrt{a^2-x^2}/a$):
\[
  \textcolor{workpurple}{= a\ln\!\left(\frac{a - \sqrt{a^2-x^2}}{x}\right)
  + \sqrt{a^2-x^2}.}
\]
So $y = -a\ln\!\bigl((a-\sqrt{a^2-x^2})/x\bigr) - \sqrt{a^2-x^2} + K$.
Applying $y(a) = 0$ gives $K = 0$. Rationalizing:
\[
  \textcolor{workpurple}{y(x) = a\ln\!\left(\frac{a + \sqrt{a^2-x^2}}{x}\right)
  - \sqrt{a^2-x^2}.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 4.}\quad
Tangency condition at dog position $(x,y)$ with owner at $(c,Y)$:
$Y - y = (c-x)y'$. Differentiate in $x$: $dY/dx = (c-x)y''$.
Speed ratio: $dY/dx = k\sqrt{1+(y')^2}$.
Equating: $(c-x)y'' = k\sqrt{1+(y')^2}$, i.e.\ $xy'' = k\sqrt{1+(y')^2}$
(after shifting).

Let $p = y'$; separate $dp/\sqrt{1+p^2} = k\,dx/x$:
\[
  \textcolor{workpurple}{\operatorname{arcsinh}(p) = k\ln x + C_1
  \;\Longrightarrow\;
  p + \sqrt{1+p^2} = A\,x^k.}
\]
$y'(c) = 0$: $A = c^{-k}$. Using $(\sqrt{1+p^2}+p)(\sqrt{1+p^2}-p) = 1$:
\[
  \textcolor{workpurple}{2p = (x/c)^k - (x/c)^{-k}
  \;\Longrightarrow\;
  y' = \frac{1}{2}\!\left[(x/c)^k - (x/c)^{-k}\right].}
\]
Integrate ($k \ne 1$):
\[
  \textcolor{workpurple}{y = \frac{1}{2}\!\left[\frac{x^{k+1}}{(k+1)c^k}
  - \frac{c^k x^{1-k}}{1-k}\right] + C_2.}
\]
$y(c) = 0$: $C_2 = ck/(1-k^2)$.

\textbf{Solution:}
$y(x) = \tfrac{1}{2}\bigl[x^{k+1}/((k+1)c^k) - c^k x^{1-k}/(1-k)\bigr] + ck/(1-k^2)$.
Capture height: $\textcolor{workpurple}{y(0) = ck/(1-k^2)}$.

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 5.}\quad
Integrate $\mu u'' = -G$ twice:
\[
  \textcolor{workpurple}{\mu u'(y) = -Gy + C_1,
  \qquad
  \mu u(y) = -\frac{G}{2}y^2 + C_1 y + C_2.}
\]
$u(0) = 0 \Rightarrow C_2 = 0$.\quad
$u(h) = 0 \Rightarrow C_1 = Gh/2$.
\[
  \textcolor{workpurple}{u(y) = \frac{G}{2\mu}\,y(h-y).}
\]
$u'(y) = (G/2\mu)(h-2y) = 0$ at $y = h/2$:
\[
  \textcolor{workpurple}{u_{\max} = \frac{G}{2\mu}\cdot\frac{h}{2}\cdot\frac{h}{2}
  = \frac{Gh^2}{8\mu}.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 6.}\quad
$\partial u/\partial x = 2x$, so $\partial v/\partial y = -2x$.
Integrate in $y$:
\[
  \textcolor{workpurple}{v(x,y) = -2xy + g(x).}
\]
With $g = 0$, find $\psi$ from $\partial\psi/\partial y = x^2 - y^2$:
\[
  \textcolor{workpurple}{\psi = x^2 y - \frac{y^3}{3} + h(x).}
\]
Check $-\partial\psi/\partial x = -2xy - h'(x) = v = -2xy$,
so $h' = 0$, $h = \text{const}$. Streamlines:
\[
  \textcolor{workpurple}{x^2 y - \frac{y^3}{3} = k.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 7.}\quad
\[
  \textcolor{workpurple}{\frac{dp}{dt} = k\bigl[(a+c)-(b+d)p\bigr]
  \;\Longrightarrow\;
  \frac{dp}{dt} + k(b+d)p = k(a+c).}
\]
$\mu = e^{k(b+d)t}$; integrate:
\[
  \textcolor{workpurple}{p(t) = \frac{a+c}{b+d}
  + \!\left(p_0 - \frac{a+c}{b+d}\right)e^{-k(b+d)t}.}
\]
$p^* = (a+c)/(b+d)$. Convergence: $k(b+d) > 0$ (automatic). \checkmark

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 8.}\quad
\textit{Parameters:} $a$ = prey growth rate; $b$ = predation coefficient;
$c$ = predator death rate; $d$ = conversion efficiency.

\textit{Equilibria:} $\dot{x} = 0 \Rightarrow x = 0$ or $y = a/b$;
$\dot{y} = 0 \Rightarrow y = 0$ or $x = c/d$:
\[
  \textcolor{workpurple}{E_0 = (0,0), \qquad E^* = (c/d,\; a/b).}
\]

\textit{Conserved quantity.} $dy/dx = y(-c+dx)/[x(a-by)]$; separate:
\[
  \textcolor{workpurple}{\left(\frac{a}{y}-b\right)dy = \left(-\frac{c}{x}+d\right)dx.}
\]
Integrate:
\[
  \textcolor{workpurple}{a\ln y - by = -c\ln x + dx + \text{const}
  \;\Longrightarrow\;
  H = dx - c\ln x + by - a\ln y = \text{const}.}
\]
$\partial H/\partial x = d - c/x = 0 \Rightarrow x = c/d$;
$\partial H/\partial y = b - a/y = 0 \Rightarrow y = a/b$.
Hessian: $\partial^2 H/\partial x^2 = c/x^2 > 0$,
$\partial^2 H/\partial y^2 = a/y^2 > 0$, cross term $= 0$: positive
definite. $E^*$ is the unique minimum of $H$.

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 9.}\quad
\textit{Equilibria:}
$(0,0)$, $(K_1, 0)$, $(0, K_2)$, and
\[
  \textcolor{workpurple}{(x^*,y^*) = \!\left(\frac{K_1 - \alpha K_2}{1-\alpha\beta},\;
  \frac{K_2 - \beta K_1}{1-\alpha\beta}\right).}
\]
Derivation: $x + \alpha y = K_1$ and $\beta x + y = K_2$; substitute
$x = K_1 - \alpha y$ into the second:
\[
  \textcolor{workpurple}{\beta K_1 + y(1-\alpha\beta) = K_2
  \;\Longrightarrow\;
  y^* = \frac{K_2 - \beta K_1}{1-\alpha\beta},
  \quad
  x^* = \frac{K_1 - \alpha K_2}{1-\alpha\beta}.}
\]

\textit{Nullclines:}
$x + \alpha y = K_1$ (axes intercepts $(K_1,0)$ and $(0,K_1/\alpha)$);
$\beta x + y = K_2$ (axes intercepts $(K_2/\beta,0)$ and $(0,K_2)$).

\textit{Outcome inequalities:}
\[
  \textcolor{workpurple}{\text{Coexistence}:\;K_1 < K_2/\beta\;\text{and}\;K_2 < K_1/\alpha
  \;\;(\alpha\beta < 1).}
\]
\[
  \textcolor{workpurple}{\text{Bistable exclusion}:\;K_1 > K_2/\beta\;\text{and}\;K_2 > K_1/\alpha
  \;\;(\alpha\beta > 1).}
\]
\[
  \textcolor{workpurple}{x\text{ wins}:\;K_1 > K_2/\beta\;\text{and}\;K_2 < K_1/\alpha.
  \qquad
  y\text{ wins}:\;K_1 < K_2/\beta\;\text{and}\;K_2 > K_1/\alpha.}
\]

\medskip

% -------------------------------------------------------------------
\noindent\textbf{Solution 10.}\quad

\noindent\textit{(a) Separable.}
\[
  \textcolor{workpurple}{\frac{dy}{1+y^2} = \frac{x\,dx}{1+x^2}
  \;\Longrightarrow\;
  \arctan y = \frac{1}{2}\ln(1+x^2) + C
  \;\Longrightarrow\;
  y = \tan\!\left(\tfrac{1}{2}\ln(1+x^2)+C\right).}
\]

\noindent\textit{(b) Linear, integrating factor.}
$\mu = x^2$; $d(x^2 y)/dx = x^5$; integrate:
\[
  \textcolor{workpurple}{x^2 y = \frac{x^6}{6} + C
  \;\Longrightarrow\;
  y = \frac{x^4}{6} + \frac{C}{x^2}.}
\]

\noindent\textit{(c) Bernoulli, $n = 3$.}
Let $v = y^{-2}$: $v' - 2v = -2x$. $\mu = e^{-2x}$:
\[
  \textcolor{workpurple}{\frac{d}{dx}(v\,e^{-2x}) = -2x\,e^{-2x}.}
\]
Integrate by parts: $\int -2x e^{-2x}\,dx = xe^{-2x} + e^{-2x}/2$:
\[
  \textcolor{workpurple}{v = x + \tfrac{1}{2} + Ce^{2x}
  \;\Longrightarrow\;
  y^2 = \frac{1}{x + \tfrac{1}{2} + Ce^{2x}}.}
\]

\noindent\textit{(d) Exact.}
$M_y = 2x = N_x$. Integrate $M$ in $x$:
\[
  \textcolor{workpurple}{F = x^2 y + \sin x + g(y);\quad
  F_y = x^2 + g'(y) = x^2 - \sin y
  \;\Longrightarrow\; g(y) = \cos y.}
\]
\[
  \textcolor{workpurple}{x^2 y + \sin x + \cos y = C.}
\]

\vspace{0.8em}

\begin{conceptbox}[Unit 8 Key Takeaway]
Every Clairaut equation $y = xy' + f(y')$ yields a
\textcolor{accentorange}{general line family $y = Cx + f(C)$} and a
\textcolor{accentorange}{singular envelope} obtained by eliminating $p$ from
$x + f'(p) = 0$. Pursuit geometry converts tangency and speed constraints
into separable second-order ODEs. The Poiseuille profile
\textcolor{accentorange}{$u = (G/2\mu)\,y(h-y)$} and the price path
\textcolor{accentorange}{$p(t) = p^* + (p_0 - p^*)e^{-k(b+d)t}$} both follow
from linear BVPs. In Lotka--Volterra dynamics the
\textcolor{accentorange}{conserved quantity $H = dx - c\ln x + by - a\ln y$}
organizes all phase-plane orbits as closed curves around the coexistence
equilibrium $(c/d, a/b)$; in two-species competition the entire outcome is
decided by \textcolor{accentorange}{$\alpha\beta < 1$ (stable coexistence) versus
$\alpha\beta > 1$ (competitive exclusion)}.
\end{conceptbox}

\end{document}
