% Unit 3: Existence, Uniqueness, and Geometry
% Condensed Cheat Sheet (Part I only)
% Prepared by Staples Education
% Content informed by the local curriculum modules (3.1 to 3.6) and the
% QUIZ_DATA micro-practice items for Unit 3.
% Shared style harness pulled from the unified preamble input.

\input{app/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 3: Existence, Uniqueness, and Geometry}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Condensed Cheat Sheet}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 3 at a Glance: Core Sub-Topics]
This unit is the geometric and theoretical heart of first-order theory. It
trades formula-hunting for understanding: how to read an equation as a picture,
and why a solution should exist and be the only one.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Slope Fields and Vector Fields} --- the equation
          $y' = f(x,y)$ as a field of tiny tangent segments, traced by integral
          curves.
    \item \textbf{The Phase Plane and Phase Portraits} --- states, trajectories,
          equilibria, and nullclines for a system of equations.
    \item \textbf{Existence and Uniqueness} --- the two continuity hypotheses
          that guarantee exactly one solution through a point.
    \item \textbf{Fixed Points and Iteration} --- solving $g(x)=x$ by repeated
          application, and when the iteration is attracting.
    \item \textbf{The Banach Fixed Point Theorem} --- a contraction on a
          complete space has one fixed point, reached from any start.
    \item \textbf{Picard Iteration} --- recasting an IVP as an integral
          equation and converging to its solution.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{3}   % subsections auto-number as the unit modules 3.1, 3.2, ...

\subsection{Slope Fields and Vector Fields}

A first order equation $\frac{dy}{dx} = f(x,y)$ hands you a \emph{number} at
every point of the plane, and a derivative is a slope. So the equation is
secretly a picture: at each point plant a short segment whose slope is
$f(x,y)$. A \emph{solution} is any curve that stays tangent to those segments
everywhere it goes.

\begin{formulabox}[Slope Field, Integral Curves, and Isoclines]
At each point $(x,y)$ the field draws a segment of slope $f(x,y)$.
\begin{itemize}[leftmargin=1.4em, itemsep=2pt]
    \item An \textbf{integral curve} is a solution that is tangent to the field
          at every point it crosses.
    \item An \textbf{isocline} is the curve $f(x,y) = c$ of all points sharing
          one slope $c$; the isocline $f(x,y)=0$ marks the flat segments.
\end{itemize}
\end{formulabox}

\begin{examplebox}[Worked Example 3.1: The field of $\frac{dy}{dx}=x+y$]
\textbf{Step 1.} Read off a slope by substituting coordinates. At $(1,2)$:
\[ f(1,2) = 1 + 2 = 3. \]
\textbf{Step 2.} Find the flat isocline by setting $f=0$:
\[ x + y = 0 \;\Longrightarrow\; y = -x. \]
Along $y=-x$ every segment is horizontal; integral curves cross it with zero
slope.
\end{examplebox}

\subsection{The Phase Plane and Phase Portraits}

For a \emph{system} $x' = f(x,y)$, $y' = g(x,y)$, the plane no longer carries
time on an axis --- each point is a complete \emph{state}, and as $t$ runs the
state point sweeps out a \emph{trajectory}. A \emph{phase portrait} is a
representative bundle of trajectories that reveals the global flow.

\begin{formulabox}[Equilibria, Nullclines, and Linear Classification]
An \textbf{equilibrium} satisfies $x'=0$ and $y'=0$ simultaneously. The
$x$-nullcline is $\{x'=0\}$ and the $y$-nullcline is $\{y'=0\}$. For a linear
system the eigenvalues of the coefficient matrix classify the origin:
\[
\begin{array}{ll}
\text{real, same sign} & \text{node (source if } +,\ \text{sink if } -)\\
\text{real, opposite signs} & \text{saddle}\\
\text{complex, nonzero real part} & \text{spiral}\\
\text{purely imaginary} & \text{center (closed orbits)}
\end{array}
\]
\end{formulabox}

\begin{examplebox}[Worked Example 3.2: Classify $x'=y,\ y'=-x$]
\textbf{Step 1.} Set both right sides to zero for the equilibrium:
\[ y = 0 \text{ and } -x = 0 \;\Longrightarrow\; (0,0). \]
\textbf{Step 2.} The coefficient matrix
$\left(\begin{smallmatrix}0&1\\-1&0\end{smallmatrix}\right)$
has eigenvalues $\pm i$ --- purely imaginary.
\[ \text{Purely imaginary} \;\Longrightarrow\; \text{a center: closed orbits around the origin.} \]
\end{examplebox}

\subsection{Existence and Uniqueness of Solutions}

Before solving, ask whether a solution even exists and whether it is the
\emph{only} one. The Big Theorem answers both with two continuity hypotheses ---
one for each conclusion.

\begin{formulabox}[The Existence and Uniqueness Theorem]
For $\dfrac{dy}{dx} = f(x,y)$, $\;y(x_0)=y_0$: if $f$ \textbf{and}
$\dfrac{\partial f}{\partial y}$ are continuous on a rectangle containing
$(x_0,y_0)$, then there is a \textbf{unique} solution on \emph{some} open
interval around $x_0$. Continuity of $f$ alone delivers \emph{existence};
continuity of $\dfrac{\partial f}{\partial y}$ secures \emph{uniqueness}.
\end{formulabox}

\begin{warningbox}[Common Pitfall: Existence Is Not Uniqueness]
A continuous $f$ with a \emph{discontinuous} $\frac{\partial f}{\partial y}$
still has a solution --- just maybe more than one. For $y' = \sqrt{y}$,
$y(0)=0$, the derivative $\frac{1}{2\sqrt{y}}$ blows up at $y=0$, and indeed
both $y=0$ and $y=\frac{x^2}{4}$ pass through the origin. The guarantee is also
only \emph{local}: never read it as a promise for all $x$.
\end{warningbox}

\begin{examplebox}[Worked Example 3.3: Where uniqueness fails for $\frac{dy}{dx}=\frac{x}{y-2}$]
\textbf{Step 1.} Identify where $f$ and $\frac{\partial f}{\partial y}$ lose
continuity --- watch the denominator:
\[ f(x,y) = \frac{x}{y-2} \text{ is undefined when } y - 2 = 0. \]
\textbf{Step 2.} Conclude:
\[ \text{The theorem fails along the line } y = 2. \]
Off that line both $f$ and $\frac{\partial f}{\partial y} = -\frac{x}{(y-2)^2}$
are continuous, so a unique solution threads each such point.
\end{examplebox}

\subsection{Fixed Points and Iteration}

A \emph{fixed point} of $g$ is a value left unmoved by the map: $g(x^*)=x^*$.
Geometrically it is where $y=g(x)$ meets the line $y=x$. Feeding outputs back as
inputs --- $x_{n+1}=g(x_n)$ --- either pulls toward such a point or pushes away.

\begin{formulabox}[Fixed Point and Stability Test]
\[ x^* \text{ is a fixed point} \iff g(x^*) = x^*. \]
The iteration $x_{n+1} = g(x_n)$ is \textbf{attracting} (stable) at $x^*$ when
\[ |g'(x^*)| < 1, \qquad \text{and \textbf{repelling} when } |g'(x^*)| > 1. \]
\end{formulabox}

\begin{examplebox}[Worked Example 3.4: Fixed points of $g(x)=x^{2}$]
\textbf{Step 1.} Solve $g(x)=x$:
\[ x^2 = x \;\Longrightarrow\; x(x-1)=0 \;\Longrightarrow\; x=0 \text{ or } x=1. \]
\textbf{Step 2.} Test stability with $|g'(x)| = |2x|$:
\[ |g'(0)| = 0 < 1 \ (\text{attracting}), \qquad |g'(1)| = 2 > 1 \ (\text{repelling}). \]
\end{examplebox}

\subsection{The Banach Fixed Point Theorem}

The deepest reason iteration converges is \emph{contraction}: a map that shrinks
the gap between any two points. On a complete space such a map cannot help but
funnel everything to a single point.

\begin{formulabox}[Contraction Mapping and Banach's Conclusion]
$g$ is a \textbf{contraction} with constant $0 \le k < 1$ if
\[ |g(x) - g(y)| \le k\,|x - y| \quad \text{for all } x,y. \]
On a \textbf{complete} space mapped into itself, Banach guarantees a
\textbf{unique} fixed point $x^*$, reached by iteration from \emph{any} start,
with the error bound
\[ |x_n - x^*| \le \frac{k^n}{1-k}\,|x_1 - x_0|. \]
\end{formulabox}

\begin{warningbox}[Common Pitfall: $k=1$ Is Not Enough]
The constant must be \emph{strictly} below one. With $k=1$ distances need not
shrink, so the iterates may never converge --- e.g. $g(x)=x+1$ satisfies
$|g(x)-g(y)|=|x-y|$ ($k=1$) and has \emph{no} fixed point at all.
\end{warningbox}

\begin{examplebox}[{Worked Example 3.5: $g(x)=\cos x$ is a contraction on $[0,1]$}]
\textbf{Step 1.} A bound $|g'(x)| \le k < 1$ makes $g$ a contraction. Here
\[ |g'(x)| = |\sin x|, \quad \text{increasing on } [0,1]. \]
\textbf{Step 2.} Its largest value on the interval is at $x=1$:
\[ |\sin x| \le \sin 1 \approx 0.841 < 1. \]
So $g$ is a contraction, and iterating cosine converges to its unique fixed
point (the Dottie number, $\approx 0.739$) from any start.
\end{examplebox}

\subsection{Picard Iteration}

Picard iteration is Banach's principle in action on differential equations.
First the IVP is recast as an \emph{integral equation}; then repeated
substitution builds the solution as a limit.

\begin{formulabox}[Integral Equation and the Picard Scheme]
The IVP $y' = f(x,y)$, $y(x_0)=y_0$ is equivalent to
\[ y(x) = y_0 + \int_{x_0}^{x} f\big(t, y(t)\big)\,dt. \]
Iterate from $\varphi_0 = y_0$:
\[ \varphi_{n+1}(x) = y_0 + \int_{x_0}^{x} f\big(t, \varphi_n(t)\big)\,dt. \]
\textbf{Picard--Lindel\"of:} if $f$ is continuous and \textbf{Lipschitz} in $y$
($|f(x,y_1)-f(x,y_2)| \le L\,|y_1-y_2|$), the operator is a contraction, so the
solution exists and is unique on some interval.
\end{formulabox}

\begin{examplebox}[Worked Example 3.6: Picard iterates for $y'=y,\ y(0)=1$]
\textbf{Step 1.} Start at $\varphi_0 = 1$ and integrate:
\[ \varphi_1 = 1 + \int_0^x 1\,dt = 1 + x. \]
\textbf{Step 2.} Feed it back:
\[ \varphi_2 = 1 + \int_0^x (1+t)\,dt = 1 + x + \frac{x^2}{2}. \]
\textbf{Step 3.} The iterates are the partial sums of a familiar series:
\[ \varphi_n \to 1 + x + \frac{x^2}{2} + \cdots = e^{x}, \]
which is exactly the true solution.
\end{examplebox}

\end{document}
