% Unit 10: Nonhomogeneous Equations and Forced Response
% Master Unit Reference Guide (Cheat Sheet, Problems, and Solutions)
% Prepared by Staples Education
% Content informed by the local curriculum modules (10.1 to 10.8) and the
% QUIZ_DATA micro-practice items for Unit 10.
% Shared style harness pulled from the unified preamble input.

\input{app/assets/markdowns/preamble.tex}

\begin{document}

% ---------------------------------------------------------------------------
% Title block
% ---------------------------------------------------------------------------
\begin{center}
    {\LARGE \bfseries Unit 10: Nonhomogeneous Equations and Forced Response}\\[6pt]
    {\large Prepared by Staples Education}\\[4pt]
    {\itshape Master Unit Reference Guide: Cheat Sheet, Problems, and Solutions}
\end{center}

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Table of contents (master guide)
% ---------------------------------------------------------------------------
\tableofcontents

\vspace{0.4em}

% ---------------------------------------------------------------------------
% Topics overview box
% ---------------------------------------------------------------------------
\begin{topicsbox}[Unit 10 at a Glance: Core Sub-Topics]
A nonhomogeneous linear equation reads $L[y] = g(x)$, and its general solution
always splits as $y = y_h + y_p$ --- the complementary solution of $L[y]=0$ plus
\emph{any} particular solution of the forcing. This unit is the toolkit for
building $y_p$ and for reading how a system answers an input.
\begin{itemize}[leftmargin=1.4em, itemsep=3pt]
    \item \textbf{Undetermined coefficients} --- guess a trial of the same shape as
          the forcing, then match coefficients.
    \item \textbf{Variation of parameters} --- promote the constants $c_1,c_2$ to
          functions for \emph{any} continuous forcing.
    \item \textbf{The annihilator method} --- an operator that kills $g$ derives
          the trial form instead of memorizing it.
    \item \textbf{Transient and steady-state} --- the homogeneous part decays, the
          particular part persists.
    \item \textbf{Unit step and impulse response} --- $u(t)$, $\delta(t)$, and the
          system's answer to a unit kick from rest.
    \item \textbf{Duhamel's principle} --- convolve the impulse response with the
          input to answer \emph{every} forcing at once.
\end{itemize}
\end{topicsbox}

\vspace{0.6em}
\hrule
\vspace{0.6em}

% ===========================================================================
% PART I: CONDENSED CHEAT SHEET
% ===========================================================================
\phantomsection
\addcontentsline{toc}{section}{Part I --- Condensed Cheat Sheet}
\section*{Part I --- Condensed Cheat Sheet}
\setcounter{section}{10}   % subsections auto-number as the unit modules 10.1, 10.2, ...

\subsection{The Method of Undetermined Coefficients}

For a constant-coefficient $L[y] = g(x)$, the general solution is
$y = y_h + y_p$: the complementary solution $y_h$ solves $L[y]=0$ and carries the
arbitrary constants, while $y_p$ matches the forcing. When $g$ is an exponential,
polynomial, sine/cosine, or a product of these, guess a trial of the \emph{same
shape} and solve for its coefficients.

\begin{formulabox}[Complementary Plus Particular]
\[ y = y_h + y_p, \qquad L[y_h] = 0, \qquad L[y_p] = g(x). \]
Only $y_h$ holds the free constants, so it alone can absorb the initial or
boundary conditions --- which is why $y_p$ is never the whole story.
\end{formulabox}

\begin{formulabox}[Trial Solution Table (No Overlap)]
\[
\begin{array}{l|l}
\textbf{Forcing } g(x) & \textbf{Trial } y_p \\ \hline
e^{\alpha x} & A e^{\alpha x} \\
\text{polynomial of degree } n & A_n x^n + \cdots + A_1 x + A_0 \\
\cos\beta x \ \text{ or } \ \sin\beta x & A\cos\beta x + B\sin\beta x \\
e^{\alpha x}\cos\beta x & e^{\alpha x}(A\cos\beta x + B\sin\beta x) \\
\end{array}
\]
For a \emph{sum} of forcings, build a particular solution for each piece and add
them (superposition for the nonhomogeneous equation).
\end{formulabox}

\begin{warningbox}[Common Pitfall: Overlap with the Homogeneous Solution]
If any term of the trial already solves $L[y]=0$, substitution collapses it to
zero and the method fails. \textbf{Multiply the offending trial by $x$} (or by a
higher power $x^k$ for a root of multiplicity $k$) until no term overlaps $y_h$.
For $y''+y=\cos x$, the bare guess $A\cos x + B\sin x$ \emph{is} the homogeneous
solution; the corrected resonant trial is $x(A\cos x + B\sin x)$.
\end{warningbox}

\subsection{Variation of Parameters}

When the forcing is not one of the friendly shapes --- $\tan x$, $\sec x$,
$e^x/x$ --- undetermined coefficients has nothing to guess. Variation of
parameters promotes the constants $c_1,c_2$ of $y_h = c_1 y_1 + c_2 y_2$ to
functions $u_1(x),u_2(x)$ and works for \emph{any} continuous $g$.

\begin{formulabox}[Variation of Parameters Formulas]
Write the equation in standard form $y'' + p y' + q y = g$ (leading coefficient
one), let $W = y_1 y_2' - y_2 y_1'$ be the Wronskian, and set $y_p = u_1 y_1 + u_2 y_2$ with
\[ u_1' = -\frac{y_2\, g}{W}, \qquad u_2' = \frac{y_1\, g}{W}, \]
\[ y_p = -\,y_1\!\int \frac{y_2\, g}{W}\,dx \;+\; y_2\!\int \frac{y_1\, g}{W}\,dx. \]
\end{formulabox}

\begin{warningbox}[Common Pitfall: Forgetting to Normalize]
The $g$ in these formulas is the right-hand side \emph{after} dividing through so
the $y''$ coefficient is one. Reading $g$ off an unnormalized equation scales the
answer incorrectly. Always normalize first, then read $g$.
\end{warningbox}

\subsection{Differential Operators and the Annihilator Method}

Writing $D = \tfrac{d}{dx}$, a constant-coefficient operator factors like a
polynomial. An \textbf{annihilator} of $g$ is an operator $A(D)$ with $A(D)g = 0$.
Applying it to both sides of $L[y]=g$ turns the problem into a larger
\emph{homogeneous} equation whose extra solutions are exactly the trial form.

\begin{methodbox}[Annihilator Workflow]
\begin{enumerate}[leftmargin=1.6em, itemsep=2pt]
    \item Factor $L(D)$ and write the complementary solution $y_h$.
    \item Pick the annihilator $A(D)$ of the forcing $g$.
    \item Apply $A(D)$: solve the homogeneous equation $A(D)L(D)\,y = 0$.
    \item Drop the terms already in $y_h$; the survivors are the trial $y_p$.
    \item Substitute into $L[y]=g$ and match coefficients.
\end{enumerate}
\end{methodbox}

\begin{formulabox}[Common Annihilators]
\[
\begin{array}{l|l}
\textbf{Function} & \textbf{Annihilator} \\ \hline
e^{\alpha x} & (D-\alpha) \\
x^{k-1} & D^{k} \\
\cos\beta x,\ \sin\beta x & (D^2 + \beta^2) \\
e^{\alpha x}\cos\beta x,\ e^{\alpha x}\sin\beta x & (D-\alpha)^2 + \beta^2 \\
\end{array}
\]
A repeated factor $(D-\alpha)^k$ signals the terms
$e^{\alpha x}, x e^{\alpha x}, \ldots, x^{k-1}e^{\alpha x}$. Annihilate a sum by
\emph{multiplying} the individual annihilators.
\end{formulabox}

\subsection{Transient and Steady-State Response}

For a \emph{stable} forced system (characteristic roots with negative real
parts), the complete solution $y = y_h + y_p$ separates by long-term behavior: the
homogeneous part fades while the particular part endures.

\begin{formulabox}[Transient Versus Steady-State]
\[ y(t) = \underbrace{y_h(t)}_{\text{transient}\ \to\ 0} \;+\; \underbrace{y_p(t)}_{\text{steady-state}}. \]
The \textbf{transient} carries the initial-condition dependence and decays because
its exponentials have negative real parts. The \textbf{steady-state} is fixed by
the forcing alone, so long after the start the response is essentially independent
of the initial data. For sinusoidal forcing the steady state is a sinusoid at the
\emph{drive} frequency, rescaled and phase-shifted, often read off the complex
gain $1/p(i\omega)$.
\end{formulabox}

\subsection{Unit Step and Impulse Response}

Two idealized inputs model switching and striking. The unit step $u(t)$ switches a
forcing on; the Dirac delta $\delta(t)$ is an instantaneous unit kick. They are
linked by calculus, and so are the responses they produce.

\begin{formulabox}[Step, Delta, and Their Responses]
\[ u(t) = \begin{cases} 0, & t<0 \\ 1, & t>0 \end{cases}, \qquad
   \delta(t) = u'(t), \qquad \int_{-\infty}^{\infty}\!\delta(t)\,dt = 1. \]
\textbf{Sifting property:} $\displaystyle\int_{-\infty}^{\infty} f(t)\,\delta(t-a)\,dt = f(a)$.
The \textbf{impulse response} $w(t)$ is the output of $L[y]=\delta(t)$ from rest;
the \textbf{step response} is its integral. Knowing $w$ unlocks every input by
convolution.
\end{formulabox}

\begin{warningbox}[Common Pitfall: Forgetting the Rest State]
The impulse response is defined from \textbf{rest} --- zero state just before the
kick. Starting with leftover energy mixes in a transient that is unrelated to the
impulse, so the clean $w(t)$ (here $w(0)=0$, $w'(0)=1$ for a second-order system)
is lost. Isolate the impulse by zeroing the initial conditions first.
\end{warningbox}

\subsection{Duhamel's Principle}

Duhamel's principle views a general input as a continuous succession of small
impulses. Each kick at time $\tau$ produces a scaled, time-shifted copy of the
impulse response, and superposition sums them into a convolution.

\begin{methodbox}[Duhamel's Convolution]
For a linear, time-invariant system starting from rest, the forced response to an
input $f(t)$ is
\[ y_p(t) = \int_0^t w(t-\tau)\,f(\tau)\,d\tau, \]
where $w$ is the impulse response (the Green's function of the initial-value
problem). The limits $0$ to $t$ encode \textbf{causality} --- only inputs already
applied can affect the present --- and the argument $t-\tau$ is the time elapsed
since each kick. \textbf{Linearity} and \textbf{time invariance} are exactly what
make the convolution valid.
\end{methodbox}

\begin{conceptbox}[Unit 10 Key Takeaway]
Every forced response is \textcolor{accentorange}{$y = y_h + y_p$}. Build $y_p$ by
\textcolor{accentorange}{undetermined coefficients} when the forcing is friendly
(remember to \textcolor{accentorange}{multiply by $x$} on overlap), by
\textcolor{accentorange}{variation of parameters} for anything continuous, or read
its form from an \textcolor{accentorange}{annihilator}. In a stable system the
\textcolor{accentorange}{transient} $y_h$ decays and the
\textcolor{accentorange}{steady-state} $y_p$ remains, and once the
\textcolor{accentorange}{impulse response $w$} is known,
\textcolor{accentorange}{Duhamel's convolution} answers every input at once.
\end{conceptbox}

% ===========================================================================
% PART II: REVIEW GUIDE (PRACTICE QUESTIONS)
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part II --- Review Guide: Practice Problems}
\section*{Part II --- Review Guide: Practice Problems}

Work the following ten problems in order. They climb through undetermined
coefficients (exponential, polynomial, sinusoidal, and the two resonance traps),
the annihilator method, variation of parameters for unfriendly forcing, the
transient/steady-state split, and a capstone on the impulse response and
Duhamel's convolution. Solutions follow in Part III.

\begin{enumerate}[leftmargin=2em]
    \item \textbf{(Undetermined coefficients --- exponential)} Solve
          $y'' - y = e^{2x}$. Give $y_h$, the trial, and the general solution.

    \item \textbf{(Undetermined coefficients --- polynomial)} Solve
          $y'' - y' - 2y = 4x^2$. Use a full quadratic trial.

    \item \textbf{(Undetermined coefficients --- sinusoid)} Solve
          $y'' + 4y = \sin x$. Explain why no factor of $x$ is needed here.

    \item \textbf{(Resonance, trig)} Solve $y'' + y = \cos x$. Identify the overlap
          with the homogeneous solution and use the corrected resonant trial.

    \item \textbf{(Resonance, exponential)} Solve $y'' - 3y' + 2y = e^{x}$.
          Show why the guess gains a factor of $x$.

    \item \textbf{(Annihilator method)} For $y'' - y' - 2y = 2e^{3x} + x$, write the
          annihilator of the forcing, deduce the trial form, and solve for $y_p$.

    \item \textbf{(Variation of parameters)} Solve $y'' + y = \sec x$, where
          undetermined coefficients cannot be applied.

    \item \textbf{(Variation of parameters, repeated root)} Find a particular
          solution of $y'' - 2y' + y = \dfrac{e^{x}}{x}$ for $x>0$.

    \item \textbf{(Transient and steady-state IVP)} Solve
          $y'' + 4y' + 3y = 6$, $y(0)=0$, $y'(0)=0$. Identify the transient and the
          steady-state pieces.

    \item \textbf{(Impulse response and Duhamel, capstone)} Find the impulse
          response $w(t)$ of $y'' + 3y' + 2y = f(t)$ (rest initial conditions), then
          write Duhamel's convolution for a general input $f$.
\end{enumerate}

% ===========================================================================
% PART III: SOLUTIONS
% ===========================================================================
\clearpage
\phantomsection
\addcontentsline{toc}{section}{Part III --- Complete Solutions}
\section*{Part III --- Complete Solutions}

\noindent\textbf{Solution 1.}\quad The homogeneous equation $y''-y=0$ has roots
$r=\pm 1$, so $y_h = c_1 e^{x} + c_2 e^{-x}$. Since $e^{2x}$ does not overlap
$y_h$, try $y_p = A e^{2x}$, so $y_p'' = 4A e^{2x}$:
\[ \textcolor{workpurple}{4A e^{2x} - A e^{2x} = 3A e^{2x} = e^{2x} \;\Rightarrow\; A = \tfrac{1}{3}.} \]
\[ \textcolor{workpurple}{y = c_1 e^{x} + c_2 e^{-x} + \tfrac{1}{3} e^{2x}.} \]

\medskip
\noindent\textbf{Solution 2.}\quad Roots of $r^2 - r - 2 = (r-2)(r+1)=0$ give
$y_h = c_1 e^{2x} + c_2 e^{-x}$. Use the complete quadratic trial
$y_p = Ax^2 + Bx + C$, with $y_p' = 2Ax+B$, $y_p'' = 2A$:
\[ \textcolor{workpurple}{2A - (2Ax+B) - 2(Ax^2+Bx+C) = -2A x^2 + (-2A-2B)x + (2A - B - 2C).} \]
Matching $-2Ax^2 + \cdots = 4x^2$ term by term:
\[ \textcolor{workpurple}{-2A = 4 \Rightarrow A=-2; \quad -2A-2B = 0 \Rightarrow B = 2; \quad 2A - B - 2C = 0 \Rightarrow C = -3.} \]
\[ \textcolor{workpurple}{y = c_1 e^{2x} + c_2 e^{-x} - 2x^2 + 2x - 3.} \]

\medskip
\noindent\textbf{Solution 3.}\quad Here $y''+4y=0$ has roots $\pm 2i$, so
$y_h = c_1\cos 2x + c_2 \sin 2x$. The forcing $\sin x$ has frequency $1\neq 2$, so
there is \emph{no} overlap and no factor of $x$ is needed. Try
$y_p = A\cos x + B\sin x$; then $y_p'' = -A\cos x - B\sin x$:
\[ \textcolor{workpurple}{(-A + 4A)\cos x + (-B + 4B)\sin x = 3A\cos x + 3B\sin x = \sin x.} \]
So $A = 0$, $B = \tfrac{1}{3}$, giving
\[ \textcolor{workpurple}{y = c_1\cos 2x + c_2\sin 2x + \tfrac{1}{3}\sin x.} \]

\medskip
\noindent\textbf{Solution 4.}\quad For $y''+y=0$ the roots are $\pm i$, so
$y_h = c_1\cos x + c_2\sin x$. The natural guess $A\cos x + B\sin x$ \emph{is}
$y_h$ --- a resonance. Multiply by $x$: $y_p = x(A\cos x + B\sin x)$. Differentiating,
$y_p'' = -x(A\cos x + B\sin x) + 2(-A\sin x + B\cos x)$, so
\[ \textcolor{workpurple}{y_p'' + y_p = 2(-A\sin x + B\cos x) = \cos x \;\Rightarrow\; B = \tfrac{1}{2},\ A = 0.} \]
\[ \textcolor{workpurple}{y = c_1\cos x + c_2\sin x + \tfrac{x}{2}\sin x.} \]

\medskip
\noindent\textbf{Solution 5.}\quad Roots of $r^2 - 3r + 2 = (r-1)(r-2)=0$ give
$y_h = c_1 e^{x} + c_2 e^{2x}$. The forcing $e^{x}$ overlaps the $e^{x}$ solution,
so the bare guess $Ae^{x}$ fails; multiply by $x$: $y_p = A x e^{x}$. Then
$y_p' = A e^{x}(1+x)$ and $y_p'' = A e^{x}(2+x)$, so
\[ \textcolor{workpurple}{A e^{x}\big[(2+x) - 3(1+x) + 2x\big] = A e^{x}(-1) = e^{x} \;\Rightarrow\; A = -1.} \]
\[ \textcolor{workpurple}{y = c_1 e^{x} + c_2 e^{2x} - x e^{x}.} \]

\medskip
\noindent\textbf{Solution 6.}\quad The operator is $L(D) = D^2 - D - 2 = (D-2)(D+1)$,
so $y_h = c_1 e^{2x} + c_2 e^{-x}$. The forcing $2e^{3x} + x$ is annihilated by
\[ \textcolor{workpurple}{A(D) = (D-3)\,D^2,} \]
since $(D-3)$ kills $e^{3x}$ and $D^2$ kills the linear term $x$. Neither $e^{3x}$
nor $\{1, x\}$ overlaps $y_h$, so the trial is $y_p = A e^{3x} + Bx + C$.
Substitute the exponential piece $Ae^{3x}$ ($y''-y'-2y = 9A - 3A - 2A = 4A$):
\[ \textcolor{workpurple}{4A e^{3x} = 2 e^{3x} \;\Rightarrow\; A = \tfrac{1}{2}.} \]
For the polynomial piece $Bx + C$ ($y'' = 0$, $y' = B$):
\[ \textcolor{workpurple}{-B - 2(Bx + C) = -2Bx + (-B - 2C) = x \;\Rightarrow\; B = -\tfrac{1}{2},\ C = \tfrac{1}{4}.} \]
\[ \textcolor{workpurple}{y = c_1 e^{2x} + c_2 e^{-x} + \tfrac{1}{2} e^{3x} - \tfrac{1}{2}x + \tfrac{1}{4}.} \]

\medskip
\noindent\textbf{Solution 7.}\quad The equation is already in standard form with
$y_1 = \cos x$, $y_2 = \sin x$, and Wronskian
$W = \cos x\cos x - \sin x(-\sin x) = 1$. With $g = \sec x$,
\[ \textcolor{workpurple}{u_1' = -\frac{y_2 g}{W} = -\sin x \sec x = -\tan x \;\Rightarrow\; u_1 = \ln|\cos x|,} \]
\[ \textcolor{workpurple}{u_2' = \frac{y_1 g}{W} = \cos x \sec x = 1 \;\Rightarrow\; u_2 = x.} \]
Hence $y_p = u_1 y_1 + u_2 y_2$:
\[ \textcolor{workpurple}{y = c_1\cos x + c_2\sin x + \cos x\,\ln|\cos x| + x\sin x.} \]

\medskip
\noindent\textbf{Solution 8.}\quad The repeated root $r=1$ of $(r-1)^2$ gives
$y_1 = e^{x}$, $y_2 = x e^{x}$, with
$W = e^{x}(e^{x} + x e^{x}) - x e^{x}\cdot e^{x} = e^{2x}$. In standard form
$g = e^{x}/x$, so
\[ \textcolor{workpurple}{u_1' = -\frac{x e^{x}\cdot (e^{x}/x)}{e^{2x}} = -1 \Rightarrow u_1 = -x, \qquad
   u_2' = \frac{e^{x}\cdot (e^{x}/x)}{e^{2x}} = \frac{1}{x} \Rightarrow u_2 = \ln x.} \]
Then $y_p = -x e^{x} + x e^{x}\ln x$; the term $-x e^{x}$ already lives in $y_h$, so
\[ \textcolor{workpurple}{y = (c_1 + c_2 x)e^{x} + x e^{x}\ln x.} \]

\medskip
\noindent\textbf{Solution 9.}\quad Roots of $r^2 + 4r + 3 = (r+1)(r+3)=0$ are
$r=-1,-3$, both negative, so $y_h = c_1 e^{-t} + c_2 e^{-3t}$ is the
\textbf{transient}. A constant forcing calls for $y_p = K$ with $3K = 6$, so the
\textbf{steady-state} is $y_p = 2$. Apply $y(0)=0$, $y'(0)=0$ to
$y = c_1 e^{-t} + c_2 e^{-3t} + 2$:
\[ \textcolor{workpurple}{c_1 + c_2 + 2 = 0, \qquad -c_1 - 3c_2 = 0 \;\Rightarrow\; c_1 = -3c_2.} \]
\[ \textcolor{workpurple}{-3c_2 + c_2 + 2 = 0 \Rightarrow c_2 = 1,\ c_1 = -3, \qquad y = -3e^{-t} + e^{-3t} + 2.} \]
As $t\to\infty$ the transient $-3e^{-t}+e^{-3t}\to 0$, leaving the steady-state
value $2$.

\medskip
\noindent\textbf{Solution (Capstone).}\quad The impulse response solves
$w'' + 3w' + 2w = \delta(t)$ from rest, which for a second-order system means
$w(0)=0$, $w'(0)=1$. Roots of $r^2+3r+2=(r+1)(r+2)$ are $r=-1,-2$, so
$w = A e^{-t} + B e^{-2t}$:
\[ \textcolor{workpurple}{w(0) = A + B = 0, \qquad w'(0) = -A - 2B = 1 \;\Rightarrow\; B = -1,\ A = 1.} \]
\[ \textcolor{workpurple}{w(t) = e^{-t} - e^{-2t}, \quad t \ge 0.} \]
By Duhamel's principle the forced response from rest to any input $f$ is
\[ \textcolor{workpurple}{y_p(t) = \int_0^t w(t-\tau)\,f(\tau)\,d\tau
   = \int_0^t \Big(e^{-(t-\tau)} - e^{-2(t-\tau)}\Big) f(\tau)\,d\tau.} \]

\vspace{0.8em}
\begin{conceptbox}[Unit 10 Key Takeaway]
The shape of the forcing chooses the tool:
\textcolor{accentorange}{undetermined coefficients} for exponentials,
polynomials, and sinusoids --- with the \textcolor{accentorange}{$\times x$
resonance fix} whenever a trial term solves $L[y]=0$ ---
\textcolor{accentorange}{variation of parameters} for $\sec x$, $e^x/x$, and other
unfriendly inputs, and the \textcolor{accentorange}{annihilator} to derive the
trial form. In a stable system the \textcolor{accentorange}{transient} decays to
the \textcolor{accentorange}{steady-state}, and the
\textcolor{accentorange}{impulse response $w(t)=e^{-t}-e^{-2t}$} feeds
\textcolor{accentorange}{Duhamel's convolution} to resolve every input at once.
\end{conceptbox}

\end{document}
